### 2011-11-04 doc/pldi2012/mapping_fig.tex
very rough draft of type system section - dummy figure in place for mapping description

\begin{figure}[t]
  \centering
  \subfigure[mapping of $dc_0$ task]{
    \label{sfig:mapping_fig:dc}
    \begin{tikzpicture}
      \partitiontree
    \end{tikzpicture}
  }
  \subfigure[mapping of $volt_0$ task]{
    \label{sfig:mapping_fig:volt}
    \begin{tikzpicture}
      \partitiontree
    \end{tikzpicture}
  }
  \subfigure[mapping completed]{
    \label{sfig:mapping_fig:done}
    \begin{tikzpicture}
      \partitiontree
    \end{tikzpicture}
  }
  \label{fig:mapping_fig}
  \caption{Mapping of tasks in region tree -- TODO: draw in little boxes and stuff}
\end{figure}