### 2011-11-04 doc/pldi2012/type_system.tex

\section{Type System}

The key to both performance and correctness in the Legion system is the
accuracy of the region usage declarations.  The runtime will place data and
schedule tasks based on these declarations, and has no way to recover if a 
task attempts to access data that lies outside the mapped regions or attempts
to write to a region that the runtime expected would only be read.

The guarantee that these region usage declarations will be adhered to is
provided by Legion's type system which is able to statically verify that a 
task will stay within its declared boundaries.  The Legion type system is
flow-sensitive, and uses judgements of the form:

\begin{center}
$\Gamma, \Phi_i, C_i \vdash e : T, \Phi_o, C_o$
\end{center}

In addition to the mapping of variables to types (i.e. $\Gamma$), the
environment includes a set of \emph{privileges} $\Phi_i$ that are possessed
and a set of \emph{constraints} $C_i$ that are known to hold before the 
expression $e$ is evaluated.  The judgement says that $e$ has some type $T$,
but also tells us that the privileges possessed after the
evaluation are a potentially-different $\Phi_o$ and that the new set of 
constraints is $C_o$.

Table~\ref{tbl:priv_const} shows the form of the individual privileges and
constraints that make up $\Phi$ and $C$.  Privileges represent the ability to
perform some operation on a region (or, as we will see, on any of its
subregions).  Constraints are used to capture the subregion relationships that
result from partitioning ($a \prec b$ if $a$ was a result of a partitioning of
$b$), the more general anywhere-in-the-subregion-tree relationship ($a \le b$ if $a$ equals $b$ or is either a direct or indirect subregion), and equality (or
inequality) of integer expressions.  More complicated constraints can be formed
through conjunctions, but neither negation nor disjunctions are permitted.  For
an example of how these work, here is the typing rule for reading an element
from a region:

\begin{center}
{\small
\begin{math}
\infrule{
\begin{array}{lc}
  \multicolumn{2}{c}{\Gamma, \Phi_1, C_1 \vdash p : T@r_1, \Phi_2, C_2} \\
  C_2 \models r_1 \le r_2 & readable(r_2) \in \Phi_2
%  \envsub{1}{1} e_1 : \rtripsub{T_1}{2}{2} \\
%  \envsub{2}{2} e_2 : \rtripsub{T_2}{3}{3} 
\end{array}
}
{
  \Gamma, \Phi_1, C_1 \vdash read(r_2, p) : T, \Phi_2, C_2
%  \envsub{1}{1} e_1; e_2 : \rtripsub{T_2}{3}{3}
}
\end{math} 
}
\end{center}

This rule tells us that in order to read from a pointer, we must be able to
provide a region ($r_2$ in this case) for which we possess read privilege that 
is known to either be, or contain, the region to which the pointer is
constrained.  Similar rules exist for the other four region access operations.

The Legion type system also has to be able to express the self-referential
nature of region relationships.  It does this through existential quantification
over regions.  The quantifiers captures both types and constraints, but does
not capture privileges.  Region relations have the following form in the
type system:

\begin{center}
$RR = \exists r_1, \ldots, r_n.\left(T, \emptyset, C\right)$
\end{center}

Explicit pack and unpack operations are used in the type system
to add and remove the existential quantifiers, shown here:

\begin{center}
\begin{math}
\begin{array}{c}
\infrule{
\begin{array}{l}
RR = \exists r_1, \ldots, r_n.\left(T_1, \emptyset, C_1\right) \\
\Gamma, \Phi_2, C_2 \vdash e : \left(RR, \Phi_3, C_3\right) \\
r'_1, \ldots, r'_n \not\in \mathit{RegionsOf}\left(\Gamma, \Phi_3, C_3\right)
\end{array}
}{
\begin{array}{l@{}l}
\Gamma, \Phi_2, C_2 \vdash unpackrr{ }e : ( & [r'_1/r_1,\ldots,r'_n/r_n]T_1, \Phi_3, \\
& C_3 \wedge [r'_1/r_1,\ldots,r'_n/r_n]C_1)
\end{array}
}
\\
\\
\infrule{
\begin{array}{l}
RR = \exists r_1, \ldots, r_n.\left(T_1, \emptyset, C_1\right) \\
\Gamma, \Phi_2, C_2 \vdash e : \left([r'_1/r_1,\ldots,r'_n/r_n]T_1, \Phi_3, C_3\right) \\
C_3 \models [r'_1/r_1,\ldots,r'_n/r_n]C_1
\end{array}
}{
\begin{array}{l}
\Gamma, \Phi_2, C_2 \vdash packrr{ }RR\mbox{ }e : \left(RR, \Phi_3, C_3\right) \\
\end{array}
}
\end{array}
\end{math}
\end{center}

Any expression whose type is a region relationship can be unpacked.  Doing so
introduces a fresh region variable for each region that was bound in the 
quantification.  The fresh region variables are substitued into the region
relaionship's type, and a corresponding substituion is performed on the
constraints that were captured by the region relationship.  These substituted
constraints are added to the current function's (flow-sensitive) constraints.

Packing an expression into a region relationship works in reverse.  If the type
of an expression can be unified with the region relationship's type and the
corresponding constraints of the region relationship can be shown to hold, the
expression can be packed into the region relationship and used interchangably
with any other instance of that region relationship.

Although the type system uses explicit packing and unpacking operations, they 
are implicit in the Legion application code.  Unpacking operations are 
automatically inserted whenever a variable comes into scope.  Packing operations
are attemped whenever a variable is written to memory or passed as an argument
to a function.  Additionally, the simultaneous assignment operation essentially
performs a pack for the result of the assignments and will also perform an
unpack if one ore more fields of the structure are left unchanged.

\begin{table}
\centering
{\small
\begin{math}
\begin{array}{cc}
\begin{tabular}{ccc}
$\phi$ & ::= & $readable(r)$ \\
  &$\mid$&$writeable(r)$ \\
  &$\mid$&$reduceable(r,f)$ \\
  &$\mid$&$allocable(r)$ \\
  &$\mid$&$freeable(r)$ \\
\\
$\Phi$ & ::= & $\{ \phi_1, \ldots, \phi_n \}$
\end{tabular}
&
\begin{array}{ccc}
C & ::= & (r_1 \prec r_2) \\
  &\mid&(r_1 \le r_2) \\
  &\mid&(r_1 * r_2) \\
  &\mid&(i_1 = i_2) \\
  &\mid&(i_1 \ne i_2) \\
  &\mid&(C_1 \wedge C_2) \\
\end{array}
\end{array}
\end{math}
}
\label{tbl:priv_const}
\caption{Privileges and Constraints}
\end{table}

Region access privileges cannot be stored or packed.  Instead, they follow the
function call tree, and it is this property that allows the runtime to make
its scheduling decisions in a distributed way.  The original source of any
privilege for a given region is the \emph{newrr} operation that created that
region, but the way in which privileges can be provided to a subtask, or
returned to a parent task, is described in the function application rule:

\begin{center}
\begin{math}
\begin{array}{c}
\infrule
{
\begin{array}{lc}
  \Gamma, \Phi_1, C_1 \vdash e_1 : \left(\left(T_i, \Phi_i, C_i\right) \rightarrow \exists r_1, \ldots, r_n . \left(T_o, \Phi_o, C_o\right), \Phi_2, C_2\right) \\
  \Gamma, \Phi_2, C_2 \vdash e_2 : \left(T_i, \Phi_3, C_3\right) \\
  r'_1, \ldots, r'_n \not\in \mathit{RegionsOf}(\Gamma, \Phi_3, C_3) \hspace{1cm} 
  \Phi_i \subset \Phi_3 \hspace{1cm}
  C_3 \models C_i
\end{array}
}
{
\begin{array}{l@{}l}
%  \envsub{1}{1} e_1 e_2 : \rtriple{& \regionexpand T_o}{\left(\Phi_3 \cup \regionexpand \Phi_o\right)}{\\ & \left(C_3 \wedge \regionexpand C_o \wedge \bigwedge_{\substack{{\footnotesize r_f \in \{ r | \regionexpand C_o \models fresh(r) \} } \\ {\tiny r_o \in \mathit{RegionsOf}(\Gamma, \Phi_3, C_3)}}} r_f * r_o\right)}
  \Gamma, \Phi_1, C_1 \vdash e_1 e_2 : \Big( & [r'_1/r_1,\ldots,r'_n/r_n]T_o, \\
 & \left(\Phi_3 \cup [r'_1/r_1,\ldots,r'_n/r_n] \Phi_o\right) , \\
 & \left(C_3 \wedge [r'_1/r_1,\ldots,r'_n/r_n] C_o\right) \Big)
\end{array}
}
\end{array}
\end{math}
\end{center}

In this rule, expression $e_1$ has the type of a Legion function.  In addition
to the function's input and output types, it specifies the privileges and
constraints that must exist at the point where the function is called.  It also
describes the privileges and constraints that exist as post-conditions of the
function invocation.  The output type, privileges, and constraints can be
captured in an existential quantification, allowing a function to return one
or more regions that were previously unknown to the caller.