### 2011-11-06 doc/pldi2012/exec.tex

\section{Execution Model}
\label{sec:exec}

In this section we present the Legion execution model and show that
it is correct in that it is guaranteed to preserve the sequential
execution semantics of programs.  The correctness argument is
instructive in that the properties that ensure correctness are also
exploited in the implementation of the Legion high-level runtime
system to achieve high performance and highly distributed scheduling
of parallel tasks.

For the purposes of describing and reasoning about the
execution model we consider an extremely restricted core language, so
simple in fact that it is not even Turing complete and is also very
inconvenient for writing programs.  The loss in expressive power, however,
is hopefully a gain in clarity and simplicity for illustrating the core ideas in
Legion.    We consider programs written in the following language:

\begin{eqnarray*}
P & := & F_1 \ldots F_n \\[.1in]
F & := & \mbox{\tt def}\ f(r_1,\ldots,r_n) =   s_1; \ldots s_n; \\[.1in]
s & := & f(r_1,\ldots,r_n) \\
&| & \deref r = x \\
& | & x = \deref r \\
&| & (r_1, r_2)_{p} = \mbox{\tt partition}(r) \\ 
%&| & \mbox{\tt close}\ p 
\end{eqnarray*}

There are four kinds of names: functions $f$, regions $r$, local
variables $x$, and partitions $p$.  A function takes only regions as
arguments and a function body consists of an ordered list of
statements.  The core language has no permissions or coherence annotations
on functions; for simplicity, in this section we simply use exclusive access in the case
that a function writes a region.  The statement $f(r_1,\ldots,r_n)$ calls a function, which
is executed for its effects on the region arguments.  The statement $x
= \deref r$ reads a value from region $r$ and stores it in local
variable $x$, and the statement $\deref r = x$ stores the value of $x$
in region $r$.

The statement $(r_1, r_2)_{p} = \mbox{\tt partition}(r)$ creates two
{\em subregions} $r_1$ and $r_2$ partitioning the parent region $r$.
All three region names ($r_1$,$r_2$,and $r$) can be used after the
partition.  Regions $r_1$ and $r_2$ are disjoint, meaning that writes
to $r_1$ have no effect on the contents of $r_2$ and vice versa, but
neither $r_1$ nor $r_2$ is disjoint from $r$.  The partition itself
also has a name $p$.  We define a relation $\rleq$ which is the
reflexive, transitive closure of
\[ r_1 \rleq p \ \ \ r_2 \rleq p \ \ \ p \rleq r \]
The $\rleq$ relation defines a {\em region forest} with regions at the
roots, and alternating levels where partitions are always children of
regions and regions are always children of partitions.  If we ignore
partitions, $\rleq$ is just the subregion relationship: $r \rleq r'$
means that $r$ is a subregion (either immmediately or transitively) of
$r'$.  Including partitions also gives information about disjointness.
Let $\lca{r}{r'}$ be the meet of $r$ and $r'$ (the least-common
ancestor of $r$ and $r'$ in the region forest).  If $r \neq r'$ and
$\lca{r}{r'} = p$ then $r$ and $r'$ are disjoint, as $r$ and $r'$ must
be in distinct subregions of the partition $p$ and partitions
guarantee disjoint subregions.  If $\lca{r}{r'} = r''$ then either $r
\rleq r' = r''$ , $r \rleq r = r''$, or $r \leq p$, $r' \leq p'$, and
$p$ and $p'$ are two distinct partitions of $r''$.  In any of these
cases $r$ and $r'$ cannot be proven disjoint using $\rleq$.

%which is used to {\em close} the partition in the $\mbox{\tt close}\
%p$ statement Intuitively, closing a partition reconciles modifications
%to the subregions with the {\em parent} region: $r$ is updated to
%incorporate any changes to $r_1$ and $r_2$.  In our source language
%{\tt close} is not available to the programmer, but is invoked by the
%language implementation to reconcile distributed copies of regions
%whenever necessary, and it is impossible for the program to observe
%discrepencies between a region and its subregions. This toy language,
%then, is closer to the level of ourintermediate representation, which
%is most appropriate for discussing our runtime system.

A program is a list of functions, with the entry point the first
function on the list.  We require that every function (except
the entry point) is called in exactly one place and that the formal
parameters are named identically to the actual parameters in the
call.  Further, local variables, new region names and
partition names in {\tt partition} are chosen to be distinct from all
other names in the program.  These restrictions allow us to avoid
dealing with variable renaming and also mean that each runtime call is
uniquely named by its function name.

The state of an execution step for a program is a map $M$ from the
variable and region names of the program to their values.  We leave the values
abstract, assuming only that each statement performs a deterministic
update of the variables/regions it writes that depends only on the
current state of the regions/variables it reads.
The sets of regions read and written by statements are defined as follows:
\[
\begin{array}{rcl}
\rread{x = \deref r} & = & \{ r \} \\
\rwrite{\deref r = x} & = & \{ r \} \\
\rwrite{(r_1, r_2)_{p} = \mbox{\tt partition}(r)} & = & \{ r_1,r_2,r \} \\[.15in]
\rwrite{f(r_1,\ldots,r_n)} & = & \bigcup_{1 \leq i \leq n}{\rwrite{s_i}} \\
\multicolumn{3}{l}{\ \ \ \ \mbox{where\ {\tt def}}\ f(r_1,\ldots,r_n): =   s_1; \ldots s_n;} \\[.15in]
\rread{f(r_1,\ldots,r_n)} & = & \bigcup_{1 \leq i \leq n}{\rread{s_i}} \\
\multicolumn{3}{l}{\ \ \ \ \mbox{where\ {\tt def}}\ f(r_1,\ldots,r_n): =   s_1; \ldots s_n;} \\[.15in]
\end{array}
\]
In all other cases the read/write set is empty (e.g., $\rread{\deref r = x} = \emptyset$).

The {\em sequential execution order} is the canonical order in which reads, writes, and partitions of regions are performed.
The sequential order is obtained by inlining all function calls---repeatedly
replace function calls $g(r_1,\ldots,r_n)$ by their function bodies until the entry function contains no function calls.
The order of statements in the entry function is then the sequential execution order. 
Two programs $P_1$ and $P_2$ are equivalent, written $P_1 \equiv P_2$, if starting in the same initial state they halt in
the same final state when evaluated in the sequential execution order.

Consider the following program, where reads, writes, and partitions are numbered in the sequential execution order:
{\small 
\begin{tabbing}
\ \ \ \ \ \= {\tt def} \= $\rm f(r_0) =$ \\
1. \>\>   $\rm (r_1,r_2)_a = \mbox{\tt partition}(r_0);$   \\
2. \>\>  $\rm (r_3,r_4)_b = \mbox{\tt partition}(r_1);$ \\
\>\>     $\rm g(r_1); h(r_2); j(r_3); k(r_3,r_4);$ \\
\> {\tt def} $\rm g(r_1) =$   \\
3. \>\>  $\rm \ast r_1 = u;$ \\ \\
\> {\tt def} $\rm h(r_2) =$ \\      
4. \>\>   $\rm \ast r_2 = v;$\\ \\
\> {\tt def} $\rm j(r_3) = $ \\
5.\>\>   $\rm w = \ast r_3;$ \\ \\
\> {\tt def} $\rm k(r_3,r_4) = $\\
6.\>\>   $\rm x = \ast r_3;$ \\
7.\>\>   $\rm (r_5,r_6)_c = \mbox{\tt partition}(r_4);$ \\ 
\>\>     $\rm m(r_5);$ \\ \\
\> {\tt def} $\rm m(r_5) = $ \\
8. \>\>  $\rm \ast r_5 = y;$
\end{tabbing}
}

Consider any two statements $s_1$ and $s_2$.  We say $s_1$ and $s_2$ are {\em independent} if
\[
\begin{array}{l}
   (\forall r_1 \in \rread{s_1} \cup \rwrite{s_1}.\forall r_2 \in \rwrite{s_2}. \exists p. \lca{r_1}{r_2} = p) \wedge \\
   (\forall r_2 \in \rread{s_2} \cup \rwrite{s_2}.\forall r_1 \in \rwrite{s_1}. \exists p. \lca{r_1}{r_2} = p)
\end{array}
\]
The definition captures the usual idea that two statements are independent if whenever one of them is a write they do not access the same memory location, but the notion of ``location'' is generalized to regions and must take into account that two regions
are disjoint only if they are in different components of some partition.

\begin{lemma}
\rm
\label{lem:independence}
If $s_1$ and $s_2$ are independent, then $s_1; s_2 \equiv s_2; s_1$.
\end{lemma}
\begin{proof}
For a contradiction, assume that $s_1$ writes a region $r_1$ and $s_2$ may read from some $r_2$ where $r_1$ and $r_2$
are not disjoint; then reversing the order of the statements clearly violates the claim.  Now $r_1 \in \rwrite{s_1}$
and $r_2 \in \rread{s_2}$, and therefore $\lca{r_1}{r_2} = p$ for some partition $p$.  But then $r_1$ and $r_2$ are disjoint.
The reasoning is symmetric if $s_1$ reads $r_1$ and $s_2$ writes $r_2$ or $s_1$ writes $r_1$ and $s_2$ writes $r_2$.
Since the read-read case cannot violate independence, the result follows.
\end{proof}
We say $s_2$ {\em depends on} $s_1$ if $s_2$ follows $s_1$ in the sequential execution order and the statements are not independent.
\begin{lemma}
\rm
\label{lem:dataflow}
Let $S = s_1; \ldots s_n;$ be the sequential execution order of reads, writes, and partitions and let $S'$ be any permutation of $S$ such that the order of dependent statements is preserved.  Then $S \equiv S'$.
\end{lemma}
\begin{proof}
By induction on the number of swaps needed to transpose $S'$ into $S$.
If $S = S'$ then we are done.  If $S$ and $S'$ are different then
there must be two adjacent statements $s_1; s_2$ in $S'$ that are
reversed with respect to the sequential execution order.  Since the order of
dependent statements is preserved by the permutation, $s_1$ and $s_2$ are also independent.
But then $s_1; s_2 \equiv s_2; s_1$ by Lemma~\ref{lem:independence} and the result follows.
\end{proof}
Lemma~\ref{lem:dataflow} captures the standard idea that any
topological sort of the statements that preserves dependencies also
preserves sequential execution semantics.  While this is the basis for
dataflow parallelism, in the large distributed memory machines we
target instruction-level dataflow parallelism is both too fine-grain
and incurs too much communication, as we must compute a global
dependence graph across the entire program.  In Legion we use a
different notion of what can be executed in parallel that works at the
granularity of functions and also requires analysis only within a function body.  We
first need a few additional definitions.  Define $\context{s}$ to be
the function call statement that invokes the function in which
statement $s$ occurs (here we assume some way of distinguishing
identical statements that occur in different functions).  Let
$\interval{f}$ be the set of statements executed by the one call to
function $f$ and all the functions $f$ transitively calls; we define
$\interval{s} = \{ s \}$ for any statement $s$ other than a function call.
Finally we say statements $s_1$ and $s_2$ are {\em siblings} if they
occur in the same function body.

\begin{lemma}
\rm
\label{lem:locality}
If $s_1$ and $s_2$ are not independent and not siblings, then $s_1$ and $\context{s_2}$ are not independent
and $s_2$ and $\context{s_1}$ are not independent.
\end{lemma}
\begin{proof}
Follows immediately from the fact that the read/writes sets of $\context{s}$ are supersets of the read/write sets of $s$.
\end{proof}
We use Lemma~\ref{lem:locality} in the following way.  Consider two
statements $s_1$ and $s_2$ such that $s_2$ depends on $s_1$.  Let $f$ be the function that is the least
common ancestor of $s_1$ and $s_2$ in the call tree of the
program---the unique function $f$ with the smallest interval such that
$s_1,s_2 \in \interval{f}$.  Then one of the following is true:
\begin{itemize}
\item $s_1$ and $s_2$ are siblings in $f$.

\item $s_1$ occurs in $f$, there is a function call $f_2(\ldots)$ in $f$ such that $s_2 \in \interval{f_2}$, and $f_2(\ldots)$
is dependent on $s_1$.

\item $s_2$ occurs in $f$, there is a function call $f_1(\ldots)$ in $f$ such that $s_1 \in \interval{f_1}$, and
$s_2$ is dependent on $f_1(\ldots)$.

\item There are two distinct function call statements $f_1(\ldots)$ and $f_2(\ldots)$ in the body of $f$ such
that $s_1 \in \interval{f_1}$ and $s_2 \in \interval{f_2}$ and $f_2(\ldots)$ is dependent on $f_1(\ldots)$.
\end{itemize}
In other words, any dependence between arbitrary statements
in different functions can be identified at a coarser granularity as a
dependence between two statements (one or both of which may be a function call) within the same function body.  
In the example program above, statements 3 and 8 are
dependent because they write into regions that are in different
partitions of region $r_0$.  This dependence means that any functions
that include these statements in their interval cannot be independent,
and in particular in the entry point function $f$ the function call
$k(r_3,r_4)$ depends on sibling $g(r_1)$.

The following lemma gives another class of statement orderings based on these observations.
\begin{lemma}
\rm
\label{lem:scheduling}
Let $S$ be the sequential execution order of reads, writes and partitions.  Let $S'$ be any permutation of $S$ such that
for any sibling statements $s_1$ and $s_2$ such that $s_2$ depends on $s_1$, all statements in $\interval{s_1}$ precede
all statements in $\interval{s_2}$.  Then $S \equiv S'$.
\end{lemma}
\begin{proof}
It suffices to show that the order of any two dependent statements $s_1$ and $s_2$ is the same in $S$ and $S'$.
From the discussion above, we know that if $s_2$ depends on $s_1$ then there are two sibling statements $s_1'$ and $s_2'$
such that $s_1 \in \interval{s_1'}$, $s_2 \in \interval{s_2'}$, and $s_2'$ depends
on $s_1'$.
\end{proof}
The advantage of Lemma~\ref{lem:scheduling} is that provided we have
a way to obtain the read and write sets for function calls, scheduling
decisions can be made completely locally for the statements in a
single function body at a time, independently of any decisions for
other functions.  This observation eliminates any global
computation to determine what statements can execute in parallel, and
furthermore allows scheduling decisions theselves to be parallelized, as the statements
in different functions can be scheduled separately.  As a practical matter
our Legion implementation only parallelizes function calls---the other
statements within a function body are executed in sequential
order.

Returning to our example above, in function $f$ the function calls
$g(r_1)$ and $h(r_2)$ are independent, as they work on disjoint
regions $r_1$ and $r_2$.  Similarly $j(r_3)$ and $k(r_3,r_4)$ are
also independent as they both only read from $r_3$.  However, $j$ and $k$
are both dependent on $g$ and $h$, as the earlier functions modify a different
partition of the same data.  Thus Legion will ensure functions $j$ and $k$
begin only after $g$ and $h$ complete, while it will allow $g$ and $h$
to run in parallel, and $j$ and $k$ to run in parallel.