### 2011-11-07 doc/pldi2012/highlevel.tex
high level

\section{High-Level Runtime} 
\label{sec:highlevel}
%In this Section we describe the high-level runtime system
%that is designed to implement the Legion programming
%and execution models while remaining machine agnostic for portability.  
%The high-level runtime system operates on top of the 
%low-level runtime which hides machine specific details
%and is described in Section \ref{sec:lowlevel}.

The major challenge in implementing the high-level
runtime is to avoid the latencies associated with distributed
memory systems and deep memory hierarchies.  To aid
in solving this problem, the high-level runtime employs two useful 
scheduling techniques.  The first, validated in 
Section~\ref{sec:exec}, is to make all scheduling decisions for a
task locally, i.e. in the context in which the
task was created.  The second is a {\em deferred execution model} 
that allows the runtime to hide the latencies associated
with moving data through deep memory hierarchies by decoupling the issuing
of operations
from when they are performed.  This model is supported
by features in the low-level runtime (see Section~\ref{sec:lowlevel}).

%The high-level runtime is responsible for guaranteeing the semantics 
%of programs written in the Legion programming model while at the same
%time extracting high performance.  This goal is challenging because
%the high-level runtime must be capable of operating on top of many 
%different architectures abstracted by the low-level runtime.  This 
%includes those with very large latencies for transferring data such
%as distributed memory clusters.  To hide these latencies, the high
%level runtime uses a deffered model of execution made possible by the
%event model of the low-level runtime.  By deffering execution, the 
%runtime can hide the latencies of data and task movement and acieve
%good throughput and high performance.

A problem with deferred execution is that deferred operations must
still be executed in the correct order if they are dependent.
To ensure correctness the high-level runtime uses an architecture
similar to a pipelined out-of-order hardware processor, but at
the coarser granularity of tasks and regions rather than instructions and registers.
The stages of the task execution pipeline seek to extract
as much task-level parallelism as possible from an application, while
preserving dependencies.  We describe some
details of the four pipeline stages: dependence analysis, mapping, execution, and
task completion.

\subsection{Dependence Analysis} 
\label{subsec:depanalysis}
The first step in the execution of a task is dependence analysis.  Within a task call, subtasks are
registered with the high-level runtime in sequential execution order.  The
registration lists the logical regions to be accessed with needed privileges and coherence.
The runtime uses this information to perform
dependence detection with each of the subtasks already registered
in the same context (i.e. called by the same parent task).  
%These checks will
%be sufficient to ensure correct program order, as was proved in Section~\ref{sec:exec}.

%We initially describe how dependences are detected from logical regions
%in Section \ref{subsec:cohdep}.  We then describe the algorithm that
%is used to perform dependence detection in Section \ref{subsec:depdetect}.
%Section \ref{subsec:dataflow} describes how data is versioned correctly
%during deferred execution.  For concreteness we describe how dependence
%analysis is performed for the circuit simulation from Section \ref{sec:ex}.

%When performing dependence analysis the runtime relies on its knowledge about
%the relationships between logical regions and partitions.  We first describe
%the the components to dependence analysis in sections \ref{subsec:regiontree},
%\ref{subsec:cohdep}, and \ref{absinsts}.  We then describe the algorithm for performing 
%dependence analysis on this data structures in section \ref{subsec:depdetect}.
%\subsubsection{Region Trees}
%\label{subsec:regiontree}
%To detect dependences between the tasks, the runtime leverages its knowledge
%about logical regions and their partitions.  To create a new logical region
%or partition, the application must invoke the corresponding call in the
%runtime.  As these calls occur, the runtime constructs a data strcture
%called a {\em region tree}.  A region tree describes the relationship between
%logical regions and partitions.  A region tree contains two types of nodes:
%region nodes and partition nodes.  Every region tree is rooted with a region
%node and alternates between region and partition nodes at each corresponding level.
%A region node represents a specific logical region and tracks the set of
%partitions of that logical region.  Similarly a partition node represents 
%a specific partitioning of a logical region and keeps track of the logical regions 
%which are subregions of the partition.  
\subsubsection{Dependencies From Coherence and Privileges}
\label{subsec:cohdep}
Conceptually, dependence detection between a new task $t$ and a previous task $t'$ requires comparing the two sets of
logical regions accessed.  For each logical region from $t'$
that may overlap with a logical region needed by $t$, the privileges
and coherence modes are compared to determine whether a dependence exists.  Read
privileges never cause conflicts, but if either 
task needs write or reduction privileges, the coherence modes are compared:

\vspace{2mm}
{\small
\begin{tabular}{c|cccc}
             & Exclusive & Atomic   & Simultaneous & Relaxed \\
\midrule
Exclusive    & Dep & Dep & Dep & Dep \\ 
Atomic       & Dep & Same & Cont & Cont \\
Simultaneous & Dep & Cont & Same & None \\
Relaxed      & Dep & Cont & None & None \\
\end{tabular}
}
\vspace{2mm}

An {\tt Dep} entry indicates a dependence, whereas {\tt None}
means no dependence.  A
{\tt Same} entry indicates a dependence unless the two tasks 
use the same {\em physical region}.  Finally, a {\tt Cont} entry
indicates a dependence unless a single writer is using the
atomic coherence mode.  When a dependence is detected, the new task 
waits for the older task to complete before it begins execution.

The table lists coherence modes we have not yet discussed.  Atomic
allows tasks to be reordered as long as the result is
serializable.  Both simultaneous and relaxed allow other tasks to
execute at the same time.  The difference comes in the restriction of
what data must be observed in the different modes.  In Simultaneous
mode, a task must see all updates to the logical region made by other
tasks operating on the same region at the same time.  In Relaxed mode,
a task is ambivalent as to whether it observes any updates to the
logical region by other tasks executing at the same time.

%For each previously registered task, 
%To detect dependencies between any two tasks using a logical region, we leverage
%the information contained in the privileges and cohrence properties provided by the
%task call.  Dependency checks are performed pair-wise between regions
%to determine whether one task depends on the other.  When checking for
%dependencies we first examine the region privileges.  If both tasks have reads privileges, then 
%there is never a dependence.  If at least one task has write privileges we use table
%\ref{tab:depsdetect} to determine whether there is a dependence.

%In this table, {\em Dependence} indicates that there will always be a dependence between
%regions while {\em None} indicates that there will never be a dependence regardless
%of mapping decisions.  In the case of {\em Same Instance} there is a dependence contingent
%upon a mapping decision.  If both tasks map to the same physical instance (described in 
%more detail in section \ref{subsec:instmang}) then there is no dependence, which if
%they map to different instances there is a dependence as the system will have to generate
%a copy between the two tasks running.
%For the case of {\tt Contingent} there are two special cases that we handle.
%One special case is Write-After-Read (WAR) dependencies.  For the WAR case, if neither coherence property
%is {\em Exclusive} then we can ensure that different physical instances
%are used to avoid an anti-dependence and increase parallelism.  If either property is {\em Exclusive} then the semantics
%of the programming model mandate that we detect a dependence and serialize the task.

%Another special case occurs if there is an {\tt Atomic} writer and either {\tt Simultaneous}
%or {\tt Relaxed} reader.  In this case it is acceptable for the tasks to run in
%parallel as it is valid for the reader to see updates form the writer, and the
%reader will not violate the atomic access semantics of the writer.
%XS\begin{center}
%\begin{table}
%\small
%\begin{tabular}{|c|c|c|c|c|} \hline
%             & Exclusive & Atomic   & Simultaneous & Relaxed \\ \hline 
%Exclusive    & Dep & Dep & Dep & Dep \\ \hline
%Atomic       & Dep & Same & Cont & Cont \\ \hline
%Simultaneous & Dep & Cont & Same & None \\ \hline
%Relaxed      & Dep & Cont & None & None \\ \hline
%\end{tabular}
%\caption{Dependence detection cases when at least one access has a write privilege. {\tt Dep}
%indicates a dependence is detected.  {\tt Same} indicates a dependence is detected
%only if the task uses different instances.  {\tt Cont} means that the dependence
%is contingent upon the permissions.  {\tt None} means there will never
%be a conflict.\label{tab:depsdetect}}
%\end{table}
%\end{center}

%\subsubsection{Abstract Instances}
%\label{subsec:absinsts}
%For the cases where tasks do not depend, they are allowed to create different physical
%instances of the same logical region.  The decision to do this however does not
%happen until the mapping phase of a task execution.  However, we need to have a place
%holder that is capable of representing all the possible valid physical instances
%from which to choose even if those instances do not exist because the prior
%tasks have not been mapped.  We call these place holders {\tt abstract instances}.

%Abstract instances represent the set of possible valid physical instances which a
%task will be capable of selecting from or adding to when it is mapped.  The value
%of abstract instances is that they allow us to remember which physical instances
%are valuable while continuing to perform dependence detection on newly registered
%tasks.  In the analogy to out-of-order processors abstract instances allow us
%to perform the same operation as register-renaming of instructions to avoid
%anti-dependences and to remember actual dependences even though our dependence detection
%analysis may run ahead of the actual execution of tasks.

\subsubsection{Dependence Detection}
\label{subsec:depdetect}

To detect dependencies efficiently, the high-level runtime maintains a
local region forest (recall Section~\ref{sec:ex}) for each task.  The
root regions are the task's region arguments, and the forest includes
all partitions and subregions created within the task.  To perform
dependence detection on a region $r$, the runtime first finds a root
region $r'$ such that $r \rleq r'$.  The runtime then walks the region
forest from $r'$ to $r$.  Each region $r''$ on the path contains a
list $L$ of registered tasks using $r''$, and potential dependencies
are discovered by comparing privileges and coherence modes with the
tasks in $L$.

The runtime next checks if a partition of $r''$ is {\em open}.  An
open partition indicates that there are active tasks using subregions
of $r''$.  No more than one partition of a node can be open
at any time.  If the open partition is is along the path to $r$,
traversal continues.  If the open partition lies along another 
path the runtime must {\em close} the currently active partition because it cannot guarantee
disjointness with the region the new task wishes to access.  To close a
partition, the runtime finds all active tasks in that
subtree and considers them all dependencies for the new task.  
With the previously active partition closed, the runtime can 
then opens the desired partition and traverses further down the tree as
needed.

When the runtime traverses an open partition, its behavior depends on the kind
of partition being traversed.  If the partition is disjoint, no other subregion
can overlap with the target region, so the state of other subregions is
ignored.  However, if the partition is not disjoint, any 
of subregion could overlap with the desired subregion, so
the runtime must close all open subregions (recording new dependencies
as it does so) before continuing the traversal.

After the runtime reaches the target logical region, it performs dependence
checks against any active tasks listed for that region.  If a dependence is 
discovered, it is added to the current task's dependence list.  It can then
be removed from the region tree's active task list, as the current task (which
will be pushed onto the active task list) {\em dominates} the previous task.
In a case where the new task has dependencies with all existing active tasks
for the target node, that whole list is replaced by a single entry for the
new task.

\subsubsection{Managing Data Versions}
\label{subsec:dataflow}

A {\em physical instance} is a copy of the data in a logical region.  Multiple
physical instances of a logical region may exist simultaneously, and it is the responsibility of the high-level runtime
to ensure tasks use the correct version.
In an out-of-order processor this problem is solved by register-renaming which maps logical
registers to a larger set of physical registers.  Our solution employs a similar technique
that maps logical regions to a larger set of {\em abstract instances}.  An abstract instance
is a place holder for the set of physical instances a task will reference when it actually runs.  
Abstract instances are needed because we may not know the actual physical instances a task will
use until its predecessors in the dependence graph are {\em mapped} 
(see Section \ref{subsec:mapping}).  
To determine the abstract instances of a region, we keep track of 
the previous valid abstract instances that were observed while traversing the region forest
during dependence analysis.  Every region in the region forest that has active tasks
has a valid abstract instance.  Every time we close a region we close all the valid
abstract instances.  When a region is reopened it must then create a new abstract instance
to reflect that new physical instances must be created (i.e., a fresh version of the data must be copied from
the parent region).  In addition, whenever a task writes a region, a new abstract instance is usually created
to represent the new version of the data.

%For each region in
%a task we record the source and destination abstract instances that correspond to the physical
%instances from which a task will read its input data and then place its output data.
%When the task is mapped as described in section \ref{subsec:mapping} we will update these
%abstract instances with the actual physical instances chosen.
%In addition to computing the dependences between tasks, we also have to compute the version
%of data that a task is permitted to use when it goes to execute since the computation.
%It is insufficient to simply record the producer task as there
%may be multiple active tasks for a logical region, all of which may produce different physical
%instances.  In addition, it is likely that the producer tasks have not been mapped
%yet which makes it impossible to know the set of physical instances from which to pull data.

%If a region
%subtree must be closed then all the abstract instances in that subtree are marked closed
%so no additional tasks can register them as sources.  We can also chose to create a new
%abstract instance to replace a currently valid abstract instance to avoid a WAR dependence.
%By creating an abstract instance we will force the mapper to choose a different physical
%instance from the existing ones when mapping is performed.

%Abstract instances correspond to a dynamic construction of the dataflow graph.  Each abstract
%instance is a node in the dataflow graph corresponding to a specific version of a logical 
%region and its instantiations as physical regions.  Tasks are then edges between abstract
%instances indicating operations on different versions of data.  By computing this information
%during dependence detection, we can record it before tasks begin executing out of order allowing
%us to guarantee program correctness.  
%Continuing
%the analogy with out-of-order processors, this implicit construction of the dataflow graph
%is annalogous to register renaming which is done in order, but allows the processor to remove
%WAR dependences and execute instructions in parallel that may have had false dependences.

\subsubsection{Circuit Execution Example}
\label{subsec:cirdependence}
To illustrate concretely how dependence analysis works, we describe the steps for
performing dependence analysis for the circuit simulation in Listing~\ref{lst:code_ex}.
For conciseness we will only show the node region forest.  At the start of the task, there exists
a single abstract instance at the root representing the valid set of physical
instances from the parent task. 
When each of the {\tt compute\_new\_current} tasks are issued they must perform dependence
analysis for two node regions.  In the {\tt private\_node} case they will first open
up the {\tt p\_nodes\_pvs} partition and then open up the {\tt p\_pvt\_nodes} sub-partition
for each of their subregions.  We use a bold circle on an edge to indicate that it has been
opened.  Since all of the private regions use a different subregion
of a disjoint partition for the private node accesses, no dependence is detected.

Similarly, when the dependence analysis is performed for the {\tt p\_ghost\_nodes} subregions,
each of the necessary partitions are opened.  There is no dependence with any of the private
nodes since all the ghost nodes traverse a different subregion of the disjoint partition
{\tt p\_nodes\_pvs}.  There is also no dependence between any tasks even though the
{\tt p\_ghost\_nodes} partition is aliased because all the ghost accesses are read-only.
The state of the region forest after dependence analysis has been performed for the first
{\tt compute\_new\_current} task can be seen in figure~\ref{sfig:mapping_fig:cnc}.

In the second phase of the computation, the {\tt distribute\_charge} tasks
are issued.  For these tasks they will first traverse the region forest
for the regions containing their private nodes.  The path to these regions are
already open, but when the dependence analysis arrives at the specific private node subregion it 
discovers a dependence between the task and the {\tt compute\_new\_currents} task that operated
on that region previously.  This dependence is recorded and the 
{\tt distribute\_charge} task becomes the new active task.

For the ghost node regions, dependence analysis discovers 
a dependence between each of the {\tt distribute\_charge} tasks and all of the
{\tt compute\_new\_current} tasks because the {\tt distribute\_charge} tasks are
performing writes and the partition is aliased.  However, the runtime will detect this as
anti-dependences which are resolved by forcing the {\tt distribute\_charge} tasks to
create new valid abstract instances rather than use the same abstract instances as
each of the {\tt compute\_new\_current} tasks.  There are no dependences detected
between each of the {\tt distribute\_charge} tasks on the ghost regions because
the {\tt Simultaneous} coherence property permits multiple tasks to be using the same
partition at the same time even if the subregions they are accessing are aliased.  The
state of the region forest after dependence analysis for the first {\tt distribute\_charge} task can
be seen in figure~\ref{sfig:mapping_fig:dc}.

For the final phase of the computation, all of the {\tt update\_voltage} tasks
are issued requiring their private node region and their shared node region.  Again
the private node paths are already open, but a write-after-read dependence is detected which
requires a new abstract instance to be created to avoid a dependence.  The
shared nodes present a more interesting case.  The first shared node to be registered
detects a dependence with every prior task since it must open the shared partition
instead of the ghost partition.  A new abstract instance is created at the 
{\tt all\_shared} node and copies are recorded for each valid instance in the ghost partition
back to the new {\tt all\_shared} abstract instance.  After this is complete, the
ghost partition can be opened.  Each corresponding {\tt update\_voltage} task will record
a dependence on the copy operation to close ghost partition as it traverses the {\tt all\_shared}
node ensuring that there is no requirement on the ordering of {\tt update\_voltage}
tasks.  The state of the region forest after the first update voltage task
has been registered can be seen in Figure~\ref{sfig:mapping_fig:volt}.

\subsection{Mapping}
\label{subsec:mapping}
After dependence analysis a task is placed on either a {\em ready queue} to be mapped or a 
The high-level runtime maintains the invariant that a task {\em maps} only after all tasks on which it depends have mapped. 
While some predecessor tasks have not mapped a task remains in the {\em waiting queue}; otherwise it is placed on the {\em ready queue}.

%Tasks are mapped observing the partial ordering created by 
%their data dependences.  By mapping tasks based on data dependences we guarantee the 
%semantics of the programming model, while still allowing tasks to execute out of order.

When the scheduler is invoked, it pulls tasks off the ready queue, maps them,
and then issue them to be executed by the low level runtime.  The process of mapping a task
consists of several different queries made by the runtime to an application level {\em mapper} object.
A mapper controls how tasks and data can be placed in the machine.  Mappers are a dynamic version of 
the static mapping files of \cite{Fatahalian06}.  
%Rather than rely on a static mapping
%we opted for a dynamic interface that could make decisions both based on application specific
%input data as well as the machine specifics both at start-up and throughout the execution
%of an application.  %The interface for a mapper is described in Section \ref{subsec:mapinter}.

%\subsubsection{Mapper Interface}
The mapper interface is a collection of eight function calls that allow the runtime to query
the application for information about how to map tasks and regions.  Mapper calls are made
from the runtime to ask where physical instances of regions should be placed in the memory hierarchy,
which physical instances should be the source for copy operations, on which processors
tasks should be run, and how task stealing should be performed.
%\label{subsec:mapinter}
%The interface for a mapper is designed to give the programmer control over how data and tasks
%are placed in the machine.  Every mapper must support the following function calls:

%\begin{itemize}
%\item {\tt rank\_initial\_region\_locations} - When a new logical region is created, the runtime
%will ask the mapper to give a ranking of physical memories of where to place the inital instance.

%\item {\tt rank\_initial\_partition\_locations} - When a new partition is created, the mapper
%must provide a ranking of memory locations for each of the new sub partitions.

%\item {\tt select\_initial\_processor} - Given a task to be mapped, select a processor to
%send this task to for mapping.

%\item {\tt target\_task\_steal} - select a processor to attempt task stealing.  This is explained
%in more detail in Section \ref{subsec:steal}.

%\item {\tt permit\_task\_steal} - also covered in Section \ref{subsec:steal}

%\item {\tt map\_task\_region} - Given a specific task and the logical region to be mapped, select
%a source physical instance for the data, and provide a ranking of memories in which to create
%or reuse an existing physical instance.

%\item {\tt rank\_copy\_targets} - Given a task and a set of valid instances, generate a set of
%rankings for new memories in which to place new phsyical instances.  This operation is used to
%manage the copy operations performed when closing a subtree as they allow the programmer to
%eagerly migrate regions through the memory hierarchy as copy ups are occuring.

%\item {\tt select\_copy\_source} - Given a set of currently valid source instances and destination
%physical instance, select the instance from which a copy should be made.
%\end{itemize}

These eight mapper calls give the programmer complete control over where all tasks are executed, and
where all data is placed within the memory hierarchy.  While this interface allows the programmer
to tune performance, it cannot affect correctness.  The high-level runtime still manages all the task dependences and the
ordering in which tasks are executed and copies are performed.  Therefore, regardless of choices
made by any mapper, the resulting program will always be executed correctly.
To make it easy for programmers to write programs, we also provide a default
 mapper that uses simple heuristics to place tasks and data.  
%For example, when
%performing task stealing, the default mapper only encourages stealing tasks in groups that use
%the same logical regions, which improves locality by moving tasks that use the same regions
%together.  The default mapper implementation makes it easy for the programmer to get an 
%initial program running.  As the programmer tunes the application, he can write mappers that
%specialize for both application specific and/or machine specific parameters.

The high-level runtime also supports the use of multiple mappers in the same application.  Each task 
call may (but does not need to) have an associated mapper,
%runtime  can choose a mapper to be used for the runtime calls made inside that
%task.  
which allows for the creation of library code that provides its own mappers indpendent
from any application level mappers.

\subsubsection{Task Stealing for Load Balance}
\label{subsec:steal}
To achieve load balancing in an application the runtime uses
task stealing.  Our task stealing algorithm is loosely based on the Cilk algorithm
\cite{CILK95} with guidance from the mapper about how stealing should be
performed.  Tasks are initially placed on the ready queue on the processor in which they
were registered.  However, a task can be moved to a different processor either by a mapper
explicitly choosing to relocate it, or by a steal request from another processor's mappers.

When a processor notices that it has no work on its ready queue, it will inquire into each
mapper asking for a target processor to attempt a steal.  A mapper can either respond with
a target processor, or indicate that no steal should be attempted.  The runtime will then
send steal requests to each processor on behalf of each specific mapper.

When a steal request is received at a processor it will determine the mapper(s) that generated
the list of steal requests.  For each mapper that issued a steal request, the runtime will
give the local instance of that mapper a list of the tasks in the ready queue that that
specific mapper is responsible for mapping.  The mapper must then determine the set which
it is going to permit to be stolen, if any.  Any tasks that the mapper marks as stolen are
sent back to the stealing processor.  By only allowing mappers to make steal requests about
the tasks they own, we give mappers complete control over where tasks are mapped.
In addition, we also give the mappers control over how steal requests are made for
the case where the mapper has better knowledge about where additional work is than the 
runtime.  Note that this mapping interface is sufficiently powerful to allow a user to
disable all stealing if desired, and to map tasks to specific processors with no interference
from the runtime.

\subsection{Execution}
\label{subsec:execution}
Once task mapping is complete the high-level runtime performs the necessary operations
to issue the task to the low level runtime.  It first creates any new physical instances
for the regions the task requires.  It then issues copy operations required
for the task to execute contignent on predecessor tasks producing the source physical instances
having finished.  Then it issues the task itself to the low level processor contingent
on all the copies for the task haveing finished as well as the the set of tasks
on which the issuing task depends having finished.  These dependences are all encoded
using {\em events} (see Section \ref{subsec:events}).  Continuing the analogy of an out-of-order
processor, the execution phase is equivalent to issuing an instruction to an
execution unit, and having the execution unit defer the instruction's execution until all its
operands are ready.  The low-level runtime contains processors that act as execution
units that will run a task once all of the events required for the task to run
have occurred.

%Once a task starts running, the runtime is asked to create the region accessors for
%the task.  The runtime investigates whether or not fast accessors can be created for
%each of the regions that the task requires.  Fast accessors are described in more
%detail in Section \ref{sec:fastaccess}.  If fast accessors can be created for every
%region in a task, then the runtime chooses to execute a task instantiated with fast
%region accessor methods.  However, if any region can not be specialized into a fast
%accessor the runtime defaults to a task instantiation that relies on {\tt get} and
%{\tt put} operations to access regions.

%The runtime also assigns a unique identifier describing the context
%in which the task is executing.  All calls into the runtime require the task
%to pass this context identifier which is what permits the runtime to know the context
%for incoming calls for creating/destroying regions and launching sub-tasks.

\subsection{Task Completion}
\label{subsec:cleanup}
We define a task to have {\tt completed} once all of the task's regions reflect
the updated state of the task and all its sub-tasks.  
%This reflects an underlying assumption 
%of the mapping process that the regions for which a task has priviliges will represent the data produced 
%by that task.  However, this is not inherently true that a physical instance of a task reflects
%this state after the task finishes executing.  
It is possible for a task to terminate but to have subtasks that are still in the execution phase.
%If a task has child tasks these tasks may
%still be executing when the parent task finishes the execution phase.  
To handle this scenario the runtime must compute the necessary copy operations
from child task physical instances back to the parent task's physical instances.  To
compute these copy operations, the runtime must first wait until all the child tasks
have been mapped.  The runtime then issues a close operation to the each of the parent
task's region forests.  These close operations record all of the necessary copies to
restore the parent task's regions.  
%We call these copy operations at the end of  a task {\em restoring copies}.
The runtime issues these copies with the required dependences
and marks the task completed once the copy operations are complete.
This is similar to the write-back phase of an out-of-order processor where store instructions cannot be
considered complete until the store effects the state of memory.

The last part of the completetion phase is garbage collecting physical instances.  Physical
instances are not a visible part of the programming model and therefore cannot be managed
directly by the programmer.  Having the mapper manage them would introduce correctness
issues.  We use a two-level reference counting scheme on both abstract instances and physical
instances to know when it is safe to reclaim physical instances.  Due to space constraints
we omit the details of this algorithm.

%Even after a task has completed it is possible for child tasks of that task not to have been 
%executed or even mapped.  This is similar
%to the write-back phase in an out-of-order processor where store instructions can only
%be considered terminated when they finish effecting the state of memory.

%The runtime needs to know the mappings for
%the child tasks as it must issue restoring copy operations so that the set of regions
%that the task operated on are up to date with any data generated by child tasks.

%\subsubsection{Restoring Copy Operations}
%\label{subsec:restore}
%\subsubsection{Garbage Collecting Physical Instances}
%\label{subsec:garbage}
%In addition to cleaning up task execution, the runtime must also deal with the problem
%of garbage collecting physical instances of regions.  Physical instances must be
%garbage collected since they are not visible to the application and would present
%a correctness problem if deleting instances was an option for the mapper.  
%Garbage collecting physical instances is a challenging problem
%due to both the distributed environment of the machine and the deffered execution
%model.  To solve this problem we employ a two-level reference counting scheme.  We
%reference count both abstract instances and as well as physical instances.

%An abstract instance maintains a count of the number of tasks that have referenced 
%that abstract instance, but have not yet been mapped.  An abstract instance also tracks
%whether or not it is open or closed.  An abstract instance is open if it is still
%a valid abstract instance in a region tree.  An abstract instance is marked closed
%once it is no longer a valid abstract instance and can therefore no longer be used
%by any new tasks that are registered.

%When a task is mapped it notifies an abstract instance that it has been used and can
%decrement its reference count.  The task mapping also notifies the abstract instance
%of the specific physical instance that was used and whether or not this use invalidates
%previous physical instances of the abstract instance.  The abstract instance maintains
%a reference count for each physical instance of it that has been made as well as
%the set that are currently valid physical instances to be used for mapping.  When a 
%task terminates, it notifies the abstract instances that it used to decrement the
%reference counts for the task's specific physical instances.

%There are two conditions under which a physical instance can be garbage collected.  The
%first condition is when a physical instance is not a valid physical instance of
%an abstract instance and its reference count is zero.  This condition occurs once no
%more tasks are using a physical instance and it can longer be mapped by any task.  The
%second condition is when an abstract instance is both closed and has a reference count
%of zero as well as the physical instance's reference count being zero.  This handles the
%case when no more tasks can be set to use an abstract instance, there are no more tasks
%to map to an abstract instance, and there are no current tasks using the physical
%instance.

%While reference counting normally doesn't allow for garbage collection of all data structures
%it will always work in this case because both our tasks and our regions are organized as 
%trees which ensures that it is impossible to create cyclic region references in our system.