### 2012-03-02 doc/runtimes/hlr.tex
Beginning to try to write up a description for the high level runtime

\section{Context Initialization}
When a task is first registered with the runtime we create a new {\em context} for the task.  
A context stores all information about that task.  The context for a task will always 
live on the processor in which the task was created even if copies of the task's context 
are made wherever the task is moved.  We call the context on the home processor of the 
task the {\em original} context.  The context for a task also knows the context of the parent 
task that created it.  This parent-child context relationship is fundamental to the propagation 
of information through the runtime.

After we've set up a context and stored all the information from the task launch, we make 
our first mapper call.  We ask the mapper to tell us whether or not this task should be spawned.  
If a task is spawned we will later ask the mapper if it would like to move the task to a remote 
processor.  Spawned tasks are also eligible to be stolen.  If a task is not spawned then it 
is required to be run on the processor on which it was created (cannot be sent anywhere and 
cannot be stolen).

\section{Dependence Analysis}
After the task's context has been initialized, we perform a dependence analysis.  This first 
dependence analysis does NOT determine any actual dependences between tasks.  This first dependence 
analysis determines {\em mapping dependences} between tasks.  A mapping dependence only implies that two 
tasks must be mapped in a specific order.  This may or may not result in them having an actual 
dependence later.  One task has a mapping dependence on a prior task if the later task uses a logical 
region which may alias with a logical region being used by the prior task and at least one of the two 
tasks is using their logical region in a write mode.  Note at this point we can almost completely 
ignore coherence properties.  If two tasks both want a region in simultaneous mode with at least 
one of them requesting a write, they will have a mapping dependence even if they ultimately don't 
have a dependence because they choose to use the same physical instance.  The one exception to this 
rule is with relaxed coherence.  Relaxed coherence will allow two tasks using aliased logical regions 
with at least one write to be mapped in any order.

To find a task's mapping dependences, we must determine which other tasks are using logical regions 
that alias with logical regions for the task we are currently registering.  This is where the type 
system makes it possible for us to only have to perform this analysis on other tasks that share 
the same parent context.  To find the tasks on which we have a mapping dependence, we will use the 
{\em logical state} of the region trees in the parent's context.  For each of the region trees for which 
the parent task has a privilege (any privilege), the parent task maintains two pieces of logical 
state information for each node in the region tree: which tasks in the parent's context are using 
the node and whether which children of the node are open.   A child node is {\em open} if it has users 
in the logical state of the region sub-tree.  Regions are only permitted to have one partition open 
at a time.  Aliased partitions are only permitted to have one child region open at a time.  Disjoint 
partitions are allowed to have any number of child regions open at a time.  Note also that both 
regions and partitions can have users.  Partitions can only be used by index space tasks.

To find mapping dependences, for each logical region that a task needs we first find the 
corresponding region for which the parent has privileges.  This can be either a region that the 
parent was passed as an argument or a region which the parent created.  We call these regions the 
{\em top-most regions} in the parent's context.  For each region that a task requires, we walk from the 
parent's top-most region down to the logical region required by the child task.  At each node we 
examine the users in the logical state of the node.  If there are any dependences between the 
tasks, we register them as mapping dependences.  As we traverse down the tree we open up nodes 
as necessary to arrive at our destination.  Once we arrive at our destination node we mark that 
we are using the current node.

If it at any point we find a path open that we don't intend to traverse, we must first {\em close}
the open subtree, since it's possible that we might use aliased logical regions from that sub-tree.  
This only occurs for regions and aliased partitions.  To close a logical sub-tree, we record all 
users of the logical subtree and move them all to be users of node in our current location.  
This is a valid operation since all these users were really using sub-pieces of the current logical 
region.  We must keep the users of the closed subtree because later tasks will also need to record 
mapping dependences on them.  We record any dependences on the tasks in the closed logical subtree 
and then continue our traversal.  Once we arrive at our destination, we also must close up all 
logical subtrees below as well to record mapping dependences on any tasks using sub-logical regions 
of our intended region.

One optimization that we perform is to replace all the users of a logical region with a single user 
as long as the single user is said to {\em dominate} all the prior users.  A task dominates all the prior 
users of a logical region if it has a mapping dependence on all prior users and all users in any 
open sub-tree.  We only check for a task being a dominator when it reaches its target node since 
it is unlikely that a task dominates all tasks in any open subtree (assuming that most partitions 
in the tree are disjoint).

After we have have found all the mapping dependences for a task, we are ready to place it on one of 
the runtime's scheduling queues.  If a task has mapping dependences we place it on the runtime's 
waiting queue of tasks that are not ready to be mapped yet.  If the task has no mapping dependences 
we place it on the runtime's ready queue of tasks that are ready to map.  As an optimization, for 
tasks that have been spawned and have no mapping dependences, we eagerly query the mapper 
for the target processor of the task.  This allows us to move tasks as soon as possible to their 
target processor.  If the target processor is not the current processor, the task along with a copy 
of its context are relocated to the target processor and then enqueued on the ready queue.

\section{Scheduling and Mapping}
When the application is started the high level runtime registers an {\em idle task} with the low-level 
runtime as a callback for whenever the low-level runtime has little or no-work to perform.  The 
high-level runtime's idle task will choose to perform one of two operations whenever it is invoked.  
If it has any tasks on its ready queue, it will map some of those tasks and launch them on the current 
processor.  If the runtime has little no ready tasks it will ask each of its mappers whether or 
not they want to attempt a task steal.  Task stealing will be discussed in more detail in a 
later section.

Tasks are currently scheduled in a very simple manner.  The oldest task on the ready queue 
will always be the next one mapped.  Changing the scheduling algorithm is easy, but may require 
coordination between mappers over how to weight tasks which is not something we've thought about 
deeply.  Scheduling at the high-level has less of an impact because the order in which tasks are 
mapped isn't necessarily the order in which they are run.  The high-level runtime simply passes 
tasks down to the low-level runtime with their corresponding event dependences.  The low-level 
runtime is the one that ultimately decides the order in which tasks run based primarily on when 
the event dependences are satisfied.

Once a task has been selected to be mapped, we first check to see if it has been spawned and if 
so whether the mapper has already selected its target processor.  If the mapper hasn't had an 
opportunity to a direct a spawned task yet, we ask it where to send the task and move it to the 
target processor.  I'll describe how remote mapping happens later in this section.

\subsection{Mapping}
To map a task, we first ask the mapper to select a location for a physical instance of each of 
the logical regions that the task needs to execute.  The mapper has the option of specifying 
a list of memories in which to try getting a physical instance or specify no memories.  If the 
mapper provides a list memories, the runtime will attempt to create or find a physical instance 
in one memory and only move onto the next if it fails to create or find a physical instance in 
the current one.  If the runtime can't create a physical instance in any of the memories it 
will fail.  If the mapper doesn't specify any memories, the runtime uses this as a signal that 
the task does not want a physical instance for the task.

To aid in the mapper's decision, the runtime provides a list of memories were physical instances with 
{\em valid} data currently reside in the memory hierarchy. A physical instance is said to be valid if
is a the physical instance or a copy the physical instance of the last task to write to the logical
region which the physical instance is representing.

\subsection{Physical Region Tree Traversal}
To be capable of providing this information and finding actual data dependences,
each context also maintains a {\em physical} state of the its topmost regions and their 
corresponding region trees.  The physical state of a node in the region tree is different for 
partitions and regions.  For regions, the physical state includes the state and set of open partitions, 
the set of valid physical instances of the particular logical region, and whether or not the physical 
instances for the region are dirty.  Regions can have multiple partitions open in a read-only state 
and only a single partition open in a write state.  Regions physical instances are marked dirty as 
soon as they have a task that performs a write.  Having dirty physical instance indicates that a 
copy will need to be made back to a parent physical instance at a later point in time.  The physical 
state of a partition node is somewhat simpler.  Partitions only keep track of their open child 
regions.  Aliased partitions can have multiple children in read-only mode or a single child open in 
write mode.  Disjoint partitions can have an arbitrary number of regions open in either read-only 
or write mode.

Similar to the algorithm for dependence analysis on the logical state, for each region that the 
mapper has requested a physical instance we walk the physical region tree from the parent context's 
top-most region down to our target region.  The steps that we perform along the way however are 
different.  Unlike before we don't need to check for any dependences explicitly on regions.  Instead 
we will discover these dependences implicitly as we build up a chain of events that will be required 
to occur before our task can be executed.  As we traverse the tree, we keep a monad corresponding to 
the last event in the chain of dependences that must occur before our task can be launched.  We call
this monad the precondition since it will ultimately define a precondition for launching the task.

As we traverse the tree, we must perform open and close operations analogous to the operations 
performed on the logical region tree.  Open operations just denote opening the physical state 
of the node.  Close operations are slightly more complicated.  To perform a close operation we must 
copy all dirty data back from the physical instances for each of the open subtrees into a physical 
instance of the current region node.  To do this we invoke the help of the mapper.  We first ask the mapper 
to provide a list of memories in which to try to create a physical instance of the current region.  We 
then go through this list of memories until we find or make a physical instance of the current region which
becomes the {\em target} physical instance.  It is an error to fail to make a target instance.

Once we have made a target instance we traverse down the subtree to be closed with the precondition monad
initialized to a null event.  At each node we issue a copy from one of the valid physical instances back
to the target instance conditioned on the precondition monad as well as the termination events of all prior
users of the physical instance.  To select the source physical instance for the copy operation we invoke
the mapper.  The event corresponding to the finish of the copy operation becomes the new state of the
precondition monad and we continue our traversal down the tree.  Once we have traversed the entire tree,
we merge the preconditions for all different subtrees back into our precondition monad and make the corresponding
event the new precondition as we continue traversing the tree.

When we arrive at our target node, we first check to see if the physical instance is a newly created.  If
it has been newly created we must copy valid data from one of the currently valid physical instances.  If the
state of our region is dirty, then we can only use valid instances from our node.  If the state of our instance
is clean, then we can traverse back up the region tree to find physical instances of parent regions that can
also be used as valid sources for our physical instance.  We must also check to see if there are any open sub-trees.  
If there are, then we must issue close operations to them as well.  If there is dirty data in the subtrees this
will result in copies back to the instance we are creating as well which will update the precondition monad.

The last step is to find the set of other tasks which are using the same physical instance as us on which we have
a dependence.  This is necessary because we didn't record any actual dependences in the original mapping dependence
analysis.  The mapping dependence analysis is an over-approximation of the actual dependences we might have since its
possible for tasks on which we have a mapping analysis to complete before we are even mapped.

To track the set of users of a physical instance, the runtime wraps a physical instance in an object called an
{\em InstanceInfo} (need a better name).  InstanceInfos track all the users of a physical instance including both
task users and copys users.  In the case of tasks, InstanceInfos track both the mode and coherence in which the
task is using a physical instance.  In the case of copies, InstanceInfos track whether the copy is a copy that
is reading or writing from the instance.  For all of these users, the InstanceInfo tracks the event corresponding
to when the user will be finished using the InstanceInfo.  Therefore whenever a new user is registered with an
InstanceInfo, the InstanceInfo returns an event that will be bound to the precondition monad indicating the event
that will trigger when all other users on which the new user has a dependence will be finished.

One complication to this is that a physical instance of a logical region can also be used as a physical instance
of any logical subregion.  To handle this case, we can have multiple InstanceInfos for the same physical
instance in the physical state of different regions in the region tree.  Each of these InstanceInfos keeps track
of the users of a physical instance for a specific logical region.  When checking for a dependences on a previous
users of a physical instance it is necessary to find dependences on users using the instance for a logical
region farther up in the region tree.  To handle this case, InstanceInfos also know about all the InstanceInfos
for the same physical instance in logical regions about them in the region tree.  When dependences are checked
on a given InstanceInfo they are also checked in all InstanceInfos farther up the tree.  This is how we are
able to avoid checking dependences when traversing down the physical state of the region tree.

Once the physical region tree state traversal is complete, the resulting event in the precondition monad
becomes a precondition for the task to be executed.  This is repeated for all the regions that the mapper
requested that the task map and the resulting events are merged to create one event that is passed as the
precondition event to the low-level runtime as the task is launched on the low-level processor.

Once the task has been mapped and launched, it notifies all other tasks which had mapping dependences on it.
If these tasks have had all their mapping dependences satisfied then they are moved to the ready queue to
be mapped.  It is important to note that a task can only perform this step if the mapper requested a physical
instance for all of the logical regions of the task.  If the mapper didn't ask for a physical instance for any
of the logical regions, then the task will not be considered mapped until all of its child tasks have been mapped.

\subsection{Remote Mapping}
Tasks can also be mapped remotely from their original context.  In order for this to be possible the region trees
as well as the physical state in the parent task's context must be moved to the region trees.  Before they can be
moved however we first must perform a {\em sanitizing} operation.  If a task is moved remotely it can be mapped
in parallel with other tasks in the same context that don't have any mapping dependence.  However, it is possible
that two or more tasks may rely on the same close operation in traversing to their needed region.  In this case
we don't want both tasks issuing the same close operation on different processors as this can lead to an
inconsistent state of the physical region tree.  Therefore before a task can be sent remotely it traverses the
physical region trees to each of its needed regions and performs any necessary close operations.  In the process
if it opens any regions it will also create valid physical instances at its needed regions using preexisting physical
instances from parent regions.  This way when the task is sent remotely, we only need to send the region trees
and physical state corresponding to the child's top-most regions instead of the parent's top-most regions.

Mapping a task remotely follows the same procedure already described, but the physical state of the region tree
only needs to be traversed from the child's top-most regions since we've already sanitized the region tree
prior to moving it.  After the mapping is complete, the child packages back up the state of all region trees
in which a physical instance was requested by the mapper.  The state is then sent back to the child's original
context which propagates the information back to the parent context on the original processor.  Note that performing
mappings remotely is a safe operation because the dependence analysis on logical regions ensures that no other
tasks in the same context could possibly map on the original processor at the same time if they touched any
of the same logical regions in the set of the child's top-most regions.

\section{Execution}

\section{Children Mapped}

\section{Restoring Copies}

\section{Termination}

\section{Garbage Collection}

\section{Region Tree Updates}

\section{Task Stealing}

\section{Utility Processors and Thread Safety}

\section{TODO}