### 2012-04-29 doc/sc2012/legion.tex

\documentclass[conference]{IEEEtran}

\bibliographystyle{IEEEtran}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{graphics}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\pdfpagewidth=8.5in
\pdfpageheight=11in

\title{Legion: Expressing Locality and Independence with Logical Regions}
\author{
\IEEEauthorblockN{Michael Bauer\IEEEauthorrefmark{1},
%\authorinfo(Stanford University, email: \mbox{mebauer@cs.stanford.edu}
Sean Treichler\IEEEauthorrefmark{1}, 
%\authorinfo(Stanford University, email: \mbox{sjt@cs.stanford.edu}
Elliott Slaughter\IEEEauthorrefmark{1}, 
% \authorinfo(Stanford University, email: \mbox{slaughter@cs.stanford.edu}
Alex Aiken\IEEEauthorrefmark{1}  
%\authorinfo(Stanford University, email: \mbox{aiken@cs.stanford.edu}
%\email{aiken@cs.stanford.edu}
}
\IEEEauthorblockA{
\IEEEauthorrefmark{1}{Computer Science Department, Stanford University}
}
}

\maketitle

\begin{abstract}
Modern parallel architectures have both heterogeneous processors and
deep, complex memory hierarchies.  We present Legion, a programming
model and runtime system for achieving high performance on these
machines.  Legion is organized around {\em logical regions}, which
express both locality and independence of program data, and {\em
  tasks}, functions that perform computations on regions.  We describe
a runtime system that dynamically extracts parallelism from
Legion programs, using a distributed, parallel scheduling algorithm
that identifies both independent tasks and nested parallelism.  Legion
also enables explicit, programmer controlled movement of data through
the memory hierarchy and placement of tasks based on locality
information via a novel mapping interface.  We evaluate our Legion
implementation using four applications: fluid-flow on a regular grid,
a three-level AMR code solving a standard heat diffusion equation, a circuit simulation,
and LINPACK.
%Running on a 4 node cluster with 8 total GPUs 
%and 4 levels of memory hierarchy, our implementation of Legion 
%achieves up to 5.9X speedup over a single CPU-GPU node on real-world applications.
\end{abstract}

\section{Introduction}
\label{sect:intro}
%Put intro text here \cite{Fatahalian06}

Modern parallel machines are increasingly complex, with deep,
distributed memory hierarchies and heterogeneous processing units.  Because the
costs of communication within these architectures vary by several orders of magnitude, 
the penalty for mistakes in the placement of data or computation is usually very poor performance.  Thus,
to achieve good performance it is crucial that the programmer
and the compiler be able to reason about {\em locality} (data resident
close to computation that uses it) and {\em independence} (computations
operating on disjoint data, and therefore not requiring communication and able to be placed in possibly distant parts of the machine).  
Most contemporary
programming systems have no facilities for the programmer to express locality
and independence.  The few languages that do focus primarily on
array-based locality \cite{Fatahalian06,CHAPEL04,UPC99} and 
avoid irregular pointer data structures.
%encode such structures in arrays, which 
%data structures encoded in arrays which are impossible for compilers
%and runtime systems to reason about with regard to locality and independence.

In this paper we describe Legion, a parallel programming system based
on using {\em logical regions} to describe the organization of data.
A logical region names a set of objects.  Logical regions are first
class values in Legion, and in particular may be passed as arguments
to distinguished functions called {\em tasks} that access the data in
those regions, providing locality information.  Logical regions may be
{\em partitioned} into (usually) disjoint {\em subregions}, providing
information for determining independence of computations.
Furthermore, computations access logical regions with particular {\em
  privileges} (e.g., {\em read-only} and {\em read-write}, among
others) and {\em coherence} (e.g., {\em exclusive access} and {\em
  atomic access}, among others).  Privileges express how a task may use
its region arguments, providing data dependence information that
is used to guide the extraction of parallelism.  For example, if two tasks
access the same region with read-only privileges the two tasks
can potentially be run in parallel.
Coherence properties express the required semantics of concurrent
accesses to the region.  For example, if the program executes $f_1(r); f_2(r)$ and tasks $f_1$
and $f_2$ both require {\em exclusive} access to region $r$, then Legion guarantees that 
$f_1(r)$ completes before $f_2(r)$ begins.   On the other hand, if
the tasks access $r$ with {\em atomic} coherence, then Legion
guarantees only atomicity of the tasks with respect to $r$: either task $f_1(r)$
appears to run entirely before $f_2(r)$ or vice versa.

Legion programs are written using logical regions, each of which
has one or more {\em physical instances} when the program executes.  Often
logical regions correspond to a single physical instance, but it is also common and useful
to have multiple physical copies of a logical region's contents in different parts of the machine 
to duplicate and distribute read-only data, to keep an old copy of a region to break a 
write-after-read dependency, or to allow independent, local reductions that are later combined.

To introduce the programming model, we begin with an extended example
of a circuit simulation in
Section~\ref{sec:ex}. This example illustrates regions, tasks,
permissions and coherence properties, the interactions between them,
and how these building blocks are assembled into a Legion
program.  Subsequent sections each describe a contribution in the
implementation and evaluation of Legion:

\input{code_ex}

\begin{itemize}

\item We define a {\em software out-of-order processor}, or SOOP, for
  scheduling tasks with region arguments in a manner analogous to how
  out-of-order hardware schedulers process instructions with register
  arguments (Section~\ref{sec:soop}).  In addition to pipelining the execution of tasks over
  several stages, our SOOP is concurrent and distributed across the machine and is
  also hierarchical and naturally extracts nested
  parallelism (because tasks may recursively spawn subtasks).

\item We use a {\em deferred execution} model that extracts parallelism
even from highly irregular, unbalanced computations.  Instead of blocking,
an operation $o$ that cannot immediately execute because it is dependent
on some other operation $o'$ waits on an event that occurs when $o'$ completes.
Events can be created for the completion of any operation in Legion, and 
the scheduler can build long chains of deferred operations.  In Section~\ref{sec:exp}
we show that dataflow-like deferred execution gives us a significant performance improvement
over bulk synchronous global synchronization primitives such as barriers.

\item A central part of orchestrating a program's execution is how
tasks are assigned (or {\em mapped}) to processors and how physical instances of logical regions are
mapped to specific memory units within the machine  (Section~\ref{sec:mapping}).  Often there is
application-specific information that, if exploited, results in better
mappings than a generic, one-size-fits-all mapping strategy.  We describe
a {\em mapping interface} that allows programmers to give the SOOP a specification
of how to map tasks and regions for a specific application, or even part of an
application.  This mapping API is carefully defined so that
any user-supplied mapping strategy can only affect the performance of applications,
not their correctness.  
%Thus, user mappings only affect the time and resources required
%to run the program and cannot change the program's output.

\item We present the results of experiments on four applications:
  fluid-flow on a regular grid, a three-level AMR code solving a
  standard heat diffusion equation, a circuit simulation, and LINPACK (Section~\ref{sec:exp}).
  We run each of these four codes and best available reference
  versions on three different clusters of multicore processors with
  attached GPUs: a small cluster of our own (4 nodes with 2 6-core CPUs and 2 GPUs per node), a medium-sized cluster
  designed for visualization applications (10 nodes with 2 6-core CPUs and 5 GPUs per node), and up to a third of the
  Keeneland supercomputer (32 nodes, 2 6-core CPUs and 3 GPUs per node)

\end{itemize}

\input{circuit}
\input{part_fig}
\input{soop}
\input{mapping}
\input{experiments}
\input{related}
\input{conclusion}

%\input{mapping_fig}

\bibstyle{IEEEtran}
\bibliography{bibliography}

\end{document}