### 2012-07-03 doc/popl2013/legion-types.tex

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}

\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{multirow}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{A Type System for Safe and Scalable Parallelism using Logical Regions}
%\title{Scalable and Safe Parallelism using Logical Regions}
\authorinfo{}{}{}
\maketitle

\begin{abstract}
Legion is a programming model that introduces 
{\em logical regions} as a mechanism for targeting large heterogeneous machines 
with complex memory hierarchies.  Logical regions are a first-class 
feature of Legion that allow programmers to specify independence 
and locality properties of program data dynamically.  We present
a type and effect system for reasoning about logical regions in Legion.  
We prove a soundness result for our type system and use it to show
that all memory accesses in Legion are safe.  Futhermore, we prove that
the soundess of our type system enables a scalable hierarchical
scheduling algorithm for Legion programs.  We implement a type checker
for Legion programs and improve performance up to 85\% by eliding
unncessary memory checks.  We demonstrate an implementation of the Legion 
runtime based on the hierarchical scheduling algorithm scales for
three real-world applications.

%Limit is 12 pages, including bibliography and appendices.
%Deadlines: \\

%\begin{tabular}{ll}
%Paper registration & July 7, 2012 3:59am PDT \\
%Paper submission & July 11, 2012 3:59am PDT \\
%Author response & September 10 - September 13, 2012 \\
%Author notification & October 1, 2012
%\end{tabular}
\end{abstract}

\section{Introduction}
\label{sect:intro}
Introduction goes here.

%\input{outline}

\input{langdef}

\input{conclusion}

{
\small
\bibliography{bibliography}
}

\end{document}