### 2012-07-05 doc/popl2013/legion-types.tex
intro

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}

\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{multirow}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{A Type System for Safe and Scalable Parallelism using Logical Regions}
%\title{Scalable and Safe Parallelism using Logical Regions}
\authorinfo{Sean Treichler \and Michael Bauer \and Alex Aiken}{Stanford University}{\{sjt,mebauer,aiken\}@cs.stanford.edu}
%\authorinfo{Sean Treichler}{Stanford Unviersity}{sjt@cs.stanford.edu}
%\authorinfo{Michael Bauer}{Stanford University}{mebauer@cs.stanford.edu}
%\authorinfo{Alex Aiken}{Stanford University}{aiken@cs.stanford.edu}
\maketitle

\begin{abstract}
Legion is a programming model that introduces logical regions as a mechanism for targeting large heterogeneous machines with complex memory hierarchies. Logical regions are first-class values in Legion, allowing programmers to specify independence and locality properties of program data dynamically. We present a type system with permission and coherence properties for reasoning about logical regions.  We prove the soundness of the  type system and use it to show that all memory accesses in Legion are safe. Furthermore, we show that type soundness directly enables a scalable, hierarchical scheduling algorithm for Legion programs. We implement a type checker for Legion programs and improve performance up to 71% by eliding memory checks the type system proves safe. We also demonstrate that an implementation of the Legion runtime based on the hierarchical scheduling algorithm scales well for three real-world applications.

%Limit is 12 pages, including bibliography and appendices.
%Deadlines: \\

%\begin{tabular}{ll}
%Paper registration & July 7, 2012 3:59am PDT \\
%Paper submission & July 11, 2012 3:59am PDT \\
%Author response & September 10 - September 13, 2012 \\
%Author notification & October 1, 2012
%\end{tabular}
\end{abstract}

\input{intro}

%\input{outline}

\input{langdef}

\input{evaluation}

\input{related}

\input{conclusion}

{
\small
\bibliography{bibliography}
}

\end{document}