### 2012-07-06 doc/popl2013/legion-types.tex
intro

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}

\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{multirow}
\usepackage{ifmtarg}
\usepackage{url}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{A Type System for Safe and Scalable Parallelism using Logical Regions}
%\title{Scalable and Safe Parallelism using Logical Regions}
\authorinfo{Sean Treichler \and Michael Bauer \and Alex Aiken}{Stanford University}{\{sjt,mebauer,aiken\}@cs.stanford.edu}
%\authorinfo{Sean Treichler}{Stanford Unviersity}{sjt@cs.stanford.edu}
%\authorinfo{Michael Bauer}{Stanford University}{mebauer@cs.stanford.edu}
%\authorinfo{Alex Aiken}{Stanford University}{aiken@cs.stanford.edu}
\maketitle

\begin{abstract}

Legion is a programming model that introduces 
{\em logical regions} as a mechanism for targeting large heterogeneous machines 
with complex memory hierarchies.  Logical regions are first-class 
values in Legion, allowing programmers to specify independence 
and locality properties of program data.  We present
a type system with permissions and coherence properties for reasoning about logical regions.  
We prove the soundness of the type system and use it to show
that all memory accesses in Legion are safe.  Furthermore, we prove that
type soundness enables a scalable, hierarchical
scheduling algorithm for Legion programs.  We show that an implementation of the type checker
for Legion programs improves performance up to 71\% by eliding
memory checks the type system proves safe.  We also demonstrate that an implementation of the Legion 
runtime based on the hierarchical scheduling algorithm scales well for
three real-world applications.

%Limit is 12 pages, including bibliography and appendices.
%Deadlines: \\

%\begin{tabular}{ll}
%Paper registration & July 7, 2012 3:59am PDT \\
%Paper submission & July 11, 2012 3:59am PDT \\
%Author response & September 10 - September 13, 2012 \\
%Author notification & October 1, 2012
%\end{tabular}
\end{abstract}

\input{intro}

%\input{outline}

\input{langdef}

\input{evaluation}

\input{related}

\input{conclusion}

{
\small
\bibliography{bibliography}
}

\end{document}