### 2012-07-08 doc/popl2013/langdef.tex
Getting rid of the ugly boxes for the type system

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}

\newcommand{\oton}[1]{{#1}_1,\ldots,{#1}_n}
\newcommand{\otok}[2]{{#2}_1,\ldots,{#2}_{#1}}
\newcommand{\dplus}{\text{+\!+}}
\newcommand{\llbracket}{[\![}
\newcommand{\rrbracket}{]\!]}
\makeatletter
\newcommand{\nonint}[1][]{\#\@ifmtarg{#1}{}{_{\!\!#1}}}
\makeatother

% fun latex tricks to make typeenv and opsenv more friendly
\makeatletter
\define@key{typeenv}{G}{\def\typeenv@G{#1}}
\define@key{typeenv}{P}{\def\typeenv@P{#1}}
\define@key{typeenv}{O}{\def\typeenv@O{#1}}
\newcommand{\typeenvx}[1][]{
{
% default values
\def\typeenv@G{\Gamma}
\def\typeenv@P{\Phi}
\def\typeenv@O{\Omega}
\setkeys{typeenv}{#1}
\typeenv@G, \typeenv@P, \typeenv@O \vdash \,
}}
\newcommand{\typeenv}[3][]{\typeenvx[#1] {#2} : {#3}}
\define@key{opsenv}{M}{\def\opsenv@M{#1}}
\define@key{opsenv}{L}{\def\opsenv@L{#1}}
\define@key{opsenv}{H}{\def\opsenv@H{#1}}
\define@key{opsenv}{S}{\def\opsenv@S{#1}}
\define@key{opsenv}{C}{\def\opsenv@C{#1}}
\newcommand{\opsenvx}[1][]{
{
% default values
\def\opsenv@M{M}
\def\opsenv@L{L}
\def\opsenv@H{H}
\def\opsenv@S{S}
\def\opsenv@C{C}
\setkeys{opsenv}{#1}
\opsenv@M, \opsenv@L, \opsenv@H, \opsenv@S, \opsenv@C \vdash \,
}}
\newcommand{\opsenv}[4][]{\opsenvx[#1] {#2} \mapsto {#3}, {#4}}
\makeatother

\begin{figure*}
\centering
{\small
\begin{tabular}{cclr|cclr}

$T$ & ::= &  & types & $bv$ & ::= & false $\;\;\;\mid\;\;\;$ true & \\
  &$\mid$& bool $\;\;\;\mid\;\;\;$ int & base types & & & & \\
  &$\mid$& $\langle T_1, T_2 \rangle$ & tuple & $iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ & \\
  &$\mid$& $T@(\oton{r})$ & pointer & & & & \\
  &$\mid$& $\text{coloring}(r)$ & region coloring & $e$ & ::= & & expressions \\
  &$\mid$& $\exists \oton{r}. T\text{ where }\Omega$ & region relationship &   &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
  &$\mid$& $\forall \oton{r}. (\oton{T}), \Phi, Q \rightarrow T_r$ & functions &   &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
& & & &   &$\mid$& $id$ &  \\
$\Omega$ & ::= & $\{ \oton{\omega} \}$ & region constraints &   &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
$\omega$ & ::= & $r_1 \leq r_2$ & subregion   &   
  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ & \\
  &$\mid$& $r_1 * r_2$ & disjointness &   
  &$\mid$ & $\text{downregion}(e, r_1,\ldots,r_n)$ & \\  
& & &  & 
  &$\mid$& $\text{read}(e_1)$ $\;\mid\;$ $\text{write}(e_1, e_2)$ & memory access \\
$\Phi$ & ::= & $\{ \oton{\phi} \}$ & privileges & 
  & $\mid$ & $\text{reduce}(id, e_1, e_2)$ & \\
$\phi$ & ::= & reads$(r)$ $\;\mid\;$ writes$(r)$ $\;\mid\;$ $\text{reduces}_{id}(r)$ &  &  
  &$\mid$& $\text{newcolor}\ r$ $\;\mid\;$ $\text{color}(e_1, e_2, e_3)$ & coloring \\
& & & &  
  &   $\mid$& $e_1 + e_2$ & integer ops \\
$Q$ & ::= & $\{ \oton{q} \}$ & coherence modes  &   
       & $\mid$& $e_1 < e_2$ & comparisons \\
$q$ & ::= & atomic$(r)$ $\;\;\;\mid\;\;\;$ simult$(r)$ & &   
  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
& & & &   
       & $\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
$v$ & ::= & & values &   
       & $\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & task calls \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & base values &   
       & $\mid$ & $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
  &$\mid$& $\langle v_1, v_2 \rangle$ & tuple & 
       & $\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
  &$\mid$& null $\;\;\;\mid\;\;\;$ $a$ & address & 
       & $\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\
  &$\mid$& $\{ (a_i, iv), \ldots \}$ & coloring & 
    & & & \\
  &$\mid$& $\langle \langle \oton{\rho}, v\rangle \rangle$ & reg. relation instance & 
   & & & \\
%$bv$ & ::= & false $\;\;\;\mid\;\;\;$ true \\
%\\
%$iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ \\
%\\
%$e$ & ::= & & expressions \\
%  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
%  &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
%  &$\mid$& $id$ &  \\
%  &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
%  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ $\;\;\;\mid\;\;\;$ $\text{downregion}(e, r_1,\ldots,r_n)$ & \\
%  &$\mid$& $\text{read}(e_1)$ $\;\;\;\mid\;\;\;$ $\text{write}(e_1, e_2)$ $\;\;\;\mid\;\;\;$ $\text{reduce}(id, e_1, e_2)$ & memory access \\
%  &$\mid$& $\text{newcolor}\ r$ $\;\;\;\mid\;\;\;$ $\text{color}(e_1, e_2, e_3)$ & coloring operations \\
%  &$\mid$& $e_1 + e_2$ & integer operations \\
%  &$\mid$& $e_1 < e_2$ & comparison operations \\
%  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
%  &$\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
%  &$\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & task calls \\
%  &$\mid$& $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
%  &$\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
%  &$\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\
\end{tabular}
}
\caption{Legion Core Language}
\label{fig:langdef}
\end{figure*}

%\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
%\newcommand{\cinfrule}[3]{\parbox{14cm}{\hfil$\infrule{#1}{#2}$\hfil}\parbox{4cm}{$\,#3$\hfil}}
%\newcommand{\finfrule}[2]{\framebox{$\infrule{#1}{#2}$}}
%\newcommand{\oldfinfrule}[2]{\vspace{10pt}\framebox{$\infrule{#1}{#2}$}\vspace{10pt}}

%\newcommand{\infx}[2]{\infrule{\begin{array}{l}{#1}\end{array}}{#2}}

\section{Legion}
\label{sec:legioncore}

Figure~\ref{fig:langdef} defines the core subset of Legion.  Types
include booleans, integers, tuples, and pointers.  Note that pointers
are annotated with a list of regions into which they may point; in a
well-typed program, the pointer is guaranteed to point to at least one
of those regions. There is a special type for {\em colorings}, which
are used to specify partitions of regions into subregions.
Functions/tasks in Legion are named, accept one or more arguments and
return a value of some type.  A function also declares the region
access privileges and coherence it requires.  Function types are
universally quantified over all region names appearing in the type.
The final value is a region relation, which stores a value and one or
more logical regions satisfying associated region constraints.  An
instance of a region relation can be read out of the heap and
unpacked, giving new local names to the logical regions contained in
the instance.
% and tracking
%the constraints between those new logical regions.

\subsection{Core Type System}
\label{subsec:coretypes}

Core Legion is explicitly typed using judgments of the form
\begin{center}
$\typeenv{e}{T}$
\end{center}
where
\begin{itemize}
\item $\Gamma$ is the type environment, mapping the names of local variables and global functions
to their types,
\item $\Phi$ is the region access privileges under which the expression is evaluated, and
\item $\Omega$ is the set of constraints that must hold between regions
\end{itemize}
As regions are hierarchical, privileges and/or constraints on a region often imply similar
privileges or constraints on subregions.  Figure~\ref{fig:closure} defines closure
operations for $\Phi^*$ and $\Omega^*$ that add implied constraints.
The type checking rules are in Figure~\ref{fig:types}.  Many of the rules are 
straightforward, but several deserve more explanation.

Core Legion has no subtyping. A pointer's region is changed only via explicit
{\em upregion} and {\em downregion} operations.  {\em Upregion} calls can be inferred,
but {\em downregion} must be accompanied by a run-time check that the pointer really does
point into the subregion(s), so {\em upregion} is made explicit as well for simplicity.

Accesses to heap memory are performed via the {\em read}, {\em write}, and {\em reduce}
operations, which are only legal if the regions in a pointer's type are ones for which the
enclosing function has privileges.
Similarly, a call to another function is permitted only if the privileges required by the 
callee are a subset of those held by the caller.  There is no way to create privileges within
a Legion function - it is limited to privileges given to it by its caller.

The {\em partition} operation uses a {\em coloring} to split a region into 
disjoint subregions.  The subregions are given names within the body of the partition,
as well as constraints asserting subregions are pairwise disjoint and contained within the parent region.
%These constraints can be used locally, or packed into region relation instances that can be unpacked and used later.

The {\em pack} expression packs together a value and the corresponding logical regions,
provided the region relation's constraints hold.  The
corresponding {\em unpack} expression gives (new) names to those regions.  For simplicitly, neither the 
{\em partition} nor the {\em unpack} expression allows shadowing of region names.

\begin{comment}
\begin{figure*}
\centering{
\framebox{$\typeenv{bv}{bool}$}
\framebox{$\typeenv{iv}{int}$}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\typeenvx e_2 : T_2
\end{array}}
{\typeenvx \langle e_1, e_2 \rangle : \langle T_1, T_2 \rangle}
\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.1}\ : T_1}
\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.2}\ : T_2}
\finfrule{\Gamma(id) = T}{\typeenvx id : T}
\framebox{$\typeenvx \text{null }T@r : T@r$}
\framebox{$\typeenvx \text{new }T@r : T@r$}
\finfrule{\typeenvx e : T@(\oton{r})}{\typeenvx \text{isnull}(e) : bool}
\finfrule
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\forall i, reads(r_i) \in \Phi^*\end{array}}
{\typeenvx \text{read}(e_1) : T}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\typeenvx e_2 : T \\
\forall i, writes(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \\
\typeenvx e_1 : T_1@(\oton{r}) \\
\typeenvx e_2 : T_2 \\
\forall i, reduces_{id}(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}
\framebox{$\typeenvx \text{newcolor }r : \text{coloring}(r)$}
\finfrule{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r) \\
\typeenvx e_2 : T@r \\
\typeenvx e_3 : int
\end{array}}
{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}
\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 + e_2 : int}
\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 < e_2 : bool}
\finfrule{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\typeenvx[G={\Gamma[id/T_1]}] e_2 : T_2
\end{array}}
{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2}
\finfrule{\begin{array}{l}\typeenvx e_1 : bool \\ \typeenvx e_2 : T \\ \typeenvx e_3 : T\end{array}}{\typeenvx \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T}
\finfrule{
\begin{array}{l}
\Gamma(id) = \forall r'_1, \ldots r'_k.(\oton{T}),\Phi', Q' \rightarrow T_r \\
\typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi'[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]}
\finfrule{
\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\
\typeenvx[O=\Omega'] e_2 : T \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset
\end{array}}
{\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}
\finfrule{
\begin{array}{l}
T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\
\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\
\typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1}
\finfrule{
\begin{array}{l}
T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \\
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset
\end{array}}
{\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3}
\finfrule{
\begin{array}{l}
\text{for }1 \leq i \leq p, \\
\Gamma(id_i) = \forall r_{i1}, \ldots r_{ik_i}. (T_{i1}, \ldots, T_{in_i}), \Phi_i, Q_i \rightarrow T_{ir} \\
\Gamma_i = \Gamma[a_{i1}/T_{i1}, \ldots, a_{in_i}/T_{in_i}] \\
\typeenv[G={\Gamma_i},P={\Phi_i},O=\emptyset]{e_i}{T_{ir}}
\end{array}}
{ 
\begin{array}{l@{ }l}
\vdash \{ & \text{task }id_1[r_{11}, \ldots, r_{1k_1}]( a_{11} : T_{11}, \ldots a_{1n_1} : T_{1n_1} ), \Phi_1, Q_1 : T_{1r} : e_1, \\
          & \ldots \\
          & \text{task }id_p[r_{p1}, \ldots, r_{pk_1}]( a_{p1} : T_{p1}, \ldots a_{pn_p} : T_{pn_p} ), \Phi_p, Q_p : T_{pr} : e_p \} : \bullet
\end{array} }
}
\caption{Legion Core Type System}
\label{fig:types}
\end{figure*}
\end{comment}

\begin{figure*}
\begin{adjustwidth}{-4em}{-4em}
\begin{center}
% Row 1: bool, int, tuple1, tuple2
\begin{minipage}{\linewidth}
%\centering
\begin{tabular}{m{4cm}m{3.5cm}m{5cm}m{5cm}}
\infax[BOOL]{\typeenv{bv}{bool}} & 
\infax[INT]{\typeenv{iv}{int}} & 
\infrule[TUPLE1]{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.1}\ : T_1} & 
\infrule[TUPLE2]{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.2}\ : T_2} 
\end{tabular}
\end{minipage}
% Row 2:  var, null, isnull, maketuple
\begin{minipage}{\linewidth}
\vspace{-0.4cm}
%\centering
\begin{tabular}{m{3cm}m{3cm}m{5cm}m{6cm}}
\infrule[VAR]{\Gamma(id) = T}{\typeenvx id : T} &
\infax[NULL]{\typeenvx \text{null }T@r : T@r} &
\infrule[ISNULL]{\typeenvx e : T@(\oton{r})}{\typeenvx \text{isnull}(e) : bool} &
\infrule[MAKETUPLE]{\typeenvx e_1 : T_1 \andalso \typeenvx e_2 : T_2}{\typeenvx \langle e_1, e_2 \rangle : \langle T_1, T_2 \rangle}
\end{tabular}
\end{minipage}
% Row 3: new, newcolor, coloring
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
%\centering
\begin{tabular}{m{4cm}m{5cm}m{9cm}}
\infax[NEW]{\typeenvx \text{new }T@r : T@r} &
\infax[NEWCOLOR]{\typeenvx \text{newcolor }r : \text{coloring}(r)} &
\infrule[COLORING]{\typeenvx e_1 : \text{coloring}(r) \andalso \typeenvx e_2 : T@r \andalso \typeenvx e_3 : int}{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}
\end{tabular}
\end{minipage}
% Row 4: let, add, compare
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
%\centering
\begin{tabular}{m{6cm}m{5cm}m{6cm}}
\infrule[LET]{\typeenvx e_1 : T_1 \\ \typeenvx[G={\Gamma[id/T_1]}] e_2 : T_2}{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} &
\infrule[ADD]{\typeenvx e_1 : int \\ \typeenvx e_2 : int}{\typeenvx e_1 + e_2 : int} &
\infrule[COMPARE]{\typeenvx e_1 : int \\ \typeenvx e_2 : int}{\typeenvx e_1 < e_2 : bool}
\end{tabular}
\end{minipage}
% Row 5: upregion, downregion
\begin{minipage}{\linewidth}
\vspace{-0.4cm}
\begin{tabular}{m{9cm}m{10cm}}
\infrule[UPREG]{\typeenvx e : T@(r'_1, \ldots r'_k) \andalso \forall i. \exists j, r'_i \leq r_j \in \Omega^*}{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})} &
\infrule[DOWNREG]{\typeenvx e : T@(r'_1, \ldots r'_k) \andalso \forall j. \exists i, r_j \leq r'_i \in \Omega^*}{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})}
\end{tabular}
\end{minipage}
% Row 6: read, write, reduce
\begin{minipage}{\linewidth}
\vspace{-0.3cm}
\begin{tabular}{m{5cm}m{5cm}m{8cm}}
\infrule[READ]{\typeenvx e_1 : T@(\oton{r}) \\ \forall i, reads(r_i) \in \Phi^*}{\typeenvx \text{read}(e_1) : T} &
\infrule[WRITE]{\typeenvx e_1 : T@(\oton{r}) \\ \typeenvx e_2 : T \\ \forall i, writes(r_i) \in \Phi^*}{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})} &
\infrule[REDUCE]{\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \andalso \typeenvx e_2 : T_2 \\
 \typeenvx e_1 : T_1@(\oton{r}) \andalso \forall i, reduces_{id}(r_i) \in \Phi^*}{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}
\end{tabular}
\end{minipage}
% Row 7: pack, if-else, task call
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{5cm}m{5cm}m{8cm}}
\infrule[PACK]{T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\ \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\ \typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]}{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} &
\infrule[IF-ELSE]{\typeenvx e_1 : bool \\ \typeenvx e_2 : T \\ \typeenvx e_3 : T}{\typeenvx \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T} &
\infrule[TASK CALL]{\Gamma(id) = \forall r'_1, \ldots r'_k.(\oton{T}),\Phi', Q' \rightarrow T_r \\ \typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\ \Phi'[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*}{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]}
\end{tabular}
\end{minipage}
% Row 8: unpack, partition
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{10cm}m{8cm}}
\infrule[UNPACK]{T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \andalso \typeenvx e_1 : T_1 \\
  \Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \andalso \Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
  \typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \andalso \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset}
  {\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3} &
\infrule[PARTITION]{\typeenvx e_1 : \text{coloring}(r_p) \\ \Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\ \typeenvx[O=\Omega'] e_2 : T \\ \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset}
  {\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}
\end{tabular}
\end{minipage}
% Row 9: program
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\infrule[PROGRAM]{\text{for }1 \leq i \leq p, \\
  \Gamma(id_i) = \forall r_{i1}, \ldots r_{ik_i}. (T_{i1}, \ldots, T_{in_i}), \Phi_i, Q_i \rightarrow T_{ir} \andalso
  \Gamma_i = \Gamma[a_{i1}/T_{i1}, \ldots, a_{in_i}/T_{in_i}] \andalso
  \typeenv[G={\Gamma_i},P={\Phi_i},O=\emptyset]{e_i}{T_{ir}}}
  {\vdash \{  \text{task }id_1[r_{11}, \ldots, r_{1k_1}]( a_{11} : T_{11}, \ldots a_{1n_1} : T_{1n_1} ), \Phi_1, Q_1 : T_{1r} : e_1, 
           \ldots
           \text{task }id_p[r_{p1}, \ldots, r_{pk_1}]( a_{p1} : T_{p1}, \ldots a_{pn_p} : T_{pn_p} ), \Phi_p, Q_p : T_{pr} : e_p \} : \bullet}
\end{minipage}
\end{center}
\end{adjustwidth}
\caption{Legion Core Type System}
\label{fig:types}
\end{figure*}

\begin{figure}
\centering{
$\begin{array}{l}
\Omega \subseteq \Omega^* \\
r_i \leq r_j \in \Omega^*  \Rightarrow r_i \leq r_i \in \Omega^*\wedge r_j \leq r_j \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j \leq r_k \in \Omega^* \Rightarrow r_i \leq r_k \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j * r_k \in \Omega^* \Rightarrow r_i * r_k \in \Omega^* \\
r_i * r_j \in \Omega^* \Rightarrow r_j * r_i \in \Omega^* \\
\\
\Phi \subseteq \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reads(r_j) \in \Phi^* \Rightarrow reads(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge writes(r_j) \in \Phi^* \Rightarrow writes(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reduces_{id}(r_j) \in \Phi^* \Rightarrow reduces_{id}(r_i) \in \Phi^* \\
reads(r) \in \Phi^* \wedge writes(r) \in \Phi^* \Rightarrow reduces_{id}(r) \in \Phi^* \\
\ \ \ \mbox{for every function identifier $id$}
\end{array}$
}
\caption{Privilege and Constraint Closure}
\label{fig:closure}
\end{figure}

\subsection{Operational Semantics}
\label{subsec:opsemantics}

Core Legion's operational semantics rules have the following form:

\begin{center}
$\opsenv{e}{v}{E}$
\end{center}

where: 
\begin{itemize}
\item $M$ is a mapping from logical regions $r_i$ to {\em physical regions} $\rho_i$, which are sets of concrete memory locations.
We write $\hat x = M \llbracket x \rrbracket$ for the application of $M$ to other objects (e.g., types), 
replacing all free logical region names $r_i$ in $x$ with corresponding physical regions $\rho_i$.  
(Note $M \llbracket x \rrbracket$ is ill-formed if it contains a free logical region name $r_j \not\in dom(M)$.)

\item $L$ is a mapping of local variables to (immutable) values.

\item $H$ is a {\em heap typing}, assigning an invariant type to every memory location.

\item $S$ is a store, mapping memory locations $l_i$ to values.

\item $C$ is a {\em clobber set}, memory locations that may be accessed by concurrently executing operations.

\item $e$ is an expression.

\item $v$ is the result of evaluatinig $e$.

\item $E$ is a {\em memory trace} of the execution, a list of {\em memory operations}
in the order performed.  A memory operation $\epsilon$ is either
$read(l, c, v)$ (a read of location $l$ with coherence mode $c$, returning value $v$),
$write(l, c, v)$ (a write to location $l$ with coherence mode $c$, storing value $v$),
or $reduce_{id}(l, c, v)$ (a reduction into location $l$ with coherence mode $c$ using an
associative function $id$ with value $v$).  Coherence modes are one of ({\em excl}, {\em atomic}, {\em simult}).
\end{itemize}

Helper functions are used in some rules:
\begin{itemize}
\item $apply(S, E) = S'$ computes the store $S'$ resulting from applying memory trace $E$ to initial store $S$.

\item $mark\_coherence(E, \hat Q)$ records the coherence of operations in trace $E$ according to coherence annotations $Q$.

\item $valid\_interleave(S, C, E', \oton{E})$ is true if $E'$ is a valid interleaving
of the memory traces $\oton{E}$.
\end{itemize}
Figure~\ref{fig:opsemfns} defines {\em apply} and {\em mark\_coherence}.  The precise definition of
{\em valid\_interleave} is given in Section~\ref{sec:coherence}.  For now, we need only that a valid interleaving
is at least an interleaving, containing each constituent trace as a subequence and no other operations.

Some rules modeling concurrent execution are locally
non-deterministic, guessing values that result from
concurrent tasks.  For example, the rule for {\tt read(e)} returns an
arbitrary value $v'$ of the correct type if the location being read is
in the clobber set $C$, the set of locations concurrent tasks may
clobber (i.e., update).\footnote{Note that this is also the reason for
including a global heap typing.}  The guess is eventually checked for
consistency by the $valid\_interleave$ predicate of some enclosing
expression, at a point where the task that wrote $v'$ is also local to
the expression being evaluated.

Another prescient operation is {\tt newcolor r}, which guesses initial
memory locations and their colors for a partition.  These locations
support partition-then-allocate, ensuring the memory mappping $M$ has
locations for the new subregions that can be allocated by {\tt new}.
As is usual, memory allocation is also non-deterministic, choosing any
location of the correct type in the region.  For simplicity we do not
require locations to be ``fresh''; our soundness results hold even if
{\tt new T@r} returns locations already in use, provided the heap
typing of the location is correct.

The most involved rule in Core Legion's semantics is for task calls.
Note that the only regions accessible to the callee (through $M'$) are the region arguments.
Also, if any regions are used with relaxed coherence (either {\em atomic} or {\em simult} instead of
{\em excl}), those locations are added to the clobber set, allowing the
impact of concurrent modifications of the store to be handled by the caller.  
%This is done by
%annotating the called function's memory trace with the relaxed coherence on a per-location basis
%and using the {\em valid\_interleave} predicate to control what parallel execution is allowed.

\begin{figure*}
\centering{
$
\begin{array}{lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \text{ where } S' = apply(S, E) \\
mark\_coherence([], \hat Q) & = & [] \\
mark\_coherence(E \dplus [ op(l, c, v) ], \hat Q) & = & mark\_coherence(E, \hat Q) \dplus \\
& & op(l, c', v), \text{ where } \\
& & c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge simult(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge atomic(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases} \\
any\_interleave([], [], \ldots, []) & = & true \\
any\_interleave([ \epsilon ] \dplus E', E_1, \ldots, [ \epsilon ] \dplus E_i, \ldots, E_n) & = & any\_interleave(E', E_1, \ldots, E_i, \ldots, E_n) \\
valid\_interleave(S, C, E', \oton{E}) & \Rightarrow & any\_interleave(E', \oton{E}) \\
\end{array}
$}
\caption{Operational Semantics Helper Functions}
\label{fig:opsemfns}
\end{figure*}

\begin{comment}
\begin{figure*}
\centering{\small
\framebox{$\opsenvx bv \mapsto bv, []$}
\framebox{$\opsenvx iv \mapsto iv, []$}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = apply(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E'}
\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.1} \mapsto v_1, E} \hspace{1cm}
\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.2} \mapsto v_2, E}
\finfrule
{L(id) = v}
{\opsenvx id \mapsto v, []}
\framebox{$\opsenvx \text{null }T@r \mapsto null, []$}
\finfrule
{\begin{array}{l}
l \in M(r) \\
H(l) = M \llbracket T \rrbracket
\end{array}}
{\opsenvx \text{new }T@r \mapsto l, []}
\finfrule
{\opsenvx e \mapsto l, E}
{\opsenvx \text{isnull}(e) \mapsto \textit{false}, E}
\finfrule
{\opsenvx e \mapsto null, E}
{\opsenvx \text{isnull}(e) \mapsto true, E}
\finfrule
{\opsenvx e \mapsto v, E}
{\opsenvx \text{upregion}(e, \oton{r}) \mapsto v, E}
\finfrule
{\begin{array}{l}
\opsenvx e \mapsto l, E \\
v = \begin{cases}
l, & \text{if $\exists i, l \in M(r_i)$}. \\
null, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenvx \text{downregion}(e, \oton{r}) \mapsto v, E}
\finfrule
{\begin{array}{l}
\opsenvx e \mapsto l, E \\
S' = \text{apply}(S, E) \\
v = \begin{cases}
S'(l), & \text{if } l \not\in C \\
v' : H(l), & \text{otherwise}
\end{cases}
\end{array}}
{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}
\finfrule{
\begin{array}{l@{}l}
K = \{ (l_1, iv_1), & \ldots, (l_p, v_p) \}, \text{ where } \\
& (\forall i \in [1,p]. l_i \in M(r)) \wedge \\
& (\forall i,j \in [1,p]. l_i \not= l_j)
\end{array}
}
{\opsenvx \text{newcolor }r \mapsto K, []}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
valid\_interleave(S, C, E', E_1, E_2, E_3)
\end{array}}
{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 + e_2 \mapsto v', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
\textit{false}, & \text{otherwise}.
\end{cases} \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 < e_2 \mapsto v', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto \textit{false}, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_3 \mapsto v_3, E_3
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1) \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E})
\vspace{1.5mm} \\
\text{task }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \\
S' = \text{apply}(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S'] e_{n+1} \mapsto v_{n+1}, E_{n+1}
\vspace{1.5mm} \\
E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \\
valid\_interleave(S, C, E'', E', E_{n+1})
\end{array}}
{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}
\finfrule{
\begin{array}{l}
\opsenv{e_1}{K}{E_1} \\
\rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'}
\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto v, E \\
\rho_i = M[r_i], \text{ for } 1 \leq i \leq k \\
v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle
\end{array}}
{\opsenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E}
\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E' }
}
\caption{Legion Core Operational Semantics}
\label{fig:semantics}
\end{figure*}
\end{comment}

\begin{figure*}
\begin{adjustwidth}{-4em}{-4em}
\begin{center}
% Row 1: bool, int, val
\begin{minipage}{\linewidth}
\begin{tabular}{m{6cm}m{6cm}m{6cm}}
\infax[OP-BOOL]{\opsenvx bv \mapsto bv, []} &
\infax[OP-INT]{\opsenvx iv \mapsto iv, []} &
\infax[OP-NULL]{\opsenvx \text{null }T@r \mapsto null, []} 
\end{tabular}
\end{minipage}
% Row 2: tuple1, tuple2, make tuple
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{9cm}m{4.5cm}m{4.5cm}}
\infrule[OP-MAKETUPLE]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S' = apply(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E'} &
\infrule[OP-TUPLE1]{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}{\opsenvx e\text{.1} \mapsto v_1, E} &
\infrule[OP-TUPLE2]{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}{\opsenvx e\text{.2} \mapsto v_2, E}
\end{tabular}
\end{minipage}
% Row 3: new, null, isnull, isnull
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{4cm}m{4.5cm}m{5cm}m{5cm}}
\infrule[OP-NEW]{l \in M(r) \\ H(l) = M \llbracket T \rrbracket}{\opsenvx \text{new }T@r \mapsto l, []} &
\infrule[OP-VAL]{L(id) = v}{\opsenvx id \mapsto v, []} &
\infrule[OP-ISNULL-FALSE]{\opsenvx e \mapsto l, E}{\opsenvx \text{isnull}(e) \mapsto \textit{false}, E} &
\infrule[OP-ISNULL-TRUE]{\opsenvx e \mapsto null, E}{\opsenvx \text{isnull}(e) \mapsto true, E}
\end{tabular}
\end{minipage}
% Row 4: let, add, compare
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{6cm}m{6cm}m{6cm}}
\infrule[OP-LET]{\opsenvx e_1 \mapsto v_1, E_1 \\ L' = L[v_1/id] \\ S' = \text{apply}(S, E_1) \\ \opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E'} &
\infrule[OP-ADD]{\opsenvx e_1 \mapsto v_1, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\ v' = v_1 + v_2 \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx e_1 + e_2 \mapsto v', E'} &
\infrule[OP-COMPARE]{\opsenvx e_1 \mapsto v_1, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\ v' = \begin{cases} true, & \text{if $v_1 < v_2$}. \\ \textit{false}, & \text{otherwise}. \end{cases} \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx e_1 < e_2 \mapsto v', E'}
\end{tabular}
\end{minipage}
% Row 5: if-else, if-else, new color
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{6.5cm}m{6.5cm}m{6cm}}
\infrule[OP-IF-ELSE-TRUE]{\opsenvx e_1 \mapsto true, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2}{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2} &
\infrule[OP-IF-ELSE-FALSE]{\opsenvx e_1 \mapsto \textit{false}, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_3 \mapsto v_3, E_3}{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3} &
\infrule[OP-NEW-COLOR]{K = \{ (l_1, iv_1), \ldots, (l_p, v_p) \}, \text{ where } \\ (\forall i \in [1,p]. l_i \in M(r)) \wedge (\forall i,j \in [1,p]. l_i \not= l_j)}{\opsenvx \text{newcolor }r \mapsto K, []}
\end{tabular}
\end{minipage}
% Row 6: partition, color
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{8.5cm}m{10.5cm}}
\infrule[OP-PARTITION]{\opsenv{e_1}{K}{E_1} \andalso \rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
  M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'} &
\infrule[OP-COLOR]{\opsenvx e_1 \mapsto K, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto l, E_2 \andalso S'' = \text{apply}(S', E_2) \\
  \opsenvx[S=S''] e_3 \mapsto v, E_3 \\
  K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
  valid\_interleave(S, C, E', E_1, E_2, E_3)}
  {\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'} 
\end{tabular}
\end{minipage}
% Row 7: upregion, downregion, read
\begin{minipage}{\linewidth}
\vspace{-0.2cm}
\begin{tabular}{m{6cm}m{6.5cm}m{6.5cm}}
\infrule[OP-UPREG]{\opsenvx e \mapsto v, E}{\opsenvx \text{upregion}(e, \oton{r}) \mapsto v, E} &
\infrule[OP-DOWNREG]{\opsenvx e \mapsto l, E \\ v = \begin{cases} l, & \text{if $\exists i, l \in M(r_i)$}. \\ null, & \text{otherwise}. \end{cases}}{\opsenvx \text{downregion}(e, \oton{r}) \mapsto v, E} &
\infrule[OP-READ]{\opsenvx e \mapsto l, E \andalso S' = \text{apply}(S, E) \\ v = \begin{cases} S'(l), & \text{if } l \not\in C \\ v' : H(l), & \text{otherwise} \end{cases}}{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}
\end{tabular}
\end{minipage}
% Row 8: write, reduce
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\begin{tabular}{m{9cm}m{9cm}}
\infrule[OP-WRITE]{\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]} &
\infrule[OP-REDUCE]{\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}
\end{tabular}
\end{minipage}
% Row 9: pack, unpack
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\begin{tabular}{m{8cm}m{10cm}}
\infrule[OP-PACK]{\opsenvx e_1 \mapsto v, E \andalso \rho_i = M[r_i], \text{ for } 1 \leq i \leq k \\
  v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle}
  {\opsenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E} &
\infrule[OP-UNPACK]{\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \andalso M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
  L' = L[v_1/id] \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E'}
\end{tabular}
\end{minipage}
% Row 10: task call
\begin{minipage}{\linewidth}
%\vspace{-0.5cm}
\infrule[OP-TASK-CALL]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = \text{apply}(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = \text{apply}(S_{n-1}, E_n) \\
  valid\_interleave(S, C, E', \oton{E}) \andalso \text{task }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
  M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = \text{apply}(S, E') \\
  C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
  \opsenvx[M=M',L=L',S=S'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})}
  {\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}
\end{minipage}
\end{center}
\end{adjustwidth}
\caption{Legion Core Operational Semantics}
\label{fig:semantics}
\end{figure*}

\section{Soundness of Effects}
\label{sec:soundness}

A key property of the Legion type system is that any expression that is well-typed is
guaranteed to access the heap only in ways permitted by the privileges under which it was typed.
A judgement $M \vdash E : \Phi$ holds if all the memory operations in memory trace $E$ are of the types and
to locations covered by the privileges in $\Phi$:
\[ 
\begin{array}{rl}
M \vdash & E : \Phi \Leftrightarrow \forall \epsilon \in E. \\
& (\epsilon = read(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reads(r) \in \Phi)\ \wedge \\
& (\epsilon = write(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge writes(r) \in \Phi)\ \wedge \\
& (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reduces_{id}(r) \in \Phi)
\end{array}
\]

Our goal is to show $M \vdash E : \Phi$ holds for the execution of an expression $e$ with static permissions $\Phi$.
As usual, even stating such a claim requires that corresponding parts of the initial static and dynamic environments be consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item $M \sim \Omega$, a region mapping satisfies the region constraints under which the expression was typed
\item $L \sim_H \Gamma$, local values have types consistent with the environment under which the expression was typed
\item $S \sim H$, the initial state of the store contains values whose types are consisten with the heap typing
\end{itemize}

Two additional properties are proven for each subexpression:
\begin{itemize}
\item $v \sim_H M \llbracket T \rrbracket$, the expression yields a value of the right type
\item $E \sim H$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these properties.

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
M \sim \Omega & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
L \sim_H \Gamma & \Leftrightarrow & \forall (id,v) \in L. v \sim_H M \llbracket \Gamma \rrbracket (id) \\
S \sim H & \Leftrightarrow & \forall (l,v) \in S. v \sim_H H(l) \\
\end{array} & 
\begin{array}{lll}
E \sim H & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow v \sim_H H(l))\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \Rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge v \sim_H \hat T_2) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
bv \sim_H bool \\
iv \sim_H int \\
null \sim_H \hat T@\rho \\
\end{array} &
\begin{array}{lll}
l \sim_H \hat T@\rho & \Leftrightarrow & l \in \rho \wedge H(l) \sim_H \hat T \\
\langle v_1, v_2 \rangle \sim_H \langle \hat T_1, \hat T_2 \rangle & \Leftrightarrow & (v_1 \sim_H \hat T_1) \wedge (v_2 \sim_H \hat T_2) \\
\langle \langle \oton{\rho}, v \rangle \rangle \sim_H \text{rr}[\oton{r}] \hat T\text{ where }\Omega & \Leftrightarrow & (v \sim_H \hat T[\rho_1/r_1, \ldots \rho_n/r_n]) \wedge (\{ (r_i, \rho_i) \} \sim \Omega) \\
K \sim_H coloring(\rho) & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\end{figure*}

\begin{thm}
\label{thm:effects}
\rm
If $\typeenv{e}{T}$ and $\opsenv{e}{v}{E}$  and $M \sim \Omega$, $L \sim_H \Gamma$ and $S \sim H$,
then
$v \sim_H M \llbracket T \rrbracket$, $E \sim H$ and $M \vdash E : \Phi$.
\end{thm}

We spare the reader the lengthy proof, which can be found in
\cite{LegionTypes12}, and merely outline the general strategy.  The
proof is a standard induction on the structure of the derivation in
the operational semantics.  The proof is organized into six pieces,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results we wish to show.

\texcomment{
For each of 
the 23 different forms a Legion expression can have, we assume the soundness of the subexpressions'
evaluations (provided we can show the consistency of the mappings and store used by that
subexpression) and prove that the result of the expression's evaluation is sound.  Although there
are a multitude of cases to be proven, many of them are similar, and benefit from
the use of the following lemmas:

\begin{lem}
\label{lemma:heapconst:apply}
If $S \sim H$ and $E \sim H$, then $\text{apply}(S, E) \sim H$.
\end{lem}

\begin{lem}
\label{lemma:heapconst:effects1}
If $E_1 \sim H$ and $E_2 \sim H$, then $E_1 \dplus E_2 \sim H$.
\end{lem}

\begin{lem}
\label{lemma:heapconst:effects2}
If $E_1 \sim H$ and $E_2 \sim H$ and $valid\_interleave(S, C, E', E_1, E_2)$, then $E' \sim H$.
\end{lem}

\begin{lem}
\label{lemma:effsound:effects1}
If $M \vdash E_1 : \Phi$ and $M \vdash E_2 : \Phi$, then $M \vdash E_1 \dplus E_2 : \Phi$.
\end{lem}

\begin{lem}
\label{lemma:effsound:effects2}
If $M \vdash E_1 : \Phi$ and $M \vdash E_2 : \Phi$ and $valid\_interleave(S, C, E', E_1, E_2)$, then $M \vdash E' : \Phi$.
\end{lem}

\begin{lem}
\label{lemma:omegaclosure}
$M \sim \Omega^*$ if and only if $M \sim \Omega$.
\end{lem}

\begin{lem}
\label{lemma:phiclosure}
$M \vdash E : \Phi^*$ if and only if $M \vdash E : \Phi$.
\end{lem}

The proofs of these lemmas can also be found here\cite{LegionTypes12}.  
}

\begin{itemize}

\item[$M \sim \Omega$] Three expressions have subexpressions that modify $M$ or $\Omega$ and therefore do not
trivially satisfy region mapping consistency.  For {\tt partition},
the consistency of the coloring preserves region mapping consistency
with respect to the constraints.  For {\tt unpack}, the consistency of
a region relation instance guarantees consistency of region mapping.
Finally, the body of a called function with an initially-empty set of
constraints, which are trivially satisfied.

\item[$L \sim_H \Gamma$] Four expressions have subexpressions that modify $L$, $\Gamma$, or $M$.  For
{\tt partition}, which only modifies $M$, the requirement that it not reuse existing
names ensures that $M \llbracket \Gamma \rrbracket$ does not change.  For {\tt let},
the value and type of the binding is obviously consistent, while the binding created in
an {\tt unpack} is less obviously so, requiring an induction over the type of the 
unpacked value to show equivalence under the new mapping.  The last case is the body of a
called function, which requires the same style of proof as for {\tt unpack} 
for each formal parameter.

\item[$S \sim H$] The heap typing consistency of all stores used in subexpressions follows directly from the following
lemma: If $S \sim H$ and $E \sim H$, then $\text{apply}(S, E) \sim H$.

\end{itemize}

With the necessary consistency conditions proven for a subexpression, the conclusion of the theorem follows for
the subexpression by the inductive hypothesis.  The remaining obligation is to prove the the three parts of the
conclusion for an entire expression:
\begin{itemize}

\item[$v \sim_H M \llbracket T \rrbracket$]    The consistency of 
{\tt upregion} is guaranteed by the type checking requirement of appropriate subregion
constraints and the mapping's consistency with those constraints, and {\tt downregion}'s
result is consistent because of the run time check.  
%(The requirement
%for appropriate constraints in the {\em downregion} typing rule is primarily for performance
%(i.e. knowing which part of the region tree to examine) and to catch downcasts that will never
%work at runtime.) 
The consistency of a {\tt read}'s result is trivial for an address in the
clobber set and uses the consistency of the store otherwise.  The consistency of a 
{\tt color}'s result depends on the pointer subexpression's consistency and the 
removal of any previous coloring of that location from the coloring set.

The remaining interesting cases arise from changes to the mapping $M$ rather than transformations on the
value $v$.  In the case of {\tt partition} and {\tt unpack}, the type system
guarantees that the subexpression's result cannot use the regions that were added to the
mapping, allowing the changes to the mapping to be ignored.  The last case is again the body
of a called function, and the same strategy that was used for the type consistency of the
formal parameters works in reverse for the function's result.

\item[$E \sim H$] The type consistency of the value in an expression's effect follows from two lemmas applied to the effects
of subexpressions:
if $E_1 \sim H$ and $E_2 \sim H$, then $E_1 \dplus E_2 \sim H$; and 
if $E_1 \sim H$ and $E_2 \sim H$ and $valid\_interleave(S, C, E', E_1, E_2)$, then $E' \sim H$.
New effects are added by {\tt write} and {\tt reduce} expressions, but consistency follows
directly from the induction hypothesis.  Finally, the consistency of the
values in a called function's memory trace is addressed in the same way as the return value.

\item[$M \vdash E : \Phi$] The proof of the crucial property of containment of effects within the available privileges
is similar in outline to the previous step.  The easy cases are covered by lemmas combining the corresponding results
for subexpressions: if $M \vdash E_1 : \Phi$ and $M \vdash E_2 : \Phi$, then $M \vdash E_1 \dplus E_2 : \Phi$, and
if $M \vdash E_1 : \Phi$ and $M \vdash E_2 : \Phi$ and $valid\_interleave(S, C, E', E_1, E_2)$, then $M \vdash E' : \Phi$.
Straightforward proofs cover {\tt read}, {\tt write}, {\tt reduce}, and one final special case for function calls. 
\end{itemize}

%% \begin{table*}
%% \centering
%% {\small
%% \begin{tabular}{l|p{5cm}|p{5cm}|p{5cm}}
%% {\bf Rule} & 
%% \multicolumn{1}{|c|}{$\mathbf{v \sim_H M \llbracket T \rrbracket}$} &
%% \multicolumn{1}{|c|}{$\mathbf{E \sim H}$} &
%% \multicolumn{1}{|c|}{$\mathbf{E : \Phi}$} \\
%% \hline
%% BoolConst & & & \\
%% IntConst & & & \\
%% TupleBuild & & & \\
%% TupleField & & & \\
%% LocalVar & & & \\
%% NullPtr & & & \\
%% NewPtr & & & \\
%% IsNull & & & \\
%% UpRegion & & & \\
%% DownRegion & & & \\
%% Read & & & \\
%% Write & & & \\
%% Reduce & & & \\
%% Newcolor & & & \\
%% Color & & & \\
%% IntMath & & & \\
%% IntCmp & & & \\
%% Let & & & \\
%% IfElse & & & \\
%% TaskCall & & & \\
%% Partition & & & \\
%% Pack & & & \\
%% Unpack & & &
%% \end{tabular}}
%% \caption{Proof pieces}
%% \end{table*}

\section{Coherence}
\label{sec:coherence}

Theorem~\ref{thm:effects} guarantees that a task produces a value of the
correct type and does so within the regions of memory for which privileges are held.
It does not, however, address {\em coherence}, what concurrent operations on the regions are permitted.
An innovation in Legion is that programmers can control coherence on a per-region basis.  Intuitively,
the different coherence modes permit different interleavings of memory traces.  For example,
in addition to be an interleaving, exclusive access requires interleaved memory traces to be:
\begin{itemize}
\item {\em coherent} - a read of location $l$ should see the effect of all writes and reductions to $l$
occuring earlier in the interleaved trace, and
\item {\em sequentially equivalent} - the value read by each read 
operation and the result of applying the interleaved trace to the original store should be the same as if each constituent trace
was applied to the store in a valid sequential execution of the program.
\end{itemize}
These requirements are relaxed for either a non-empty clobber set (representing
operations that may occur concurrently as part of evaluating other functions) or for coherence
modes other than {\em excl}.

\begin{figure*}
$$
\begin{array}{lll}
valid\_interleave(S, C, E', \oton{E}) & = & any\_interleave(E', \oton{E}) \wedge \\
& & coherent(S, L_{excl}(E', C), L_{excl}(E', C), E')\ \wedge \\
& & seq\_equiv(S, L_{excl}(E', C), L_{excl}(E', C), E', \oton{E})\ \wedge \\
& & coherent(S, L_{atomic}(E', C), \emptyset, E')\ \wedge \\
& & \exists \oton{\pi}, (\oton{\pi}) \text{ is a permutation of } (1,\ldots,n)\ \wedge \\
& & \ \ \ seq\_equiv(S, L_{atomic}(E', C), \emptyset, E', E_{\pi_1}, \ldots, E_{\pi_n}) \\
coherent(S, L_1, L_2, []) & = & \text{true} \\
coherent(S, L_1, L_2, \epsilon \dplus E) & = &\
\begin{cases}
(l \in L_2 \Rightarrow S(l) = v)\ \wedge \\
\ \ \ coherent(S, L_1, L_2, E), & \text{if $\epsilon = read(l, c, v)$} \\
coherent(apply(S, \epsilon), L_1, L_2 \cup \{ l \}, E), & \text{if $\epsilon = write(l, c, v)$ and $l \in L_1$} \\
coherent(apply(S, \epsilon), L_1, L_2, E), & \text{otherwise}
\end{cases} \\
seq\_equiv(S, L_1, L_2, E', \oton{E}) & = & coherent(S, L_1, L_2, E_1 \dplus \ldots \dplus E_2)\ \wedge \\
& & \forall l \in L_1. apply(S, E')(l) = apply(S, E_1 \dplus \ldots \dplus E_n)(l) \\
L_{excl}(E, C) & = & \{ l : op(l, excl, v) \text{ in } E  \} \setminus C \\
L_{atomic}(E, C) & = & \{ l : op(l, atomic, v) \text{ in } E \} \setminus (C \cup L_{excl}(E, C)) \\
\end{array}
$$
\caption{Valid Interleaving Test}
\label{fig:validinterleave}
\end{figure*}

Figure~\ref{fig:validinterleave} formally defines {\em
valid\_interleave}, which we briefly explain.  Consider
traces $\oton{E}$.  The {\em exclusive location set} $L_{excl}$ is the
set of locations used in an operation with coherence mode $excl$ in
some trace $E_i$, excepting any locations in the clobber set $C$.
Similarly, the {\em atomic location set} $L_{atomic}$ is the set of
locations in an operation with coherence mode $atomic$, but no
operations with coherence mode $excl$, again excepting any location in
the clobber set.  The $coherent(S,L_1,L_2,E)$ test checks that all
reads of locations in $L_2$, and any read of a location $l \in L_1$
that comes after a write to $l$, return the value most recently written
to the store.  (If $L_2 = \emptyset$ initially, atomic reads without a preceding write are not checked, effectively defering
these checks to point where traces ordered before $E$ are included.)  The $seq\_equiv$
test verifies every read in $E'$ with a location in $L_1$ 
is coherent if operations are performed in 
program order and that the store that results from applying $E'$
matches the store resulting from applying the
constituent traces in program order for all locations in $L_1$.  For
the atomic location set, we allow the constituent traces to be
permuted, but all locations must be checked with the same permutation.

Flexibility in interleaving operations is the key to achieving parallel execution, but
implementing the precise criteria for checking a valid interleaving is
much too expensive.  We now consider practical mechanisms that generate valid interleavings and 
under what conditions these approaches may be used.

%
% FIX ME   I don't understand this paragraph.  I didn't see this discussed
% earlier; I took that comment out.  I don't know what the pronoun ``the one in which''
% refers to.  The clobber set?  The interleaving?
%
The first option is to serialize the constituent traces in program
order.  Although each constituent trace can be assumed to be coherent,
that coherence is in the context of a clobber set that may differ from
the one in which the interleaving must occur. 
This difference can result in cases where no valid interleaving
exists.  
%
% FIX ME  Why?
%
However, we can show that as long as any valid interleaving
exists, the serialization of the constituent traces is valid.
To prove this result, we use the following lemma, the proof
of which follows directly from the definition of $apply$.

\begin{lem}
\label{lemma:applychain}
\rm
For any initial store $S$ and memory traces $E_1$ and $E_2$,
$$apply(S, E_1 \dplus E_2) = apply(apply(S, E_1), E_2).$$
\end{lem}

\begin{thm}
\label{thm:sequential}
\rm
Let $S_0$ be a store, $C$ a clobber set and $\oton{E}$ memory traces to be
interleaved.  Let $E' = E_1 \dplus \ldots \dplus E_n$, in program order.  If there exists any $E''$ that is a
valid interleaving of $\oton{E}$, then $E'$ is also valid interleaving of $\oton{E}$.
\end{thm}

\begin{proof}
To prove $valid\_interleave(S_0, C, E', \oton{E})$, we need to show the four conditions:
\begin{enumerate}
\item $any\_interleave(E', \oton{E})$ is proven by induction on
the length of $E_1$, followed by $E_2$, and so on.
\item $coherent(S_0, C, E')$ is shown by considering each constituent trace's subsequence in $E'$
in order, using Lemma~\ref{lemma:applychain} with the coherence of each $E_i$.
\item $seq\_equiv(S_0, C, E', \oton{E})$ follows trivially by matching the structure of $E'$
to the corresponding  pieces $\oton{E}$.
\item $coherent(S_0, C, E')$ is shown using the sequential equivalence of $E''$.  The value 
returned by each read in $E''$ is equal to the value in the store that results
from the application of all previous constituent traces and the operations before the read in its
own trace.  This is exactly the sequence of preceding operations of the corresponding read in $E'$,
and Lemma~\ref{lemma:applychain} allows us to transform the chain of $apply$ operations into a 
single $apply$ operation on the whole sequence, satisfying the coherence condition for the read
in $E'$.  This property applies for all reads in $E''$, and since $E'$ and $E''$ are
interleavings of the same traces, they have the same reads, so all reads
in $E'$ are coherent.
\end{enumerate}
\end{proof}

Next we examine how to modify the serialized interleaving to produce valid interleavings allowing parallel execution. 
We consider when we may swap the order of two adjacent memory operations in a valid interleaving.
\begin{itemize}
\item The two operations must have come from different constituent
traces, or the swap would violate $any\_interleave$.
\item If one is a read of
a location $l$ (that is not in the clobber set), the other may not be a write or reduce to $l$, 
as changing the order would almost certainly change the value read.
\item If one is a write of a location $l$, the other may not be a write or reduce to $l$,
as the store that would result from applying those operations in the opposite order may
differ in $l$.
\item Finally, if both operations are reductions to the same location, they may be swapped if they
use the same reduction function, as reduction functions are assumed to commute.  
%(We 
%follow the example of other work (i.e. DPJ) and leave proofs of commutativity to the programmer,
%possibly with the aid of other tools.)  
However, if the reduction functions have different names, swapping the two operations might result in a different store.
\end{itemize}

To summarize this, we define a {\em noninterference} operator on two effects that have been
annotated with which constituent trace they came from as follows:
$$
\begin{array}{l@{ }l}
op_1^{E_1}(l_1, c_1,& v_1)\ \nonint\ op_2^{E_2}(l_2, c_2, v_2) \Leftrightarrow E_1 \not= E_2\ \wedge \\
& \big( l_1 \not= l_2 \vee (op_1 = read \wedge op_2 = read)\ \vee \\
& (op_1 = reduce_{id_1} \wedge op_2 = reduce_{id_2} \wedge id_1 = id_2) \big)
\end{array}
$$

\begin{lem}
\label{lem:nonintswap}
\rm
Let $S$ be a store, $C$ a clobber set, $\oton{E}, E'_a, E'_b$ memory traces, and
$\epsilon_1, \epsilon_2$ two annotated memory operations such that $\epsilon_1 \nonint \epsilon_2$.
Then,
$valid\_interleave(S, C, E'_a \dplus \epsilon_1 \dplus \epsilon_2 \dplus E'_b, \oton{E})$
if and only if
$valid\_interleave(S, C, E'_a \dplus \epsilon_2 \dplus \epsilon_1 \dplus E'_b, \oton{E}).$
\end{lem}
\texcomment{
\begin{proof}
As before, we consider the three conditions needed to be a valid interleaving.  We describe the
implication in the ``forward'' direction, but the arguments are all symmetric.
\begin{enumerate}
\item If $E'_a \dplus \epsilon_1 \dplus \epsilon_2 \dplus E'_b$ is an interleaving of $\oton{E}$,
the proof must first ``pop off'' all of $E'_a$, leaving $\epsilon_1 \dplus \epsilon_2 \dplus E'_b$ for
the interleaved subtrace.  The next two steps must pop $\epsilon_1$ and $\epsilon_2$, leaving $E'_b$.
They cannot be in the same subtrace because $\epsilon_1 \nonint \epsilon_2$, so we could also pop
$\epsilon_2$ and then $\epsilon_1$, resulting in a proof that $E'_a \dplus \epsilon_2 \dplus \epsilon_1 \dplus E'_b$ is also an interleaving of the constituent traces.
\item For coherence, we must show the value read (if any) by $\epsilon_1$ does not change, the value
read (if any) by $\epsilon_2$ does not change, and that any read in $E'_b$ sees no change.  With the
help of Lemma~\ref{lemma:applychain}, we can eliminate the common $E'_a$ from all sequences by
using $S' = apply(S, E'_a)$.  Similarly to show that $apply(S', \epsilon_1 \dplus \epsilon_2 \dplus E) = apply(S', \epsilon_2 \dplus \epsilon_1 \dplus E)$ for any $E$, we need only to show that it holds
for $E = []$.  The proof is now reduced to showing the following:
\begin{enumerate}
\item $\epsilon_1 = read(l, c, v) \Rightarrow S'(l) = apply(S',\epsilon_2)(l)$ - since $\epsilon_1$
is a read, either $\epsilon_2$ must be a read (in which case $apply(S, \epsilon_2) = S'$) or must be
to a different location, so the result of applying just it to $S'$ cannot change the value in 
location $l$.
\item $\epsilon_2 = read(l, c, v) \Rightarrow apply(S', \epsilon_1)(l) = S'(l)$ - this is the same
as above, with $\epsilon_1$ and $\epsilon_2$ switched.
\item $apply(S', \epsilon_1 \dplus \epsilon_2) = apply(S', \epsilon_2 \dplus \epsilon_1)$ - we must
show that the two store operations $apply(\bullet, \epsilon_1)$ and $apply(\bullet, \epsilon_2)$
commute.  Modifications to two different locations commute, as do two reads (which make no change
to the store).  Finally, two reductions using the same function to the same location also commute.
\end{enumerate}
\item The proof that sequential equivalence is maintained comes down to showing the final store
is identical after the swap of the two operations, and this is merely a special case of the third
piece of the coherence proof.
\end{enumerate}
\end{proof}
}
The proof is included in \cite{LegionTypes12}.
We now extend the noninterference operator to entire memory traces, providing a way to say that
no operation in a trace interferes with any operation in another:
$$E_1 \nonint E_2 \Leftrightarrow \displaystyle\bigwedge_{\epsilon_1 \text{ in } E_1, \epsilon_2 \text{ in } E_2} \epsilon_1 \nonint \epsilon_2$$
and observe that this property permits arbitrary ordering (i.e. parallel execution) of the 
corresponding subexpressions.

\begin{lem}
\label{lem:noninteffects}
\rm
Let $S$ be an initial store, $C$ be a clobber set, $\oton{E}$ be memory traces such that $E_i \nonint E_j$
for every $1 \leq i < j \leq n$.  Then any trace $E'$ satisfying $any\_interleave(E', \oton{E})$
is a valid interleaving.
\end{lem}
\texcomment{
\begin{proof}
We will use Theorem~\ref{thm:sequential} and Lemma~\ref{lem:nonintswap} to make repeated swaps
of adjacent operations to transform $E'$ into $E_1 \dplus \ldots \dplus E_n$.  The symmetry of
Lemma~\ref{lem:nonintswap} is then used to show that the validity of $E_1 \dplus \ldots \dplus E_n$
under Theorem~\ref{thm:sequential} implies the validity of $E'$.

The swapping algorithm as follows: Associate with each operation in $E'$ the index of that operation
in $E_1 \dplus \ldots \dplus E_n$.  Any consectutive pair of operations $\epsilon_1$ and $\epsilon_2$
for which $index(\epsilon_1) > index(\epsilon_2)$ is a {\em misordered pair}.  For any such pair,
$\epsilon_1$ and $\epsilon_2$ will be from different constituent traces (if they were from the same
trace, their misordering would mean that $E'$ was not an interleaving of $\oton{E}$).  Every pair of
traces is noninterfering, so $\epsilon_1 \nonint \epsilon_2$ and they can be swapped while preserving
validity of the trace.  At most $n^2/2$ swaps are needed to eliminate all misordered pairs.
\end{proof}
}
We next extend noninterference to privileges.  We 
use the region mapping $M$ to translate the logical regions named in privileges into sets of
locations (i.e. physical regions), and define noninterference between individual privileges and
sets of privileges:
$$
\begin{array}{l@{ }l}
priv_1(r_1) & \nonint[M] priv_2(r_2) \Leftrightarrow (M(r_1) \cap M(r_2) = \emptyset)\ \vee \\
& (priv_1 = reads \wedge priv_2 = reads)\ \vee \\
& (priv_1 = reduces_{id_1} \wedge priv_2 = reduces_{id_2} \wedge id_1 = id_2)
\end{array}
$$
%% \begin{center}
%% \begin{tabular}{llc}
%% \multicolumn{1}{c}{$\phi_1$} & \multicolumn{1}{c}{$\phi_2$} & $\phi_1 \nonint[M] \phi_2$? \\
%% \multirow{3}{*}{$reads(r_1)$} & $reads(r_2)$ & true \\
%% & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% & $reduces_{id_2}(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% \cline{1-1}
%% \multirow{3}{*}{$writes(r_1)$} & $reads(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% & $reduces_{id_2}(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% \cline{1-1}
%% \multirow{3}{*}{$reduces_{id_1}(r_1)$} & $reads(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%% & $reduces_{id_2}(r_2)$ & $\begin{array}{c}id_1 = id_2 \vee \\ M(r_1) \cap M(r_2) = \emptyset\end{array}$ \\
%% \end{tabular}
%% \end{center}

$$\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2$$

\begin{lem}
\label{lem:nonintpriv}
\rm
Let $M$ be a region mapping and $E_1$ and $E_2$ two memory traces such that $M \vdash E_1 : \Phi_1$ and $M \vdash E_2 : \Phi_2$.  If $\Phi_1$ and $\Phi_2$ are noninterfering under $M$, then $E_1$ and $E_2$ must be noninterfering.
\end{lem}
\texcomment{
\begin{proof}
The proof is by contradiction.  Assume $E_1 \cancel\nonint E_2$.  Then there must be some $\epsilon_1$ in $E_1$ and some $\epsilon_2$ in $E_2$ such that $\epsilon_1 \cancel\nonint \epsilon_2$.  This requires that
$\epsilon_1$ and $\epsilon_2$ be operations to the same location $l$ and not both be reads or both
reductions using the same function name.  Assume $\epsilon_1 = read(l, c_1, v_1)$ and $\epsilon_2 = write(l, c_2, v_2)$.  The definition of $M \vdash E_1 : \Phi_1$ tells us that there must be some $r_1$ satisfying $l \in M(r_1) \wedge reads(r_1) \in \Phi_1$.  Similarly, there must be some $r_2$ satisfying $l \in M(r_2) \wedge writes(r_2) \in \Phi_2$.  Letting $\phi_1 = reads(r_1)$ and $\phi_2 = writes(r_2)$, we have $M(r_1) \cap M(r_2) \not= \emptyset$ (it contains at least $l$), so $\phi_1 \cancel{\nonint[M]} \phi_2$ and $\Phi_1 \cancel{\nonint[M]} \Phi_2$.  There are seven other cases to consider (using $id_1 \not= id_2$ for the reduce-reduce case), but all yield the same result.
\end{proof}
}
We are now able to state the key result that allows the Legion runtime to safely run subtasks in parallel.

\begin{thm}
\label{thm:parallelexec}
\rm
Let $\oton{e}$ be well-typed Legion expressions, each with its own privileges $\Phi_i$.
Let $M$ be a region mapping, $L$ a local value mapping,
$H$ a heap typing, and $S$ be an initial store satisfying
$M \sim \Omega$, $L \sim_H \Gamma$, and $S \sim H$.
If $\Phi_i \nonint[M] \Phi_j$ for $1 \leq i < j \leq n$, then any parallel execution of expressions
$\oton{e}$ results in a valid interleaving of memory operations.
\end{thm}

The proof follows directly from Lemmas~\ref{lem:noninteffects}
and \ref{lem:nonintpriv}.  This result holds even if the
clobber set $C$ is non-empty, allowing two ``locally independent''
subtasks to run in parallel even if one or both of them is interacting
(in a programmer-permitted way) with a third subtask.

We are now able to explain an important aspect of the Legion
implementation that is different from other systems and relies on the
soundness of privileges.  Noninterference of memory effects is
something that can only be determined after evaluation of an
expression is completed, and only at great expense, as illustrated at
one extreme by work on transactional
memory \cite{Harris05}.  At the other extreme are systems
that check privileges statically, such as Jade \cite{Rinard98} and
DPJ \cite{Bocchino11}.  In contrast, Legion checks noninterference of privileges
at runtime, which is much simpler and more efficient than checking
noninterference of dynamic effects.  Even though the privileges
themselves are static, the region mapping $M$ is dynamic.  Testing
noninterference on the privileges with physical regions allows
parallel execution in many more cases than a purely static analysis
can achieve.

\subsection{Relaxed Coherence}
\label{subsec:relaxed}

The analysis above does not consider the coherence modes of operations,
making the worst-case assumption that all operations use $excl$ mode.  However, if two or more
subexpressions contain calls to tasks using $atomic$ or $simult$ coherence modes
on some regions, executions besides serial program order may 
result in valid interleavings even if constituent traces have locations in common.  We discuss
two cases of interest.  In both the previous sections noninterference % Is noninterference the right subject here?
is repeated with minor modifications, so for brevity we limit
discussion to a description of the differences rather than complete proofs.

The first important case occurs when all apparently-interfering operations have coherence modes
of either $atomic$ or $simult$.  In this case, the conditions required for
coherence and sequential equivalence are trivially satisfied for locations in $L_{excl}$.  The
requirements for locations in $L_{atomic}$ allow any permutation of the constituent traces,
so the runtime is free to serialize the execution of the subexpressions in any order, not just
program order.

To convert the required analysis to one that can be performed on regions, we modify
the noninterference operator for effects to consider only locations in $L_{excl}$ and show that
traces that are noninterfering under this modified check satisfy the requirement for atomic
execution.  We then modify the noninterference operator on privileges to allow 
$M(r_1) \cap M(r_2) \not= \emptyset$ if either $atomic(r_1)$ or $simult(r_1)$ is in $Q_1$ and
either $atomic(r_2)$ or $simult(r_2)$ is in $Q_2$. We then show that subexpressions with 
privileges that are {\em atomically noninterfering} (under a dynamic region mapping
$M$) always have effects compatible with atomically-reordered execution.  Finally,
in the special case where each subexpression has $atomic(r)$ or $simult(r)$ in $Q_i$ for every $r$
in $\Phi_i$, we can eliminate the dynamic check entirely.

The second case is a subset of the first, in which all apparently-interfering operations have
the $simult$ coherence mode.  In this case, all the coherence and sequential equivalence 
checks are trivially satisfied, and any interleaving of the constituent traces is valid.
We again modify the noninterference criteria for effects, traces, and privileges, this time
only allowing $simult$ coherence annotations. We show that subexpression with privileges
that are {\em simultaneously noninterfering} (under mapping $M$) can be executed
in parallel, which includes all the cases covered by Theorem~\ref{thm:parallelexec} as well
as those where the programmer has used the $simult$ coherence mode to explicitly allow parallel
execution.  Again, a special (but important) case exists where all privileges for each
subexpression are paired with $simult$ coherence mode declarations, allowing the safety of
parallel execution to be determined at compile time.

\section{Hierarchical Scheduling}
\label{sec:scheduling}

The enforcement of a (locally) valid interleaving at every step in the operational semantics
captures the hierarchical nature of the scheduler used by the Legion runtime.  As the cost of
a scheduling decision is quadratic in the number of tasks (more precisely, the number of 
privileges for those tasks), it should be clear that a hierarchical scheduling algorithm will
be vastly more efficient than a global algorithm that must consider the privileges of every
concurrently executing expression.
However, a scheduling algorithm that makes decisions hierarchically must do so with limited
information, and must respect all required scheduling dependencies while ideally imposing as
few extra dependencies as possible.  The following theorem forms the basis of the 
Legion scheduler.

\texcomment{
Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$.  Now let $e_3$ be some subexpression of $e_1$ and $e_4$ be some 
subexpression of $e_2$, with corresponding privileges $\Phi_3$ and $\Phi_4$.
By Theorem~\ref{thm:parallelexec}, if the runtime can verify $\Phi_1 \nonint[M] \Phi_2$, then 
expressions $e_1$ and $e_2$ may be evaluated in parallel.  The evaluation of the subexpressions
$e_3$ and $e_4$ will therefore be allowed to run in parallel, and we must show that this is
permissible.  We do so by proving that any memory traces that can result from the evaluation
of the subexpression must be noninterfering (allowing parallel execution by
Lemma~\ref{lem:noninteffects}).
}
\begin{thm}
\label{thm:hiersched}
\rm
Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$ respectively .  Let $e_3$ be a subexpression of $e_1$ and $e_4$ be a 
subexpression of $e_2$.  Any memory traces $E_3$ of $e_3$ and $E_4$ of $e_4$ resulting from evaluation of
$e_1$ and $e_2$ (with the usual consistent $M$, $L$, $H$, and $S$), must be
noninterfering.
\end{thm}
\texcomment{
\begin{proof}
Let $E_1$ be the result of an evaluation of $e_1$, with $E_3$ being the memory trace from the
evaluation of $e_3$ within the expression tree.  $E_1$ will be formed from a tree of valid
interleavings, each of which must include all the memory operations of each consitituent trace.
By induction over the number of intermediate subexpressions between $e_1$ and $e_3$, we
can show that any memory operation $\epsilon_3$ that is in $E_3$ must also be in $E_1$.
Similarly, any memory operation $\epsilon_4$ that is in $E_4$ must also be in $E_2$.  We have
$E_1 \nonint E_2$ from Lemma~\ref{lem:nonintpriv}, so and therefore $\epsilon_3 \nonint \epsilon_4$.  Since
this holds for all $\epsilon_3$ in $E_3$ and all $\epsilon_4$ in $E_4$, we have $E_3 \nonint E_4$.
\end{proof}
} The Legion task scheduler uses Theorem~\ref{thm:hiersched} in the
following way.  Sibling task calls (tasks invoked within the same
function body) are checked for non-interference of their (dynamic)
privileges.  If they interfere they are serialized; otherwise they are
considered for parallel execution.  Although determining that $e_1$
and $e_2$ are noninterfering guarantees $e_3$ and $e_4$ are
noninterfering, the converse is not true.  It is possible for $e_1$
and $e_2$ to be judged to be potentially interfering and be serialized
even when the subexpressions are noninterfering.  This loss of
precision can impact performance, and did in a few instances in our
experiments.  
%When the performance impact is significant, two options
%exist to improve matters.  
In cases where $e_1$ and $e_2$ are in fact
noninterfering, region partitions can be further refined to make the
disjointness visible in the privileges.  Alternately, the code can be
refactored to lift the subexpressions up to the same level, allowing
the dynamic test based on the region mapping to infer noninterference
in more cases.  
%Although this can increase the overhead of scheduling
%decisions, judicious use of this technique in practice has not shown
%any measurable performance degradation.