### 2012-07-10 doc/popl2013/example.tex
Major work on the example and some edits to the intro

\section{Circuit Example}
\label{sec:example}

Listing~\ref{lst:circuit_ex} shows code for a
circuit simulation written in the core Legion language introduced in
Section~\ref{sec:legioncore} (with one exception described below).  The
circuit simulation takes as input a large graph of circuit elements and wires
connecting the elements.  To perform the simulation in parallel the graph
is partitioned into pieces.  After partitioning
the graph the simulation is run for a number of time steps.  Each time step
consists of three phases: calculating the current on each wire, distributing
charge to the nodes to which each wire is connected, and updating the voltage
of each node.

The top level function in the application is {\tt simulate\_circuit} (line 15).  The {\tt simulate\_circuit} 
function is parameterized on the regions {\tt rl}, {\tt rw}, and {\tt rn} (line 15) which 
implies that the when the function is invoked it must have privileges for the regions 
bound to those parameters.  Furthermore, these privileges must be for both reading 
and writing these regions (line 17).  The only logical regions which {\tt simulate\_circuit} will 
be allowed to access are these regions and their logical subregions.  This property will be 
enforced by the type system introduced in Section~\ref{sec:legioncore}.  {\tt simulate\_circuit}
takes pointers to the lists of nodes and wires that describe the graph.  The nodes
are stored in the {\tt rn} region while wires are stored in the {\tt rw} region.  The
lists live in the {\tt rl} region.

The first step in {\tt simulate\_circuit} is to partition the graph into pieces.  Partitioning
is done by constructing {\em colorings} using the {\tt color\_circuit} function (line 18-19).  A
coloring is a mapping from region elements to colors with each color corresponding to a new subregion
to be created.  The algorithm for performing a coloring is an application specific decision made by 
the programmer.  Note, that in the case of circuit, it would be impossible to
pick a single static partitioning scheme that would do well for all graphs.  Dynamic partitioning
provides the programmer the flexibility to make partitioning decisions based on input data.

The {\tt color\_circuit} function returns three colorings.  The first coloring is a total 
coloring that maps each node in the graph the piece that owns it.  The third coloring is a total
coloring that maps each wire in the graph to a piece which owns on of its nodes.  The second coloring captures a
different property of the graph: those nodes which must be shared between different pieces of the
graph to perform computations in parallel.  These are often referred to as {\em ghost nodes}.  We
describe how ghost node regions are used in conjunction with the owned nodes later.
Note that its possible some ghost nodes are shared between multiple pieces, and can be colored multiple ways.  Thus,
the second coloring is referred to as a {\em multicoloring} because it might map the same element
to multiple colors.

After creating the colorings, the application partitions the region into {\em subregions} based on 
the colorings (lines 21-25).  For simplicity, our example only partitions the circuit into two pieces.  Lines
21 and 23 create two different partitions with two different subregions of the {\tt rn} region.  The first partition is
{\em disjoint} because it is based on a coloring while the second partition is {\em aliased}
because it is based on a multicoloring which does not imply disjointness of subregions.  While 
aliased partitions based on multicolorings are not a part of core Legion described in Section~\ref{sec:legioncore}, 
they are functionally equivalent to creating single partitions for each color in the multicoloring.

Disjoint partitions introduce
disjointness constraints into the static environment indicating that all the subregions in the partition are 
disjoint (e.g. {\tt rn0}$*${\tt rn1} on line 21 says {\tt rn0} is disjoint from {\tt rn1}).
Both disjoint and aliased partitions introduce constraints into the environment indicating that all
created regions are logical subregions of the partitioned parent region (e.g. {\tt rn0}$<${\tt rn}).

The circuit example illustrates an example of partitioning called {\em allocate-then-partition}
where the data structure is allocated first and parititoned later.  Alternatively, a
{\em partition-then-allocate} scheme would allow the creation and paritioning of regions that
are later populated with data.  Sequoia supports a static form of the former while DPJ supports
a static form of the later.  Legion enables both approaches dynamically.

After partitioning the circuit into pieces, the application creates instances of {\tt CircuitPiece} 
(defined on lines 9-12) for each piece (lines 28-33).  {\tt CircuitPiece} is a bounded existential 
type called a {\em region relationship}.  
Region relationships allow the programmer to {\em pack} a group of regions together and remember properties about
them such as disjointness and sub-region relationships (lines 29 and 33).  The type system verifies 
the properties hold statically when packing so when region relationships are {\em unpacked} the same properties can
be reintroduced into the static checking environment (lines 40-41).  One important feature of Legion
is that privileges cannot be packed in a region relationship but can only be passed through function calls.

The {\tt execute\_time\_steps} function represents the primary loop of the simulation (lines 37-48).  Here
we see the importance of having multiple views onto the same logical region.  Both the the 
{\tt calc\_new\_currents} and {\tt distribute\_charge} function calls bind the owned region of a piece 
and the ghost region of a piece.  In the case of {\tt calc\_new\_currents} these regions only need
read privileges, allowing both instances of {\tt calc\_new\_currents} to be run in parallel.  In the
case of {\tt distribute\_charge} the privilege is for a reduction which can also be done 
in parallel because of the atomic and commutative nature of reductions.  If it were only possible to have a single
partition of the region {\tt rn} it would have been impossible to describe these data sharing patterns.

Listing~\ref{lst:circuit_leaf} shows the leaf functions for each phase of the computation.  Each
function iterates over the list of wires or nodes for its piece.  Each function specifies the privileges
that it must have for all of its regions (lines 2,14,23).  These privileges are statically checked
to match the operations performed inside of each function (e.g. read and write).  In the case of {\em reduce}
the privilege must also specify which reduction function is being used (line 14).

In addition to privileges, functions can also specify {\em coherence} on regions.  Coherence specifies what 
values the function is allowed to see from other functions using the same region.  If not otherwise specified,
coherence defaults to {\em exclusive} which means the function must appear to execute as if the whole program
were run sequentialy, which we refer to as {\em program order}.  Line 14 shows an example of a relaxed
coherence mode called {\tt atomic} which says that reductions to {\tt rn} and {\tt rg} must appear atomic relative to other
functions using those regions.  The most relaxed coherence mode is {\tt simult} which means multiple functions
all with simultaneous coherence are allowed to access the region concurrently.

% This is a description of how the listings should be formatted.
% It can go anywhere before the listings.
\lstset{
  captionpos=b,
  language=Haskell,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  morekeywords={function,rr,int,float,bool,isnull,partition,as,downregion,upregion,reads,writes,rdwrs,reduces,read,write,reduce,using,unpack,pack,coloring,multicoloring,color,newcolor,atomic,simultaneous},
  deletekeywords={float,head,min,max}
}

\begin{lstlisting}[float={t},label={lst:circuit_leaf},caption={Circuilt Leaf Functions}]
function calc_new_currents[rl,rw,rn,rg] ( ptr_list : wire_list<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn,rg), writes(rw) : bool =
  if isnull(ptr_list) then true else
  let wire_node : wire_list<rl,rw,rn,rg> = read(ptr_list) in
  let wire : CircuitWire<rn,rg> = read(wire_node.1) in
  let in_node : CircuitNode = read(wire.1) in
  let out_node : CircuitNode = read(wire.2) in
  let current : float = (in_node.1 - out_node.1) /  wire.3 in 
  let new_wire : CircuitWire<rn,rg> = <wire.1,wire.2,wire.3,current> in
  let _ : CircuitWire<rn,rg>@rw = write(wire_node.1, new_wire) in
      calc_new_currents[rl,rw,rn,rg](wire_node.2)

function distribute_charge[rl,rw,rn,rg] ( ptr_list : wire_list<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn), reduces(reduce_charge,rn,rg), atomic(rn,rg) : bool =
  if isnull(ptr_list) then true else
  let wire_node : wire_list<rl,rw,rn,rg> = read(ptr_list) in
  let wire : CircuitWire<rn,rg> = read(wire_node.1) in
  let _ : CircuitNode@rn = reduce(reduce_charge, wire.1, wire.4) in
  let _ : CircuitNode@(rn,rg) = reduce(reduce_charge, wire.2, wire.4) in
      distribute_charge[rl,rw,rn,rg](wire_node.2)

function update_voltage[rl,rn] ( ptr_list : node_list<rl,rn>@rl ), 
      reads(rl,rn), writes(rn) : bool = 
  if isnull(ptr_list) then true else
  let node_node : node_list<rl,rn> = read(ptr_list) in
  let node : CircuitNode = read(node_node.1) in
  let voltage : float = (node.3/node.4) in
  let new_node : CircuitNode = <voltage,node.2,node.3,node.4> in
  let _ : CircuitNode@rn = write(node_node.1, new_node) in
      update_voltage[rl,rn](node_node.2)

-- Reduction function for distribute charge
function reduce_charge ( node : CircuitNode, current : float ) : CircuitNode =
    let new_charge : float = node.3 + current in
        < node.1,new_charge,node.3,node.4>
\end{lstlisting}