### 2012-07-10 doc/popl2013/example.tex

\section{Circuit Example}
\label{sec:example}

Listing~\ref{lst:circuit_ex} shows a circuit simulation written (with
one exception) in the core Legion language (see
Section~\ref{sec:legioncore}).  The input is a graph of circuit
elements and connecting wires.  To perform parallel
simulation the graph is partitioned into pieces, after which the
simulation is run for a number of time steps.  Each time step has
three phases: calculate each wire's current, distribute charge to
the nodes to which each wire is connected, and update each node's voltage.

The top-level function is {\tt simulate\_circuit} (line 15), which is
parameterized on regions {\tt rl}, {\tt rw}, and {\tt rn} (line 15).
The wires and nodes live in regions {\tt rw} and {\tt rn},
respectively.  The {\tt rl} region holds lists of the nodes and wires.
When {\tt simulate\_circuit} is invoked it must have read and
write privileges for the region parameters
(line 17); the type system enforces that
{\tt simulate\_circuit} only accesses these regions and their
subregions.

Partitioning of the graph is done using {\em colorings} built in the
{\tt color\_circuit} function (line 18-19).  A coloring maps region
elements to colors, with each color corresponding to a new subregion
to be created.  Note that it is impossible to pick a single static
partitioning scheme that would do well for all graphs.  Dynamic
partitioning using colorings provides the programmer the flexibility
to make partitioning decisions based on input data.

The {\tt color\_circuit} function returns three colorings.  The first
maps each node to the graph piece that owns it.
The third maps each wire to a graph piece that
owns one of its nodes.  The second coloring captures, for each piece $p$,
the nodes in other pieces on the boundary of $p$, often referred to as
{\em ghost nodes}.  
%We describe how ghost node regions are used in
%conjunction with the owned nodes shortly.  
Note that 
ghost nodes may be on the boundary of multiple pieces and therefore colored
multiple ways.  The second coloring is a {\em multicoloring} because
it may map the same element to multiple colors.

After creating the colorings, the application partitions the region
into {\em subregions} according to the colorings (lines 21-25).  For
simplicity, our example only partitions the circuit into two pieces.
Lines 21 and 23 create two different partitions with two different
subregions of the {\tt rn} region.  The first partition uses a
coloring and is therefore {\em disjoint}; the second partition uses a
multicoloring and is {\em aliased}---the subregions may not be
disjoint. While aliased partitions based on multicolorings are not a
part of core Legion described in Section~\ref{sec:legioncore}, they
are functionally equivalent to creating single partitions for each
color in the multicoloring.
Disjoint partitions introduce region disjointness constraints into the static
environment indicating that all the subregions in the partition are
disjoint (e.g. {\tt rn0}$*${\tt rn1} on line 21 says {\tt rn0} is
disjoint from {\tt rn1}).  Both disjoint and aliased partitions
introduce constraints into the environment indicating that all created
regions are logical subregions of the partitioned parent region
(e.g. {\tt rn0}$\leq${\tt rn}).

The circuit simulation uses an {\em allocate-then-partition} style of
computation, where a large data structure is first allocated and then
partitioned.  Legion also supports {\em partition-then-allocate},
where empty regions are first partitioned and later populated with
data.  Both are useful; e.g., Sequoia \cite{Fatahalian06} and DPJ
\cite{Bocchino09} support static forms of allocate-then-partition and
partition-then-allocate, respectively.  Legion supports both
approaches dynamically.

After partitioning the circuit into pieces, the application creates
instances of {\tt CircuitPiece} (defined on lines 9-12) for each piece
(lines 28-33).  {\tt CircuitPiece} is a {\em region relationship}, a
bounded existential type.  Region relationships allow the programmer
to {\em pack} a group of regions together and remember properties
about them such as disjointness and subregion relationships (lines 29
and 33).  The type system verifies the properties hold statically when
packing; when region relationships are {\em unpacked} the same
properties are reintroduced into the static checking environment
(lines 40-41).  A key feature underlying the soundness of the Legion
type system is that privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the new
regions it finds in the packed value.  For example, on line 40
function {\tt execute\_time\_steps} has read/write privileges for
region {\tt rn0} because it has read/write privileges for {\tt rn} and
the {\tt CircuitPiece} region relationship constraints ensure $\tt rn0
\leq rn$ (line 12).

The {\tt execute\_time\_steps} function (lines 37-48)  
also illustrates the importance of using different partitions to give
multiple views onto the same logical region.  Both {\tt
  calc\_new\_currents} and {\tt distribute\_charge} 
use the owned and ghost regions of a piece, which are from different partitions. In
{\tt calc\_new\_currents} these regions only need read
privileges, allowing both instances of {\tt calc\_new\_currents} to be
run in parallel.  In {\tt distribute\_charge} the
privilege is for a reduction which can also be done in parallel
because of the atomic and commutative nature of reductions.  No
single partition of the nodes
describes these data sharing patterns.

Listing~\ref{lst:circuit_leaf} shows the leaf functions for each phase.
Each function iterates over the list of wires or
nodes for its piece.  Each function specifies the region privileges it
requires (lines 2,14,23).  
%These privileges
%are statically checked to match the operations performed inside of
%each function (e.g. read and write).  
In the case of {\em reduce} the
privilege must also specify which reduction function is used
(line 14).

In addition to privileges, functions can also specify {\em coherence}
on regions.  Coherence specifies what updates the function may
see from other functions using the same region.  If not otherwise
specified, coherence defaults to {\em exclusive}, meaning the
function must appear to execute in program order.  Line 14 shows
an example of {\tt atomic}, a relaxed coherence mode requiring
that reductions to {\tt rn} and {\tt rg} must appear atomic relative
to other functions using those regions.  The most relaxed coherence
mode is {\tt simult}; all functions with 
simultaneous coherence are allowed to access the region concurrently.

% This is a description of how the listings should be formatted.
% It can go anywhere before the listings.
\lstset{
  captionpos=b,
  language=Haskell,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  morekeywords={function,rr,int,float,bool,isnull,partition,as,downregion,upregion,reads,writes,rdwrs,reduces,read,write,reduce,using,unpack,pack,coloring,multicoloring,color,newcolor,atomic,simultaneous},
  deletekeywords={float,head,min,max}
}

\begin{lstlisting}[float={t},label={lst:circuit_leaf},caption={Circuilt Leaf Functions}]
function calc_new_currents[rl,rw,rn,rg] ( ptr_list : wire_list<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn,rg), writes(rw) : bool =
  if isnull(ptr_list) then true else
  let wire_node : wire_list<rl,rw,rn,rg> = read(ptr_list) in
  let wire : CircuitWire<rn,rg> = read(wire_node.1) in
  let in_node : CircuitNode = read(wire.1) in
  let out_node : CircuitNode = read(wire.2) in
  let current : float = (in_node.1 - out_node.1) /  wire.3 in 
  let new_wire : CircuitWire<rn,rg> = <wire.1,wire.2,wire.3,current> in
  let _ : CircuitWire<rn,rg>@rw = write(wire_node.1, new_wire) in
      calc_new_currents[rl,rw,rn,rg](wire_node.2)

function distribute_charge[rl,rw,rn,rg] ( ptr_list : wire_list<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn), reduces(reduce_charge,rn,rg), atomic(rn,rg) : bool =
  if isnull(ptr_list) then true else
  let wire_node : wire_list<rl,rw,rn,rg> = read(ptr_list) in
  let wire : CircuitWire<rn,rg> = read(wire_node.1) in
  let _ : CircuitNode@rn = reduce(reduce_charge, wire.1, wire.4) in
  let _ : CircuitNode@(rn,rg) = reduce(reduce_charge, wire.2, wire.4) in
      distribute_charge[rl,rw,rn,rg](wire_node.2)

function update_voltage[rl,rn] ( ptr_list : node_list<rl,rn>@rl ), 
      reads(rl,rn), writes(rn) : bool = 
  if isnull(ptr_list) then true else
  let node_node : node_list<rl,rn> = read(ptr_list) in
  let node : CircuitNode = read(node_node.1) in
  let voltage : float = (node.3/node.4) in
  let new_node : CircuitNode = <voltage,node.2,node.3,node.4> in
  let _ : CircuitNode@rn = write(node_node.1, new_node) in
      update_voltage[rl,rn](node_node.2)

-- Reduction function for distribute charge
function reduce_charge ( node : CircuitNode, current : float ) : CircuitNode =
    let new_charge : float = node.3 + current in
        < node.1,new_charge,node.3,node.4>
\end{lstlisting}