### 2012-07-16 doc/popl2013full/langdef.tex
proof

\newtheorem{thm}{Theorem}
\newtheorem{lem}{Lemma}

\newcommand{\oton}[1]{{#1}_1,\ldots,{#1}_n}
\newcommand{\otok}[2]{{#2}_1,\ldots,{#2}_{#1}}
\newcommand{\dplus}{\text{+\!+}}
\newcommand{\llbracket}{[\![}
\newcommand{\rrbracket}{]\!]}
\newcommand{\tuple}[2]{\langle #1, #2 \rangle}
\makeatletter
% macros for consistency flavors
\newcommand{\simH}{\sim_{\!H}}
% usually: \mapconsist{\Omega}
\newcommand{\mapconsist}[2][M]{#1 \sim #2}
% usually: \localconsist{L}{\Gamma}
\newcommand{\localconsist}[3][M]{#2 \simH \@ifmtarg{#1}{#3}{#1 \llbracket #3 \rrbracket}}
% usually: \storeconsist{S}{H}
\newcommand{\storeconsist}[2]{#1 \sim #2}
% usually: \valueconsist{v}{T}
\newcommand{\valueconsist}[3][M]{#2 \simH \@ifmtarg{#1}{#3}{#1 \llbracket #3 \rrbracket}}
% usually: \traceconsist{E}
\newcommand{\traceconsist}[2][H]{#2 \sim #1}
% usually: \privconsist{E}{\Phi}
\newcommand{\privconsist}[3][M]{#2 \@ifmtarg{#1}{:}{:_{\!#1}} #3}

\newcommand{\nonint}[1][]{\#\@ifmtarg{#1}{}{_{\!\!#1}}}
\makeatother

% override BCP's typesetrule function to left-justify multiple lines of hypotheses
\renewcommand{\typesetrule}[2]{%
   \setrulebody{%
      \frac{\begin{array}{@{}l@{}}#1\end{array}}%
           {\begin{array}{@{}l@{}}#2\end{array}}}}

% fun latex tricks to make typeenv and opsenv more friendly
\makeatletter
\define@key{typeenv}{G}{\def\typeenv@G{#1}}
\define@key{typeenv}{P}{\def\typeenv@P{#1}}
\define@key{typeenv}{O}{\def\typeenv@O{#1}}
\newcommand{\typeenvx}[1][]{
{
% default values
\def\typeenv@G{\Gamma}
\def\typeenv@P{\Phi}
\def\typeenv@O{\Omega}
\setkeys{typeenv}{#1}
\typeenv@G, \typeenv@P, \typeenv@O \vdash \,
}}
\newcommand{\typeenv}[3][]{\typeenvx[#1] {#2} : {#3}}
\define@key{opsenv}{M}{\def\opsenv@M{#1}}
\define@key{opsenv}{L}{\def\opsenv@L{#1}}
\define@key{opsenv}{H}{\def\opsenv@H{#1}}
\define@key{opsenv}{S}{\def\opsenv@S{#1}}
\define@key{opsenv}{C}{\def\opsenv@C{#1}}
\newcommand{\opsenvx}[1][]{
{
% default values
\def\opsenv@M{M}
\def\opsenv@L{L}
\def\opsenv@H{H}
\def\opsenv@S{S}
\def\opsenv@C{C}
\setkeys{opsenv}{#1}
\opsenv@M, \opsenv@L, \opsenv@H, \opsenv@S, \opsenv@C \vdash \,
}}
\newcommand{\opsenv}[4][]{\opsenvx[#1] {#2} \mapsto {#3}, {#4}}
\makeatother

\begin{figure*}
\centering
{\small
\begin{tabular}{cclr|cclr}

$T$ & ::= &  & types & $bv$ & ::= & false $\;\;\;\mid\;\;\;$ true & \\
  &$\mid$& bool $\;\;\;\mid\;\;\;$ int & base types & & & & \\
  &$\mid$& $\langle T_1, T_2 \rangle$ & tuple & $iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ & \\
  &$\mid$& $T@(\oton{r})$ & pointer & & & & \\
  &$\mid$& $\text{coloring}(r)$ & region coloring & $e$ & ::= & & expressions \\
  &$\mid$& $\exists \oton{r}. T\text{ where }\Omega$ & region relationship &   &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
  &$\mid$& $\forall \oton{r}. (\oton{T}), \Phi, Q \rightarrow T_r$ & functions &   &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
& & & &   &$\mid$& $id$ &  \\
$\Omega$ & ::= & $\{ \oton{\omega} \}$ & region constraints &   &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
$\omega$ & ::= & $r_1 \leq r_2$ & subregion   &   
  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ & \\
  &$\mid$& $r_1 * r_2$ & disjointness &   
  &$\mid$ & $\text{downregion}(e, r_1,\ldots,r_n)$ & \\  
& & &  & 
  &$\mid$& $\text{read}(e_1)$ $\;\mid\;$ $\text{write}(e_1, e_2)$ & memory access \\
$\Phi$ & ::= & $\{ \oton{\phi} \}$ & privileges & 
  & $\mid$ & $\text{reduce}(id, e_1, e_2)$ & \\
$\phi$ & ::= & reads$(r)$ $\;\mid\;$ writes$(r)$ $\;\mid\;$ $\text{reduces}_{id}(r)$ &  &  
  &$\mid$& $\text{newcolor}\ r$ $\;\mid\;$ $\text{color}(e_1, e_2, e_3)$ & coloring \\
& & & &  
  &   $\mid$& $e_1 + e_2$ & integer ops \\
$Q$ & ::= & $\{ \oton{q} \}$ & coherence modes  &   
       & $\mid$& $e_1 < e_2$ & comparisons \\
$q$ & ::= & atomic$(r)$ $\;\;\;\mid\;\;\;$ simult$(r)$ & &   
  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
& & & &   
       & $\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
$v$ & ::= & & values &   
       & $\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & function calls \\
  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & base values &   
       & $\mid$ & $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
  &$\mid$& $\langle v_1, v_2 \rangle$ & tuple & 
       & $\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
  &$\mid$& null $\;\;\;\mid\;\;\;$ $a$ & address & 
       & $\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\
  &$\mid$& $\{ (a_i, iv), \ldots \}$ & coloring & 
    & & & \\
  &$\mid$& $\langle \langle \oton{\rho}, v\rangle \rangle$ & reg. relation instance & 
   & & & \\
%$bv$ & ::= & false $\;\;\;\mid\;\;\;$ true \\
%\\
%$iv$ & ::= & 0 $\;\;\;\mid\;\;\;$ 1 $\ldots$ \\
%\\
%$e$ & ::= & & expressions \\
%  &$\mid$& $bv$ $\;\;\;\mid\;\;\;$ $iv$ & constants \\
%  &$\mid$& $\langle e_1, e_2 \rangle$ $\;\;\;\mid\;\;\;$ $e$.1 $\;\;\;\mid\;\;\;$ $e$.2 & tuple \\
%  &$\mid$& $id$ &  \\
%  &$\mid$& $\text{new}\ T@r$ $\;\;\;\mid\;\;\;$ $\text{null }T@r$ $\;\;\;\mid\;\;\;$ $\text{isnull}(e)$ & \\
%  &$\mid$& $\text{upregion}(e, r_1,\ldots,r_n)$ $\;\;\;\mid\;\;\;$ $\text{downregion}(e, r_1,\ldots,r_n)$ & \\
%  &$\mid$& $\text{read}(e_1)$ $\;\;\;\mid\;\;\;$ $\text{write}(e_1, e_2)$ $\;\;\;\mid\;\;\;$ $\text{reduce}(id, e_1, e_2)$ & memory access \\
%  &$\mid$& $\text{newcolor}\ r$ $\;\;\;\mid\;\;\;$ $\text{color}(e_1, e_2, e_3)$ & coloring operations \\
%  &$\mid$& $e_1 + e_2$ & integer operations \\
%  &$\mid$& $e_1 < e_2$ & comparison operations \\
%  &$\mid$& $\text{let}\ id : T = e_1 \text{in}\ e_2$ &  \\
%  &$\mid$& $\text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3$ &  \\
%  &$\mid$& $id[r_1, \ldots, r_n](e_1,\ldots,e_n)$ & function calls \\
%  &$\mid$& $\text{partition}\ r_p\text{ using }e_1\text{ as }\oton{r}\text{ in }\ e_2$ &  \\
%  &$\mid$& $\text{pack}\ e_1\ \text{as}\ T[r_1,\ldots,r_n]$ &  \\
%  &$\mid$& $\text{unpack}\ e_1\ \text{as}\ id : T[r_1,\ldots,r_n]\ \text{in}\ e_2$ &  \\
\end{tabular}
}
\caption{Core Legion}
\label{fig:langdef}
\vspace{-5mm}
\end{figure*}

%\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
%\newcommand{\cinfrule}[3]{\parbox{14cm}{\hfil$\infrule{#1}{#2}$\hfil}\parbox{4cm}{$\,#3$\hfil}}
%\newcommand{\finfrule}[2]{\framebox{$\infrule{#1}{#2}$}}
%\newcommand{\oldfinfrule}[2]{\vspace{10pt}\framebox{$\infrule{#1}{#2}$}\vspace{10pt}}

%\newcommand{\infx}[2]{\infrule{\begin{array}{l}{#1}\end{array}}{#2}}

\section{Legion}
\label{sec:legioncore}

Figure~\ref{fig:langdef} defines Core Legion, a subset of the full Legion language
that still illustrates all of the important issues.  Types
include booleans, integers, tuples, and pointers.  Pointers
are annotated with a list of regions---a non-null pointer must point to a 
location that is contained in at least one of the regions. There is a special
type for {\em colorings}, which
are used to specify partitions of regions into subregions.
Functions in Legion are named, accept one or more arguments and
return a value of some type.  A function also declares the region
access privileges and coherence it requires.  A function may be executed
within the current task or run in parallel as a separate subtask---this decision is made
by the Legion task scheduler.  Function types are
universally quantified over all region names appearing in the type.
The final type is a region relationship, an instance of which captures a value and
one or more regions satisfying associated constraints.  A region relation instance
can be written into the heap and later read and
unpacked, giving new local names to the regions contained in
the instance.

\subsection{Core Type System}
\label{subsec:coretypes}

Core Legion is explicitly typed using judgments of the form
$$\typeenv{e}{T}$$
where:
\begin{itemize}
\item $\Gamma$ is the type environment, mapping the names of local variables and global functions
to their types,
\item $\Phi$ is the region access privileges under which the expression is evaluated, and
\item $\Omega$ is the set of constraints that must hold between regions
\end{itemize}
As regions are hierarchical, privileges and/or constraints on a region often imply similar
privileges or constraints on subregions.  Figure~\ref{fig:closure} defines closure
operations for $\Phi^*$ and $\Omega^*$ that explicitly add these implied constraints.
The type checking rules are in Figure~\ref{fig:types}.  Many of the rules are 
straightforward, but several deserve more explanation.

Core Legion has no subtyping. A pointer's region is changed only via explicit
{\em upregion} and {\em downregion} operations.  {\em Upregion} calls can be inferred,
but {\em downregion} must be accompanied by a run-time check that the pointer really does
point into the subregion(s), so {\em upregion} is made explicit as well for simplicity.

Accesses to heap memory are performed via the {\em read}, {\em write}, and {\em reduce}
operations, which are only legal if the regions in a pointer's type are ones for which the
enclosing function has privileges.
Similarly, a call to another function is permitted only if the privileges required by the 
callee are a subset of those held by the caller.  There is no way to create privileges within
a Legion function - it is limited to privileges given to it by its caller.  (The
top-level task of a Core Legion program begins with a single, empty region from which other
regions can be partitioned and full read and
write privileges.)

The {\em partition} operation uses a {\em coloring} to split a region into 
disjoint subregions.  The subregions are given names within the body of the partition,
as well as constraints asserting subregions are pairwise disjoint and contained within the parent region.
%These constraints can be used locally, or packed into region relation instances that can be unpacked and used later.

The {\em pack} expression packs together a value and the corresponding logical regions,
provided the region relation's constraints hold.  The
corresponding {\em unpack} expression gives (new) names to those regions.  For simplicity, in Core Legion neither the 
{\em partition} nor the {\em unpack} expression allows shadowing of region names.

\begin{comment}
\begin{figure*}
\centering{
\framebox{$\typeenv{bv}{bool}$}
\framebox{$\typeenv{iv}{int}$}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\typeenvx e_2 : T_2
\end{array}}
{\typeenvx \langle e_1, e_2 \rangle : \langle T_1, T_2 \rangle}
\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.1}\ : T_1}
\finfrule{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.2}\ : T_2}
\finfrule{\Gamma(id) = T}{\typeenvx id : T}
\framebox{$\typeenvx \text{null }T@r : T@r$}
\framebox{$\typeenvx \text{new }T@r : T@r$}
\finfrule{\typeenvx e : T@(\oton{r})}{\typeenvx \text{isnull}(e) : bool}
\finfrule
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall i. \exists j, r'_i \leq r_j \in \Omega^* \\
\end{array}}
{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\typeenvx e : T@(r'_1, \ldots r'_k) \\
\forall j. \exists i, r_j \leq r'_i \in \Omega^* \\
\end{array}}
{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\forall i, reads(r_i) \in \Phi^*\end{array}}
{\typeenvx \text{read}(e_1) : T}
\finfrule
{\begin{array}{l}
\typeenvx e_1 : T@(\oton{r}) \\
\typeenvx e_2 : T \\
\forall i, writes(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})}
\finfrule
{\begin{array}{l}
\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \\
\typeenvx e_1 : T_1@(\oton{r}) \\
\typeenvx e_2 : T_2 \\
\forall i, reduces_{id}(r_i) \in \Phi^*
\end{array}}
{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}
\framebox{$\typeenvx \text{newcolor }r : \text{coloring}(r)$}
\finfrule{\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r) \\
\typeenvx e_2 : T@r \\
\typeenvx e_3 : int
\end{array}}
{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}
\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 + e_2 : int}
\finfrule{\begin{array}{l}\typeenvx e_1 : int \\ \typeenvx e_2 : int\end{array}}{\typeenvx e_1 < e_2 : bool}
\finfrule{\begin{array}{l}
\typeenvx e_1 : T_1 \\
\typeenvx[G={\Gamma[T_1/id]}] e_2 : T_2
\end{array}}
{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2}
\finfrule{\begin{array}{l}\typeenvx e_1 : bool \\ \typeenvx e_2 : T \\ \typeenvx e_3 : T\end{array}}{\typeenvx \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T}
\finfrule{
\begin{array}{l}
\Gamma(id) = \forall r'_1, \ldots r'_k.(\oton{T}),\Phi', Q' \rightarrow T_r \\
\typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\
\Phi'[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*
\end{array}}
{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]}
\finfrule{
\begin{array}{l}
\typeenvx e_1 : \text{coloring}(r_p) \\
\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\
\typeenvx[O=\Omega'] e_2 : T \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset
\end{array}}
{\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}
\finfrule{
\begin{array}{l}
T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\
\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\
\typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]
\end{array}}
{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1}
\finfrule{
\begin{array}{l}
T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \\
\typeenvx e_1 : T_1 \\
\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \\
\Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
\typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \\
\{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset
\end{array}}
{\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3}
\finfrule{
\begin{array}{l}
\text{for }1 \leq i \leq p, \\
\Gamma(id_i) = \forall r_{i1}, \ldots r_{ik_i}. (T_{i1}, \ldots, T_{in_i}), \Phi_i, Q_i \rightarrow T_{ir} \\
\Gamma_i = \Gamma[a_{i1}/T_{i1}, \ldots, a_{in_i}/T_{in_i}] \\
\typeenv[G={\Gamma_i},P={\Phi_i},O=\emptyset]{e_i}{T_{ir}}
\end{array}}
{ 
\begin{array}{l@{ }l}
\vdash \{ & \text{function }id_1[r_{11}, \ldots, r_{1k_1}]( a_{11} : T_{11}, \ldots a_{1n_1} : T_{1n_1} ), \Phi_1, Q_1 : T_{1r} : e_1, \\
          & \ldots \\
          & \text{function }id_p[r_{p1}, \ldots, r_{pk_1}]( a_{p1} : T_{p1}, \ldots a_{pn_p} : T_{pn_p} ), \Phi_p, Q_p : T_{pr} : e_p \} : \bullet
\end{array} }
}
\caption{Legion Core Type System}
\label{fig:types}
\end{figure*}
\end{comment}

\begin{figure*}
\begin{adjustwidth}{-4em}{-4em}
\begin{center}
% Row 1: bool, int, tuple1, tuple2
\begin{minipage}{\linewidth}
%\centering
\begin{tabular}{m{4cm}m{3.5cm}m{5cm}m{5cm}}
\infax[T-Bool]{\typeenv{bv}{bool}} & 
\infax[T-Int]{\typeenv{iv}{int}} & 
\infrule[T-Tuple1]{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.1}\ : T_1} & 
\infrule[T-Tuple2]{\typeenvx e : \langle T_1,T_2 \rangle}{\typeenvx e\text{.2}\ : T_2} 
\end{tabular}
\end{minipage}
% Row 2:  var, null, isnull, maketuple
\begin{minipage}{\linewidth}
\vspace{-0.4cm}
%\centering
\begin{tabular}{m{3cm}m{3cm}m{5cm}m{6cm}}
\infrule[T-Var]{\Gamma(id) = T}{\typeenvx id : T} &
\infax[T-Null]{\typeenvx \text{null }T@r : T@r} &
\infrule[T-IsNull]{\typeenvx e : T@(\oton{r})}{\typeenvx \text{isnull}(e) : bool} &
\infrule[T-MakeTuple]{\typeenvx e_1 : T_1 \andalso \typeenvx e_2 : T_2}{\typeenvx \langle e_1, e_2 \rangle : \langle T_1, T_2 \rangle}
\end{tabular}
\end{minipage}
% Row 3: new, newcolor, coloring
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
%\centering
\begin{tabular}{m{4cm}m{5cm}m{9cm}}
\infax[T-New]{\typeenvx \text{new }T@r : T@r} &
\infax[T-Newcolor]{\typeenvx \text{newcolor }r : \text{coloring}(r)} &
\infrule[T-Color]{\typeenvx e_1 : \text{coloring}(r) \andalso \typeenvx e_2 : T@r \andalso \typeenvx e_3 : int}{\typeenvx \text{color}(e_1, e_2, e_3) : \text{coloring}(r)}
\end{tabular}
\end{minipage}
% Row 4: let, add, compare
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
%\centering
\begin{tabular}{m{6cm}m{5cm}m{6cm}}
\infrule[T-Let]{\typeenvx e_1 : T_1 \\ \typeenvx[G={\Gamma[T_1/id]}] e_2 : T_2}{\typeenvx : \text{let}\ id : T_1 \text{in}\ e_2 : T_2} &
\infrule[T-Add]{\typeenvx e_1 : int \\ \typeenvx e_2 : int}{\typeenvx e_1 + e_2 : int} &
\infrule[T-Compare]{\typeenvx e_1 : int \\ \typeenvx e_2 : int}{\typeenvx e_1 < e_2 : bool}
\end{tabular}
\end{minipage}
% Row 5: upregion, downregion
\begin{minipage}{\linewidth}
\vspace{-0.4cm}
\begin{tabular}{m{9cm}m{10cm}}
\infrule[T-UpRgn]{\typeenvx e : T@(r'_1, \ldots r'_k) \andalso \forall i. \exists j, r'_i \leq r_j \in \Omega^*}{\typeenvx upregion(e,\oton{r}) : T@(\oton{r})} &
\infrule[T-DownRgn]{\typeenvx e : T@(r'_1, \ldots r'_k)
% \andalso \forall j. \exists i, r_j \leq r'_i \in \Omega^*
}{\typeenvx downregion(e,\oton{r}) : T@(\oton{r})}
\end{tabular}
\end{minipage}
% Row 6: read, write, reduce
\begin{minipage}{\linewidth}
\vspace{-0.3cm}
\begin{tabular}{m{5cm}m{5cm}m{8cm}}
\infrule[T-Read]{\typeenvx e_1 : T@(\oton{r}) \\ \forall i, reads(r_i) \in \Phi^*}{\typeenvx \text{read}(e_1) : T} &
\infrule[T-Write]{\typeenvx e_1 : T@(\oton{r}) \\ \typeenvx e_2 : T \\ \forall i, writes(r_i) \in \Phi^*}{\typeenvx \text{write}(e_1, e_2) : T@(\oton{r})} &
\infrule[T-Reduce]{\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1 \andalso \typeenvx e_2 : T_2 \\
 \typeenvx e_1 : T_1@(\oton{r}) \andalso \forall i, reduces_{id}(r_i) \in \Phi^*}{\typeenvx \text{reduce}(id, e_1, e_2) : T_1@(\oton{r})}
\end{tabular}
\end{minipage}
% Row 7: pack, if-else, task call
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{5cm}m{5cm}m{8cm}}
\infrule[T-Pack]{T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1 \\ \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^* \\ \typeenvx e_1 : T_2[r_1/r'_1,\ldots,r_k/r'_k]}{\typeenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] : T_1} &
\infrule[T-IfElse]{\typeenvx e_1 : bool \\ \typeenvx e_2 : T \\ \typeenvx e_3 : T}{\typeenvx \text{if}\ e_1\ \text{then}\ e_2\ \text{else}\ e_3 : T} &
\infrule[T-Call]{\Gamma(id) = \forall r'_1, \ldots r'_k.(\oton{T}),\Phi', Q' \rightarrow T_r \\ \typeenvx e_i : T_i[r_1/r'_1,\ldots,r_k/r'_k] \\ \Phi'[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*}{\typeenvx id[\otok{k}{r}](\oton{e}) : T_r[r_1/r'_1,\ldots,r_k/r'_k]}
\end{tabular}
\end{minipage}
% Row 8: unpack, partition
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{10cm}m{8cm}}
\infrule[T-Unpack]{T_1 = \exists r'_1, \ldots, r'_n.\ T_2\text{ where }\Omega_1 \andalso \typeenvx e_1 : T_1 \\
  \Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id] \andalso \Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \\
  \typeenvx[G=\Gamma',O=\Omega'] e_2 : T_3 \andalso \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T_1, T_3) = \emptyset}
  {\typeenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 : T_3} &
\infrule[T-Part]{\typeenvx e_1 : \text{coloring}(r_p) \\ \Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j \\ \typeenvx[O=\Omega'] e_2 : T \\ \{ \oton{r} \} \cap \textit{regions\_of}(\Gamma, T) = \emptyset}
  {\typeenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 : T}
\end{tabular}
\end{minipage}
% Row 9: program
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\infrule[T-Program]{\text{for }1 \leq i \leq p, \\
  \Gamma(id_i) = \forall r_{i1}, \ldots r_{ik_i}. (T_{i1}, \ldots, T_{in_i}), \Phi_i, Q_i \rightarrow T_{ir} \andalso
  \Gamma_i = \Gamma[a_{i1}/T_{i1}, \ldots, a_{in_i}/T_{in_i}] \andalso
  \typeenv[G={\Gamma_i},P={\Phi_i},O=\emptyset]{e_i}{T_{ir}}}
  {\begin{array}{@{}l@{}l@{}}
   \vdash \{ & \text{function }id_1[r_{11}, \ldots, r_{1k_1}]( a_{11} : T_{11}, \ldots a_{1n_1} : T_{1n_1} ), \Phi_1, Q_1 : T_{1r} : e_1, \\
           & \ldots \\
           & \text{function }id_p[r_{p1}, \ldots, r_{pk_1}]( a_{p1} : T_{p1}, \ldots a_{pn_p} : T_{pn_p} ), \Phi_p, Q_p : T_{pr} : e_p \} : \bullet
   \end{array}
  }
\end{minipage}
\end{center}
\end{adjustwidth}
\caption{Legion Core Type System}
\vspace{-5mm}
\label{fig:types}
\end{figure*}

\begin{figure}
\centering{
$\begin{array}{l}
\Omega \subseteq \Omega^* \\
r_i \leq r_j \in \Omega^*  \Rightarrow r_i \leq r_i \in \Omega^*\wedge r_j \leq r_j \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j \leq r_k \in \Omega^* \Rightarrow r_i \leq r_k \in \Omega^* \\
r_i \leq r_j \in \Omega^* \wedge r_j * r_k \in \Omega^* \Rightarrow r_i * r_k \in \Omega^* \\
r_i * r_j \in \Omega^* \Rightarrow r_j * r_i \in \Omega^* \\
\\
\Phi \subseteq \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reads(r_j) \in \Phi^* \Rightarrow reads(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge writes(r_j) \in \Phi^* \Rightarrow writes(r_i) \in \Phi^* \\
r_i \leq r_j \in \Omega^* \wedge reduces_{id}(r_j) \in \Phi^* \Rightarrow reduces_{id}(r_i) \in \Phi^* \\
reads(r) \in \Phi^* \wedge writes(r) \in \Phi^* \Rightarrow reduces_{id}(r) \in \Phi^* \\
\ \ \ \mbox{for every function identifier $id$}
\end{array}$
}
\caption{Privilege and Constraint Closure}
\vspace{-5mm}
\label{fig:closure}
\end{figure}

\subsection{Operational Semantics}
\label{subsec:opsemantics}

Core Legion's operational semantics rules have the following form:
$$\opsenv{e}{v}{E}$$
where: 
\begin{itemize}
\item $M$ is a mapping from logical regions $r_i$ to {\em physical regions} $\rho_i$, which are sets of concrete memory locations.
We write $\hat x = M \llbracket x \rrbracket$ for the application of $M$ to other objects (e.g., types), 
replacing all free logical region names $r_i$ in $x$ with corresponding physical regions $\rho_i$.  
(Note $M \llbracket x \rrbracket$ is ill-formed if $x$ contains a free logical region name $r_j \not\in dom(M)$.)

\item $L$ is a mapping of local variables to (immutable) values.

\item $H$ is a {\em heap typing}, assigning an invariant type to every memory location.  The heap typing
serves two important roles.  First, it is used to bound the values that can be used for reads when 
concurrent operation may make the exact value uncertain.  Second, it greatly simplifies the consistency
check of pointer values, allowing a simple test against the heap typing for a location in lieu of a
potentially unbounded walk through the store for complicated pointer data structures.

\item $S$ is a store, mapping memory locations $l_i$ to values.

\item $C$ is a {\em clobber set}, memory locations that may be accessed by concurrently executing operations.

\item $e$ is an expression.

\item $v$ is the result of evaluating $e$.

\item $E$ is a {\em memory trace} of the execution, a list of {\em memory operations}
in the order performed.  A memory operation $\epsilon$ is either
$read(l, c, v)$ (a read of location $l$ with coherence mode $c$, returning value $v$),
$write(l, c, v)$ (a write to location $l$ with coherence mode $c$, storing value $v$),
or $reduce_{id}(l, c, v)$ (a reduction into location $l$ with coherence mode $c$ using an
associative function $id$ with value $v$).  Coherence modes are one of ({\em excl}, {\em atomic}, {\em simult}).
\end{itemize}

Helper functions are used in some rules:
\begin{itemize}
\item $apply(S, E)$ computes the store that results from applying memory trace $E$ to initial store $S$.

\item $mark\_coh(E, \hat Q)$ records the coherence of operations in trace $E$ according to coherence annotations $\hat Q$.

\item $valid\_interleave(S, C, E', \oton{E})$ is true if $E'$ is a valid interleaving
of the memory traces $\oton{E}$.
\end{itemize}
Figure~\ref{fig:opsemfns} defines {\em apply} and {\em mark\_coherence}.  The precise definition of
{\em valid\_interleave} is given in Section~\ref{sec:coherence}.  For now, we need only that a valid interleaving
is at least an interleaving, containing each constituent trace as a subsequence and no other operations.

Some rules modeling concurrent execution are locally
non-deterministic, guessing values that result from
concurrent tasks.  For example, the rule for {\tt read(e)} returns an
arbitrary value $v'$ of the correct type if the location being read is
in the clobber set $C$, the set of locations concurrent tasks may
clobber (i.e., update).  The guess is eventually checked for
consistency by the $valid\_interleave$ predicate of some enclosing
expression, at a point where the task that wrote $v'$ is also local to
the expression being evaluated.

Another prescient operation is {\tt newcolor r}, which guesses initial
memory locations and their colors for a partition.  These locations
support partition-then-allocate, ensuring the memory mapping $M$ has
locations for the new subregions that can be allocated by {\tt new}.
As is usual, memory allocation is also non-deterministic, choosing any
location of the correct type in the region.  For simplicity we do not
require locations to be ``fresh''; our soundness results hold even if
{\tt new T@r} returns locations already in use, provided the heap
typing of the location is correct.

The most involved rule in Core Legion's semantics is for function calls.
The only regions accessible to the callee (through $M'$) are the region arguments.
Also, if any regions are used with relaxed coherence (either {\em atomic} or {\em simult} instead of
{\em excl}), those locations are added to the clobber set, allowing the
impact of concurrent modifications of the store to be handled by the caller.  
%This is done by
%annotating the called function's memory trace with the relaxed coherence on a per-location basis
%and using the {\em valid\_interleave} predicate to control what parallel execution is allowed.

\begin{figure*}
\centering{
$
\begin{array}{@{}l@{}}
\begin{array}{@{}l@{\hspace{-0.25in}}l@{}} 
\begin{array}{@{}lll}
apply(S, []) & = & S \\
apply(S, E \dplus [ read(l, c, v) ]) & = & apply(S, E) \\
apply(S, E \dplus [ write(l, c, v) ]) & = & apply(S, E)[v/l] \\
apply(S, E \dplus [ reduce_{id}(l, c, v) ]) & = & S'[id(S'(l), v)/l], \\
& & \text{ where } S' = apply(S, E)
\end{array} &
\begin{array}{@{}ll}
mark\_coh([], \hat Q) & = [] \\
mark\_coh([ op(l, c, v) ] \dplus E, \hat Q) & = [ op(l, c', v) ] \dplus mark\_coh(E, \hat Q), \\
\multicolumn{2}{@{}l}{ \hspace{1.0in} \text{ where } c' =
\begin{cases}
simult, & \text{if }\exists \rho. l \in \rho \wedge simult(\rho) \in \hat Q \\
atomic, & \text{if }\exists \rho. l \in \rho \wedge atomic(\rho) \in \hat Q \\
excl, & \text{otherwise}
\end{cases}}
\end{array}
\end{array} \\
\begin{array}{@{}lll}
any\_interleave([], [], \ldots, []) & = & true \\
any\_interleave([ \epsilon ] \dplus E', E_1, \ldots, [ \epsilon ] \dplus E_i, \ldots, E_n) & = & any\_interleave(E', E_1, \ldots, E_i, \ldots, E_n) \\
valid\_interleave(S, C, E', \oton{E}) & \Rightarrow & any\_interleave(E', \oton{E}) \\
\end{array}
\end{array}
$}
\caption{Operational Semantics Helper Functions}
\label{fig:opsemfns}
\vspace{-5mm}
\end{figure*}

\begin{comment}
\begin{figure*}
\centering{\small
\framebox{$\opsenvx bv \mapsto bv, []$}
\framebox{$\opsenvx iv \mapsto iv, []$}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = apply(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E'}
\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.1} \mapsto v_1, E} \hspace{1cm}
\finfrule
{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}
{\opsenvx e\text{.2} \mapsto v_2, E}
\finfrule
{L(id) = v}
{\opsenvx id \mapsto v, []}
\framebox{$\opsenvx \text{null }T@r \mapsto null, []$}
\finfrule
{\begin{array}{l}
l \in M(r) \\
H(l) = M \llbracket T \rrbracket
\end{array}}
{\opsenvx \text{new }T@r \mapsto l, []}
\finfrule
{\opsenvx e \mapsto l, E}
{\opsenvx \text{isnull}(e) \mapsto \textit{false}, E}
\finfrule
{\opsenvx e \mapsto null, E}
{\opsenvx \text{isnull}(e) \mapsto true, E}
\finfrule
{\opsenvx e \mapsto v, E}
{\opsenvx \text{upregion}(e, \oton{r}) \mapsto v, E}
\finfrule
{\begin{array}{l}
\opsenvx e \mapsto l, E \\
v = \begin{cases}
l, & \text{if $\exists i, l \in M(r_i)$}. \\
null, & \text{otherwise}.
\end{cases}
\end{array}}
{\opsenvx \text{downregion}(e, \oton{r}) \mapsto v, E}
\finfrule
{\begin{array}{l}
\opsenvx e \mapsto l, E \\
S' = \text{apply}(S, E) \\
v = \begin{cases}
S'(l), & \text{if } l \not\in C \\
v' : H(l), & \text{otherwise}
\end{cases}
\end{array}}
{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto l, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}
\finfrule{
\begin{array}{l@{}l}
K = \{ (l_1, iv_1), & \ldots, (l_p, v_p) \}, \text{ where } \\
& (\forall i \in [1,p]. l_i \in M(r)) \wedge \\
& (\forall i,j \in [1,p]. l_i \not= l_j)
\end{array}
}
{\opsenvx \text{newcolor }r \mapsto K, []}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto K, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto l, E_2 \\
S'' = \text{apply}(S', E_2) \\
\opsenvx[S=S''] e_3 \mapsto v, E_3 \\
K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
valid\_interleave(S, C, E', E_1, E_2, E_3)
\end{array}}
{\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = v_1 + v_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 + e_2 \mapsto v', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\
v' = \begin{cases}
true, & \text{if $v_1 < v_2$}. \\
\textit{false}, & \text{otherwise}.
\end{cases} \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx e_1 < e_2 \mapsto v', E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E'}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto true, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_2 \mapsto v_2, E_2
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto \textit{false}, E_1 \\
S' = \text{apply}(S, E_1) \\
\opsenvx[S=S'] e_3 \mapsto v_3, E_3
\end{array}}
{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3}
\finfrule
{\begin{array}{l}
\opsenvx e_1 \mapsto v_1, E_1 \\
S_1 = \text{apply}(S, E_1) \\
\opsenvx e_2 \mapsto v_2, E_2 \\
\ldots \\
S_n = \text{apply}(S_{n-1}, E_n) \\
valid\_interleave(S, C, E', \oton{E})
\vspace{1.5mm} \\
\text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \\
L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \\
S' = \text{apply}(S, E') \\
C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
\opsenvx[M=M',L=L',S=S'] e_{n+1} \mapsto v_{n+1}, E_{n+1}
\vspace{1.5mm} \\
E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \\
valid\_interleave(S, C, E'', E', E_{n+1})
\end{array}}
{\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}
\finfrule{
\begin{array}{l}
\opsenv{e_1}{K}{E_1} \\
\rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'}
\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto v, E \\
\rho_i = M[r_i], \text{ for } 1 \leq i \leq k \\
v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle
\end{array}}
{\opsenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E}
\finfrule{
\begin{array}{l}
\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \\
M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
L' = L[v_1/id] \\
S' = \text{apply}(S, E_1) \\
\opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \\
valid\_interleave(S, C, E', E_1, E_2)
\end{array}}
{\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E' }
}
\caption{Legion Core Operational Semantics}
\label{fig:semantics}
\end{figure*}
\end{comment}

\begin{figure*}
\begin{adjustwidth}{-4em}{-4em}
\begin{center}
% Row 1: bool, int, val
\begin{minipage}{\linewidth}
\begin{tabular}{m{6cm}m{6cm}m{6cm}}
\infax[E-Bool]{\opsenvx bv \mapsto bv, []} &
\infax[E-Int]{\opsenvx iv \mapsto iv, []} &
\infax[E-Null]{\opsenvx \text{null }T@r \mapsto null, []} 
\end{tabular}
\end{minipage}
% Row 2: tuple1, tuple2, make tuple
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{9cm}m{4.5cm}m{4.5cm}}
\infrule[E-MakeTuple]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S' = apply(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \langle e_1, e_2 \rangle \mapsto \langle v_1, v_2 \rangle, E'} &
\infrule[E-Tuple1]{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}{\opsenvx e\text{.1} \mapsto v_1, E} &
\infrule[E-Tuple2]{\opsenvx e \mapsto \langle v_1, v_2 \rangle, E}{\opsenvx e\text{.2} \mapsto v_2, E}
\end{tabular}
\end{minipage}
% Row 3: new, null, isnull, isnull
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{4cm}m{4.5cm}m{5cm}m{5cm}}
\infrule[E-New]{l \in M(r) \\ H(l) = M \llbracket T \rrbracket}{\opsenvx \text{new }T@r \mapsto l, []} &
\infrule[E-Var]{L(id) = v}{\opsenvx id \mapsto v, []} &
\infrule[E-IsNull-F]{\opsenvx e \mapsto l, E}{\opsenvx \text{isnull}(e) \mapsto \textit{false}, E} &
\infrule[E-IsNull-T]{\opsenvx e \mapsto null, E}{\opsenvx \text{isnull}(e) \mapsto true, E}
\end{tabular}
\end{minipage}
% Row 4: let, add, compare
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{6cm}m{6cm}m{6cm}}
\infrule[E-Let]{\opsenvx e_1 \mapsto v_1, E_1 \\ L' = L[v_1/id] \\ S' = \text{apply}(S, E_1) \\ \opsenvx[L=L',S=S'] e_2 \mapsto v_2, E_2 \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx \text{let }id : T = e_1\text{ in }e_2 \mapsto v_2, E'} &
\infrule[E-Add]{\opsenvx e_1 \mapsto v_1, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\ v' = v_1 + v_2 \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx e_1 + e_2 \mapsto v', E'} &
\infrule[E-Compare]{\opsenvx e_1 \mapsto v_1, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2 \\ v' = \begin{cases} true, & \text{if $v_1 < v_2$}. \\ \textit{false}, & \text{otherwise}. \end{cases} \\ valid\_interleave(S, C, E', E_1, E_2)}{\opsenvx e_1 < e_2 \mapsto v', E'}
\end{tabular}
\end{minipage}
% Row 5: if-else, if-else, new color
\begin{minipage}{\linewidth}
\vspace{-0.7cm}
\begin{tabular}{m{6.5cm}m{6.5cm}m{6cm}}
\infrule[E-IfElse-T]{\opsenvx e_1 \mapsto true, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_2 \mapsto v_2, E_2}{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_2, E_1 \dplus E_2} &
\infrule[E-IfElse-F]{\opsenvx e_1 \mapsto \textit{false}, E_1 \\ S' = \text{apply}(S, E_1) \\ \opsenvx[S=S'] e_3 \mapsto v_3, E_3}{\opsenvx \text{if }e_1\text{ then }e_2\text{ else }e_3 \mapsto v_3, E_1 \dplus E_3} &
\infrule[E-Newcolor]{K = \{ (l_1, iv_1), \ldots, (l_p, iv_p) \}, \text{ where } \\ (\forall i \in [1,p]. l_i \in M(r)) \wedge (\forall i,j \in [1,p]. l_i \not= l_j)}{\opsenvx \text{newcolor }r \mapsto K, []}
\end{tabular}
\end{minipage}
% Row 6: partition, color
\begin{minipage}{\linewidth}
\vspace{-0.6cm}
\begin{tabular}{m{8.5cm}m{10.5cm}}
\infrule[E-Part]{\opsenv{e_1}{K}{E_1} \andalso \rho_i = \{ l : (l, i) \in K \}, \text{ for } 1 \leq i \leq k \\
  M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[M=M',S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2 \mapsto v, E'} &
\infrule[E-Color]{\opsenvx e_1 \mapsto K, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto l, E_2 \andalso S'' = \text{apply}(S', E_2) \\
  \opsenvx[S=S''] e_3 \mapsto v, E_3 \\
  K' = \{ (l,v) \} \cup \{ (l_i,v_i) : (l_i,v_i) \in K \wedge l \not= l_i \} \\
  valid\_interleave(S, C, E', E_1, E_2, E_3)}
  {\opsenvx \text{color}(e_1, e_2, e_3) \mapsto K', E'} 
\end{tabular}
\end{minipage}
% Row 7: upregion, downregion, read
\begin{minipage}{\linewidth}
\vspace{-0.2cm}
\begin{tabular}{m{6cm}m{6.5cm}m{6.5cm}}
\infrule[E-UpRgn]{\opsenvx e \mapsto v, E}{\opsenvx \text{upregion}(e, \oton{r}) \mapsto v, E} &
\infrule[E-DownRgn]{\opsenvx e \mapsto l, E \\ v = \begin{cases} l, & \text{if $\exists i, l \in M(r_i)$}. \\ null, & \text{otherwise}. \end{cases}}{\opsenvx \text{downregion}(e, \oton{r}) \mapsto v, E} &
\infrule[E-Read]{\opsenvx e \mapsto l, E \andalso S' = \text{apply}(S, E) \\ v = \begin{cases} S'(l), & \text{if } l \not\in C \\ v' : H(l), & \text{otherwise} \end{cases}}{\opsenvx \text{read}(e) \mapsto v, E \dplus [ read(l, excl, v) ]}
\end{tabular}
\end{minipage}
% Row 8: write, reduce
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\begin{tabular}{m{9cm}m{9cm}}
\infrule[E-Write]{\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{write}(e_1, e_2) \mapsto l, E' \dplus [ write(l, excl, v) ]} &
\infrule[E-Reduce]{\opsenvx e_1 \mapsto l, E_1 \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[S=S'] e_2 \mapsto v, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{reduce}(id, e_1, e_2) \mapsto l, E' \dplus [ reduce_{id}(l, excl, v) ]}
\end{tabular}
\end{minipage}
% Row 9: pack, unpack
\begin{minipage}{\linewidth}
\vspace{-0.5cm}
\begin{tabular}{m{8cm}m{10cm}}
\infrule[E-Pack]{\opsenvx e_1 \mapsto v, E \andalso \rho_i = M[r_i], \text{ for } 1 \leq i \leq k \\
  v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle}
  {\opsenvx \text{pack}\ e_1\ \text{as}\ T_1[\otok{k}{r}] \mapsto v', E} &
\infrule[E-Unpack]{\opsenvx e_1 \mapsto \langle \langle \otok{k}{\rho} , v_1 \rangle \rangle, E_1 \andalso M' = M[\rho_1/r_1, \ldots, \rho_k/r_k] \\
  L' = L[v_1/id] \andalso S' = \text{apply}(S, E_1) \\
  \opsenvx[M=M',L=L',S=S'] e_2 \mapsto v_2, E_2 \andalso valid\_interleave(S, C, E', E_1, E_2)}
  {\opsenvx \text{unpack}\ e_1\ \text{as}\ id : T_1[\otok{k}{r}] \text{in}\ e_2 \mapsto v_2, E'}
\end{tabular}
\end{minipage}
% Row 10: task call
\begin{minipage}{\linewidth}
%\vspace{-0.5cm}
\infrule[E-Call]{\opsenvx e_1 \mapsto v_1, E_1 \andalso S_1 = \text{apply}(S, E_1) \andalso \opsenvx e_2 \mapsto v_2, E_2 \andalso \ldots \andalso S_n = \text{apply}(S_{n-1}, E_n) \\
  valid\_interleave(S, C, E', \oton{E}) \andalso \text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1} \\
  M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \} \andalso L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \} \andalso S' = \text{apply}(S, E') \\
  C' = C \cup \{ l : \exists \rho. atomic(\rho) \in M' \llbracket Q' \rrbracket \vee simult(\rho) \in M' \llbracket Q' \rrbracket \} \\
  \opsenvx[M=M',L=L',S=S',C=C'] e_{n+1} \mapsto v_{n+1}, E_{n+1} \andalso E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket) \andalso valid\_interleave(S, C, E'', E', E_{n+1})}
  {\opsenvx id[\otok{k}{r}](\oton{e}) \mapsto v_{n+1}, E''}
\end{minipage}
\end{center}
\end{adjustwidth}
\caption{Legion Core Operational Semantics}
\label{fig:semantics}
\end{figure*}

\section{Soundness of Privileges}
\label{sec:soundness}

A key property of the Legion type system is that any expression that is well-typed is
guaranteed to access the heap only in ways permitted by the privileges under which it was typed.
A judgment $\privconsist{E}{\Phi}$ holds if all the memory operations in memory trace $E$ are of the types and
to locations covered by the privileges in $\Phi$:
\[ 
\begin{array}{l}
\privconsist{E}{\Phi} \Leftrightarrow \forall \epsilon \in E. \\
\quad\quad (\epsilon = read(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reads(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = write(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge writes(r) \in \Phi)\ \wedge \\
\quad\quad (\epsilon = reduce_{id}(l, c, v) \Rightarrow \exists r. l \in M(r) \wedge reduces_{id}(r) \in \Phi)
\end{array}
\]

Our goal is to show $\privconsist{E}{\Phi}$ holds for any execution of an expression $e$ with static privileges $\Phi$.
As usual, even stating such a claim requires that corresponding parts of the initial static and dynamic environments be consistent.
For our results, three consistency properties are needed:
\begin{itemize}
\item $\mapconsist{\Omega}$, a region mapping satisfies the region constraints under which the expression was typed
\item $\localconsist{L}{\Gamma}$, local values have types consistent with the environment under which the expression was typed
\item $\storeconsist{S}{H}$, the initial state of the store contains values whose types are consistent with the heap typing
\end{itemize}

\noindent Two additional properties are proven for each subexpression:
\begin{itemize}
\item $\valueconsist{v}{T}$, the expression yields a value of the right type
\item $\traceconsist{E}$, values written to the store match the heap types of their locations
\end{itemize}
Figure~\ref{fig:constprop} gives the precise definition of these properties.

\begin{figure*}
\centering{\footnotesize
$
\begin{array}{l}
\begin{array}{ll}
\begin{array}{lll}
\mapconsist{\Omega} & \Leftrightarrow & (\forall r_i, r_j. r_i \leq r_j \in \Omega \Rightarrow M(r_i) \subseteq M(r_j))\ \wedge \\
              &                 & (\forall r_i, r_j. r_i * r_j \in \Omega \Rightarrow M(r_i) \cap M(r_j) = \emptyset) \\
\localconsist{L}{\Gamma} & \Leftrightarrow & \forall (id,v) \in L. \valueconsist[]{v}{M \llbracket \Gamma \rrbracket (id)} \\
\storeconsist{S}{H} & \Leftrightarrow & \forall (l,v) \in S. \valueconsist[]{v}{H(l)} \\
\end{array} & 
\begin{array}{lll}
\traceconsist{E} & \Leftrightarrow & (\forall l, c, v. write(l,c, v) \in E \Rightarrow \valueconsist[]{v}{H(l)})\ \wedge \\
& & (\forall id, l, v. reduce_{id}(l, v) \in E \Rightarrow \\
& & (M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1) \wedge H(l) = \hat T_1 \wedge \valueconsist[]{v}{\hat T_2}) \\
\end{array}
\end{array} \vspace{3mm} \\
\begin{array}{l@{\hspace{1cm}}l}
\begin{array}{l}
\valueconsist[]{bv}{bool} \\
\valueconsist[]{iv}{int} \\
\valueconsist[]{null}{\hat T@\rho} \\
\end{array} &
\begin{array}{lll}
\valueconsist[]{l}{\hat T@\rho} & \Leftrightarrow & l \in \rho \wedge H(l) = \hat T \\
\valueconsist[]{\tuple{v_1}{v_2}}{\tuple{\hat T_1}{\hat T_2}} & \Leftrightarrow & (\valueconsist[]{v_1}{\hat T_1}) \wedge (\valueconsist[]{v_2}{\hat T_2}) \\
\valueconsist[]{\langle \langle \oton{\rho}, v \rangle \rangle}{\text{rr}[\oton{r}]\ \hat T\text{ where }\Omega} & \Leftrightarrow & (\valueconsist[]{v}{T[\rho_1/r_1, \ldots \rho_n/r_n]}) \wedge (\mapconsist[\{ (r_i, \rho_i) \}]{\Omega}) \\
\valueconsist[]{K}{coloring(\rho)} & \Leftrightarrow & \forall l_1, v_1. (l_1, v_1) \in K \Rightarrow (l_1 \in \rho\ \wedge \\
& & \forall l_2, v_2. (l_2, v_2) \in K \Rightarrow (l_1 \not= l_2) \vee (v_1 = v_2)) \\
\end{array}
\end{array}
\end{array}
$
}
\caption{Consistency Properties}
\label{fig:constprop}
\vspace{-5mm}
\end{figure*}

\begin{thm}
\label{thm:effects}
\rm
If $\typeenv{e}{T}$ and $\opsenv{e}{v}{E}$  and $\mapconsist{\Omega}$, $\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$,
then
$\valueconsist{v}{T}$, $\traceconsist{E}$ and $\privconsist{E}{\Phi}$.
\end{thm}

The proof is a standard induction on the structure of the derivation
For each of 
the Core Legion expressions there are six separate parts to the proof,
one for each of the three consistency properties that must be shown
for the execution environment of subexpressions and one for each of
the three properties of the execution's results.
Many of the cases are similar, and benefit from
the use of the following straighforward lemmas:
\begin{lem}
\label{lemma:heapconst:apply}
\rm
If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.
\end{lem}

\begin{lem}
\label{lemma:heapconst:effects1}
\rm
If $\traceconsist{E_1}$ and $\traceconsist{E_2}$, then $\traceconsist{E_1 \dplus E_2}$.
\end{lem}

\begin{lem}
\label{lemma:heapconst:effects2}
\rm
If $\traceconsist{E_1}$ and $\traceconsist{E_2}$ and \\
$valid\_interleave(S, C, E', E_1, E_2)$, then $\traceconsist{E'}$.
\end{lem}

\begin{lem}
\label{lemma:effsound:effects1}
\rm
If $\privconsist{E_1}{\Phi}$ and $\privconsist{E_2}{\Phi}$, then $\privconsist{E_1 \dplus E_2}{\Phi}$.
\end{lem}

\begin{lem}
\label{lemma:effsound:effects2}
\rm
If $\privconsist{E_1}{\Phi}$ and $\privconsist{E_2}{\Phi}$ and \\ 
$valid\_interleave(S, C, E', E_1, E_2)$, then $\privconsist{E'}{\Phi}$.
\end{lem}

\begin{lem}
\label{lemma:omegaclosure}
\rm
$\mapconsist{\Omega^*}$ if and only if $\mapconsist{\Omega}$.
\end{lem}

\begin{lem}
\label{lemma:phiclosure}
\rm
$\privconsist{E}{\Phi^*}$ if and only if $\privconsist{E}{\Phi}$.
\end{lem}

We first give an outline of the idea behindthe proof.  For each subexpression of a compound expression,
there are three consistency properties we must prove to be able to apply induction: \\

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\mapconsist{\Omega}$ - Three expressions have subexpressions that modify $M$ or $\Omega$ and therefore do not
trivially satisfy region mapping consistency.  For {\tt partition},
the consistency of the coloring preserves region mapping consistency
with respect to the constraints.  For {\tt unpack}, the consistency of
a region relation instance guarantees consistency of region mapping.
Finally, the body of a called function uses an initially-empty set of
constraints, which are trivially satisfied.

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\localconsist{L}{\Gamma}$ -  Four expressions have subexpressions that modify $L$, $\Gamma$, or $M$.  For
{\tt partition}, which only modifies $M$, the requirement that it not reuse existing
names ensures that $M \llbracket \Gamma \rrbracket$ does not change.  For {\tt let},
the value and type of the binding is obviously consistent, while the binding created in
an {\tt unpack} is less obviously so, requiring an induction over the type of the 
unpacked value to show equivalence under the new mapping.  The last case is the body of a
called function, which requires the same style of proof as for {\tt unpack} 
for each formal parameter.

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\storeconsist{S}{H}$ -  The heap typing consistency of all stores used in subexpressions follows directly from the following
lemma: If $\storeconsist{S}{H}$ and $\traceconsist{E}$, then $\storeconsist{\text{apply}(S, E)}{H}$.

\vspace{4pt}
With the necessary consistency conditions proven for a subexpression, the conclusion of the theorem follows for
the subexpression by the inductive hypothesis.  The remaining obligation is to prove the the three parts of the
conclusion for an entire expression:

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\valueconsist{v}{T}$ - The consistency of 
{\tt upregion} is guaranteed by the type checking requirement of appropriate subregion
constraints and the mapping's consistency with those constraints, and {\tt downregion}'s
result is consistent because of the run time check.  
%(The requirement
%for appropriate constraints in the {\em downregion} typing rule is primarily for performance
%(i.e. knowing which part of the region tree to examine) and to catch downcasts that will never
%work at runtime.) 
The consistency of a {\tt read}'s result is trivial for an address in the
clobber set and uses the consistency of the store otherwise.  The consistency of a 
{\tt color}'s result depends on the pointer subexpression's consistency and the 
removal of any previous coloring of that location from the coloring set.

\hangindent=2\parindent \hangafter=1 \indent\indent\indent %
The remaining interesting cases arise from changes to the mapping $M$ rather than transformations on the
value $v$.  In the case of {\tt partition} and {\tt unpack}, the type system
guarantees that the subexpression's result cannot use the regions that were added to the
mapping, allowing the changes to the mapping to be ignored.  The last case is again the body
of a called function, and the same strategy that was used for the type consistency of the
formal parameters works in reverse for the function's result.

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\traceconsist{E}$ -  The type consistency of the values in an expression's memory
trace follows from the following lemma applied to the traces of subexpressions:
if $\traceconsist{E_1}$ and $\traceconsist{E_2}$ and $any\_interleave(E', E_1, E_2)$, then $\traceconsist{E'}$.
New memory operations are added by {\tt write} and {\tt reduce} expressions, but consistency follows
directly from the induction hypothesis.  Finally, the consistency of the
values in a called function's memory trace is addressed in the same way as the return value.

\vspace{4pt}
\hangindent=2\parindent \hangafter=1 \noindent %
$\privconsist{E}{\Phi}$ -  The proof of the crucial property of containment of heap accesses within the available privileges
is similar in outline to the previous step.  The easy cases are covered by a lemma combining the corresponding results
for subexpressions: if $\privconsist{E_1}{\Phi}$ and $\privconsist{E_2}{\Phi}$ and $any\_interleave(E', E_1, E_2)$, then $\privconsist{E'}{\Phi}$.
Straightforward proofs cover {\tt read}, {\tt write}, {\tt reduce}, and one final special case for function calls.

\begin{proof}
The proof of the theorem is by cases on the structure of the derivation of $e$.
\begin{itemize}

\item
Case [E-Bool]. By assumption, 
$$\opsenv{bv}{bv}{[]}$$ and
$$\typeenv{bv}{T}$$ and
$\mapconsist{\Omega}$ and
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply is [T-Bool], so $T = bool$.
Therefore $\valueconsist{bv}{bool}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Int].  
By assumption, 
$$\opsenv{iv}{iv}{[]}$$ and 
$$\typeenv{iv}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Int], so $T = int$.
Therefore $\valueconsist{iv}{int}$ and 
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$ all follow trivially.

\item Case [E-Null].
By assumption, 
$$\opsenv{\text{null\ }T@r}{null}{[]}$$ and 
$$\typeenv{\text{null}}{T'}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Null], so $T' = T@r$.
Therefore $\valueconsist{null}{T@r}$, $\traceconsist{[]}$ and $\privconsist{[]}{\Phi}$ all follow trivially.

\item
Case [E-MakeTuple].
By assumption, 
$$\opsenv{\langle e_1, e_2 \rangle}{\langle v_1,v_2 \rangle}{E'}$$ and 
$$\typeenv{\langle e_1, e_2 \rangle}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-MakeTuple], so $T = \langle T_1, T_2 \rangle$.

For the subexpression $e_1$, 
from the form of rule [E-MakeTuple], we know that $$\opsenv{e_1}{v_1}{E_1}$$
From the form of rule [T-MakeTuple], we know that $$\typeenv{e_1}{T_1}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from the form of rule [E-MakeTuple], we know that $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From the form of rule [T-MakeTuple], we know that $$\typeenv{e_2}{T_2}$$
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

From the preceding steps we can conclude that $\valueconsist{\langle v_1, v_2\rangle}{\langle T_1, T_2 \rangle}$.
Finally, from [E-MakeTuple] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$ and
Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by Lemma~\ref{lemma:effsound:effects2}
$\privconsist{E'}{\Phi}$.

\item Case [E-Tuple1].
By assumption, 
$$\opsenv{e\text{.1}}{v_1}{E}$$ and 
$$\typeenv{e\text{.1}}{T_1}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
From the form of [E-Tuple1] we know that
$$\opsenv{e}{\langle v_1, v_2 \rangle}{E}$$
The only type rule that can apply is [T-Tuple1], and from the form of the rule we know
$$\typeenv{e}{\langle T_1, T_2 \rangle}$$
By induction, we have $\valueconsist{\langle v_1, v_2 \rangle}{\langle T_1, T_2 \rangle}$, $\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$.
It follows immediately that $\valueconsist{v_1}{T_1}$ also holds.

\item Case [E-Tuple2] is symmetric to case [E-Tuple1].

\item Case[E-New].
By assumption, 
$$\opsenv{\text{new }T@r}{l}{[]}$$ and 
$$\typeenv{\text{new }T@r}{T@r}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-New] rule is that $l \in M(r)$ and $H(l) = M \llbracket T \rrbracket$,
which establishes $\valueconsist{l}{T}$.  
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.
\item Case [E-Var].
By assumption, 
$$\opsenv{id}{v}{[]}$$ and 
$$\typeenv{id}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The hypothesis of the [E-Var] rule is $L(id) = v$.
From $\localconsist{L}{\Gamma}$ we have $\valueconsist{L(id)}{\Gamma(id)}$.
The hypothesis of the [T-Var] rule is $\Gamma(id) = T$.
Putting these three facts together we conclude $\valueconsist{v}{T}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item Case [E-IsNull-F].
By assumption, 
$$\opsenv{\text{isnull}(e)}{\textit{false}}{E}$$ and 
$$\typeenv{\text{isnull}(e)}{bool}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
From the form of [E-IsNull-F] we know that
$\opsenv{e}{l}{E}$, and
and from the form of the type rule [T-IsNull] we know
$\typeenv{e}{T@(\oton{r})}$.
Clearly $\valueconsist{\textit{false}}{bool}$, and
$\traceconsist{E}$ and 
      $\privconsist{E}{\Phi}$ follow by induction.

\item Case [E-IsNull-T] is almost identical to case [E-IsNull-F], with $l$ replaced by \textit{null} and
\textit{false} replaced by \textit{true}.

\item Case [E-Let].
By assumption, 
$$\opsenv{\text{let } id:T_1 = e_1 \text{ in } e_2}{v_2}{E'}$$ and 
$$\typeenv{\text{let } id:T_1 = e_1 \text{ in } e_2}{T_2}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-Let], we know that $$\opsenv{e_1}{v_1}{E_1}$$
From type rule [T-Let], we know that $$\typeenv{e_1}{T_1}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Let], we have $$\opsenv[L=L',S=S']{e_2}{v_2}{E_2}$$
where $L' = L[v_1/id]$ and $S' = apply(S,E_1)$.  
From [T-Let], we have $\typeenv[G={\Gamma[T_1/id]}]{e_2}{T_2}$.
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From $\localconsist{L}{\Gamma}$ and $\valueconsist{v_1}{T_1}$ we conclude $\localconsist{L[v_1/id]}{\Gamma[T_1/id]}$.
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

Finally, from [E-Let] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$
Then, by Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by 
Lemma~\ref{lemma:effsound:effects2} $\privconsist{E'}{\Phi}$.

\item Case [E-Add].
By assumption, 
$$\opsenv{e_1 + e_2}{v'}{E'}$$ and 
$$\typeenv{e_1 + e_2}{T}$$ and 
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and $\storeconsist{S}{H}$.
The only type rule that can apply is [T-Add], so $T = int$.

For the subexpression $e_1$, 
from [E-Add] we know $$\opsenv{e_1}{v_1}{E_1}$$
From [T-Add], we know that $$\typeenv{e_1}{int}$$
Then, by induction, we have $\valueconsist{v_1}{T_1}$, $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Add], we know $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From [T-Add] we know $$\typeenv{e_2}{int}$$
Then, by induction, we have $\valueconsist{v_2}{T_2}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

It is immediate that $\valueconsist{v_1 + v_2}{int}$.
Finally, from [E-Add] we also have $$valid\_interleave(S,C,E',E_1,E_2)$$ and
Lemma~\ref{lemma:heapconst:effects2}, $\traceconsist{E'}$ and by Lemma~\ref{lemma:effsound:effects2},
$\privconsist{E'}{\Phi}$.

\item Case [E-Compare] is isomorphic to [E-Add].

\item Case [E-IfElse-T].
 By assumption, 
$$\opsenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{v_2}{E}$$ and 
$$\typeenv{\text{ if } e_1 \text{ then } e_2 \text{ else } e_3}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For the subexpression $e_1$, 
from [E-IfElse-T] we know $$\opsenv{e_1}{true}{E_1 \dplus E_2}$$
From [T-IfElse], we know that $$\typeenv{e_1}{bool}$$
Clearly $\valueconsist{true}{bool}$, and by induction we have  $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-IfElse-T] we know $$\opsenv[S=S']{e_2}{v_2}{E_2}$$
where $S' = apply(S,E_1)$.  By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S' = \text{apply}(S,E)}{H}$.
From [T-IfElse-T] we know $$\typeenv{e_2}{T}$$
Then, by induction, we have $\valueconsist{v_2}{T}$, $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

Finally, by Lemma~\ref{lemma:heapconst:effects1}, $\traceconsist{E_1 \dplus E_2}$ and by Lemma~\ref{lemma:effsound:effects1},
$\privconsist{E_1 \dplus E_2}{\Phi}$.

\item Case [E-IfElse-F] is nearly the same as [E-IfElse-T] with {\em false} replacing {\em true} and $e_3$ replacing $e_2$.

\item Case [E-Newcolor].
By assumption, 
$$\opsenv{\text{newcolor } r}{K}{[]}$$ 
where  $K = \{ (l_1,v_1),\ldots,(l_p,i_p) \}$ such that
$l_i \in M(r)$ for all $i$ and $l_1,\ldots,l_p$ are pairwise distinct locations.
In addition, we have the assumptions
$$\typeenv{\text{newcolor } r}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Newcolor], so can infer that $T = \text{coloring}(r)$.
From the structure of $K$ it follows that $\valueconsist{K}{\text{coloring(r)}}$.
The other two conclusions,
$\traceconsist{[]}$ and 
$\privconsist{[]}{\Phi}$, follow trivially.

\item  Case [E-Color].
By assumption, 
$$\opsenv{\text{color($e_1$, $e_2$, $e_3$)}}{K'}{E'}$$ and
$$\typeenv{\text{color($e_1$, $e_2$,$e_3$)}}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The only type rule that can apply if [T-Color], so we know $T = \text{coloring}(r)$.

For the subexpression $e_1$, 
from [E-Color] we know $$\opsenv{e_1}{K}{E_1}$$
From [T-Color], we know that $$\typeenv{e_1}{\text{coloring}(r)}$$
By induction it follows that $\valueconsist{K}{\text{coloring(r)}}$ and $\traceconsist{E_1}$ and 
      $\privconsist{E_1}{\Phi}$.

For the subexpression $e_2$, 
from [E-Color] we know $$\opsenv[S=S']{e_2}{l}{E_2}$$ where $S' = \text{apply}(S,E_1)$.
From [T-Color], we know that $$\typeenv{e_2}{T'@r}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S'}{H}$.
By induction, $\valueconsist{l}{T'@r}$ and $\traceconsist{E_2}$ and 
      $\privconsist{E_2}{\Phi}$.

For the subexpression $e_3$, 
from [E-Color] we know $$\opsenv[S=S'']{e_3}{v}{E_3}$$ where $S'' \text{apply}(S',E_2)$.
From [T-Color], we know that $$\typeenv{e_3}{int}$$
By Lemma~\ref{lemma:heapconst:apply}, $\storeconsist{S''}{H}$.
By induction, $\valueconsist{v}{int}$ and $\traceconsist{E_3}$ and 
      $\privconsist{E_3}{\Phi}$.

Now $K'$ is built from $K$ by extending $K$ with a mapping for location $l$.  First, because $\valueconsist{l}{T'@r}$ and
$\valueconsist{K}{\text{coloring(r))}$ and because $K'$ excludes any previous mapping for $l$ (guaranteeing the the coloring is a function), we conclude $\valueconsist{K'}{\text{coloring}(r)}}$.  From [E-Color], we also know
$$valid\_interleave(S, C, E', E_1, E_2, E_3)$$ and by applying Lemma~\ref{lemma:heapconst:effects2} twice
we can conclude $\traceconsist{E'}$.  Similarly, by applying Lemma~\ref{lemma:effsound:effects2} twice
we can conlude $\privconsist{E'}{\Phi}$.

\item  Case [E-UpRgn].
By assumption, $$\opsenv{\text{upregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{upgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-UpRgn], so we also know $T = T' @ (\oton{r})$.

From [E-UpRgn] we know $$\opsenv{e}{v}{E}$$ and
from [T-UpRgn] we know $$\typeenv{e}{T' @ (r_1',\ldots,r_k')}$$
By induction, we can conclude that $\valueconsist{v}{T'@(r_1',\ldots,r_k')}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T'@(\oton{r})}$, we first observe that from
$\valueconsist{v}{T'@(r_1',\ldots,r_k')}$ we know
there is an $i$ such that $v \in M(r_i')$ and 
$H(v) = \hat T'$.   From [T-UpRgn]
we know $\forall i. \exists j. r_i' \leq r_j \in \Omega^*$.  Combining this with
$\mapconsist{\Omega}$ and Lemma~\ref{lemma:omegaclosure}, we know there
is some $r_j$ such that $v \in M(r_j)$.
Therefore $\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-DownRgn].
By assumption, $$\opsenv{\text{downregion}(e,\oton{r})}{v}{E}$$ and
$$\typeenv{\text{downgregion}(e,\oton{r})}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.
The type rule must be [T-DownRgn], so we also know $T = T' @ (\oton{r})$.

From [E-DownRgn] we know $$\opsenv{e}{l}{E}$$ and
from [T-DownRgn] we know $$\typeenv{e}{T' @ (r_1',\ldots,r_k')}$$
By induction, we can conclude that $\valueconsist{l}{T'@(r_1',\ldots,r_k')}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T'@(\oton{r})}$ there are two cases.
If $v = null$ then $\valueconsist{null}{T'@(\oton{r})}$ follows by definition.
Otherwise, $v = l$ and $l \in M(r_i)$ for some $i$.  We also know $H(v) = \hat T'$
because $\valueconsist{l}{T'@(r_1',\ldots,r_k')}$.  Therefore
$\valueconsist{v}{T'@(\oton{r})}$.

\item Case [E-Read].
By assumption, $\opsenv{\text{read}(e)}{v}{E \dplus [ read(l,excl,v) ]}$ and
$\typeenv{\text{read}(e)}{T}$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

From [E-Read] we have $$\opsenv{e}{l}{E}$$ and
from [T-Read] we know $$\typeenv{e}{T @ (\oton{r})}$$
By induction, we can conclude that $\valueconsist{l}{T@(\oton{r})}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

To show $\valueconsist{v}{T}$ there are two cases.
If $v = M(l)$ then the claim follows from $\valueconsist{l}{T@(\oton{r})}$.
If $v$ is instead some $v'$ of type $H(l)$ the result also follows because $H(l) = \hat{T}$.
Next, $\traceconsist{E \dplus [ read(l,excl,v)]}$ follows immediately from $\traceconsist{E}$ as
any read operation is consistent (the resriction is only on writes and reductions).

Finally, we must show $\privconsist{E \dplus [ read(l,excl,v)]}{\Phi}$.
Since we have already shown $\privconsist{E}{\Phi}$, it suffices to show $\privconsist{[ read(l,excl,v) ]}{\Phi}$.
By $\valueconsist{l}{T@(\oton{r})}$, there is some $r_i$ such that $l \in M(r_i)$.  Furthermore,
from [T-Read] we have $reads(r_i) \in \Phi^*$, which shows $\privconsist{[ read(l,excl,v)]}{\Phi^*}$.
The result follows from Lemma~\ref{lemma:phiclosure}.

\item Case [E-Write].
By assumption, $$\opsenv{\text{write}(e_1,e_2)}{l}{E' \dplus [ write(l,excl,v) ]}$$ and
$$\typeenv{\text{write}(e_1,e_2)}{T@(\oton{r})}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For subexpression $e_1$,
From [E-Write] we have $$\opsenv{e_1}{l}{E_1}$$ and
from [T-Write] we know $$\typeenv{e_1}{T @ (\oton{r})}$$.
By induction, we can conclude that $\valueconsist{l}{T@(\oton{r})}$,
$\traceconsist{E_1}$ and $\privconsist{E_1}{\Phi}$.

For subexpression $e_2$,
From [E-Write] we have $$\opsenv[S=S']{e_2}{v}{E_2}$$ where
$S' = \text{apply}(S,E_1)$.
From [T-Write] we know $\typeenv{e_2}{T}$.
Using Lemma~\ref{lemma:heapconst:apply} it follows that $\storeconsist{S'}{H}$. 
Then by induction, we can conclude that $\valueconsist{v}{T}$,
$\traceconsist{E_2}$ and $\privconsist{E_2}{\Phi}$.

We have already shown $\valueconsist{l}{T@(\oton{r})}$, so it only remains to show
$\traceconsist{E' \dplus [ write(l,excl,v) ]}$ and 
$\privconsist{E' \dplus [ write(l,excl,v) ]}{\Phi}$.
From [E-Write] we have $$valid\_interleave(S,C,E',E_1,E_2)$$ and therefore by Lemma~\ref{lemma:heapconst:effects2} we have
$\traceconsist{E'}$.  To show $\traceconsist{[ write(l,excl,v)]}$ we must show $\valueconsist{v}{H(l)}$,
which follows from $\valueconsist{v}{T}$ and $\valueconsist{l}{T@(\oton{r})}$ and $\storeconsist{S}{H}$.
Thus, $\traceconsist{E' \dplus [ write(l,excl,v) ]}$.

To show $\privconsist{E' \dplus [ write(l,excl,v) ]}{\Phi}$, we first observe that
$valid\_interleave(S,C,E',E_1,E_2)$ and Lemma~\ref{lemma:effsound:effects2} imply
$\privconsist{E'}{\Phi}$.  To show $\privconsist{[ write(l,excl,v) ]}{\Phi}$, we observe that because
$\valueconsist{l}{T@(\oton{r})}$, there is some $r_i$ such that $l \in M(r_i)$.  Furthermore,
from [T-Write] we have $writes(r_i) \in \Phi^*$, which shows $\privconsist{[ write(l,excl,v)]}{\Phi^*}$.
The result follows from Lemma~\ref{lemma:phiclosure}.

\item Case[E-Reduce].
By assumption, $$\opsenv{\text{reduce}(id,e_1,e_2)}{l}{E' \dplus [ reduce_{id}(l,excl,v) ]}$$ and
$$\typeenv{\text{reduce}(id,e_1,e_2)}{T_1@(\oton{r})}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For subexpression $e_1$,
From [E-Reduce] we have $$\opsenv{e_1}{l}{E_1}$$ and
from [T-Reduce] we know $$\typeenv{e_1}{T @ (\oton{r})}$$.
By induction, we can conclude that $\valueconsist{l}{T@(\oton{r})}$,
$\traceconsist{E_1}$ and $\privconsist{E_1}{\Phi}$.

For subexpression $e_2$,
From [E-Reduce] we have $$\opsenv[S=S']{e_2}{v}{E_2}$$ where
$S' = \text{apply}(S,E_1)$.
From [T-Reduce] we know $$\typeenv{e_2}{T_2}$$
Using Lemma~\ref{lemma:heapconst:apply} it follows that $\storeconsist{S'}{H}$. 
Then by induction, we can conclude that $\valueconsist{v}{T_2}$,
$\traceconsist{E_2}$ and $\privconsist{E_2}{\Phi}$.

We have already shown $\valueconsist{l}{T_1@(\oton{r})}$, so it only remains to show
$\traceconsist{E' \dplus [ reduce_{id}(l,excl,v) ]}$ and 
$\privconsist{E' \dplus [ reduce_{id}(l,excl,v) ]}{\Phi}$.
From [E-Reduce] we have $valid\_interleave(S,C,E',E_1,E_2)$ and therefore by Lemma~\ref{lemma:heapconst:effects2} we have
$\traceconsist{E'}$.  To show $\traceconsist{[ reduce_{id}(l,excl,v)]}$ we must show three things.
First, we must show $M \llbracket \Gamma \rrbracket (id) = (\hat T_1, \hat T_2), \emptyset, \emptyset \rightarrow \hat T_1$ which
follows from the assumption $\Gamma(id) = (T_1, T_2), \emptyset, \emptyset \rightarrow T_1$ in [T-Reduce].
Second, we must show $H(l) = \hat T_1$, which follows from $\valueconsist{l}{T_1@(\oton{r})}$.
Finally, we have already shown $\valueconsist{v}{\hat T_2}$.
Thus, $$\traceconsist{E' \dplus [ reduce_{id}(l,excl,v) ]}$$.

To show $$\privconsist{E' \dplus [ reduce_{id}(l,excl,v) ]}{\Phi}$$, we first observe that
$valid\_interleave(S,C,E',E_1,E_2)$ and Lemma~\ref{lemma:effsound:effects2} imply
$\privconsist{E'}{\Phi}$.  To show $\privconsist{[ reduce_{id}(l,excl,v) ]}{\Phi}$, we observe that because
$\valueconsist{l}{T@(\oton{r})}$, there is some $r_i$ such that $l \in M(r_i)$.  Furthermore,
from [T-Reduce] we have $reduce_{id}(r_i) \in \Phi^*$, which shows $\privconsist{[ reduce_{id}(l,excl,v)]}{\Phi^*}$.
The result follows from Lemma~\ref{lemma:phiclosure}.

\item Case [E-Pack].
By assumption, $$\opsenv{\text{pack } e_1 \text{ as } T_1[\otok{k}{r}]}{v'}{E}$$ and
$$\typeenv{\text{pack } e_1 \text{ as } T_1[\otok{k}{r}]}{T_1}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

From [E-Pack] we have $\opsenv{e_1}{v}{E}$ and
from [T-Pack] we know $\typeenv{e_1}{T_2[r_1/r'_1,\ldots,r_k/r'_k]}$.
By induction, we can conclude that $\valueconsist{v}{T_2[r_1/r'_1,\ldots,r_k/r'_k]}$,
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

The only thing remaining to show is $\valueconsist{v'}{T_1}$.
From [E-Pack] we have   $v' = \langle \langle \otok{k}{\rho}, v \rangle \rangle$.
From [T-Pack] we have $T_1 = \exists r'_1, \ldots r'_n.\ T_2\text{ where }\Omega_1$
and $\Omega_1[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Omega^*$.  
%
% FIXME  The next sentence can't be made fully formal with the current setup.
%
The result
follows from the definition of value consistency on region relationships.

\item Case [E-Unpack].
By assumption, $$\opsenv{\text{unpack } e_1 \text{ as } id : T_1[\otok{k}{r}] \text{ in } e_2}{v_2}{E'}$$ and
$$\typeenv{\text{unpack } e_1 \text{ as } id : T_1[\otok{k}{r}] \text{ in } e_2}{T}$$ and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For subexpression $e_1$, from [E-UnPack] we have $$\opsenv{e_1}{\langle \langle \otok{k}{\rho},v \rangle\rangle}{E}$$ and
from [T-UnPack] we know $$\typeenv{e_1}{T_1 = \exists r'_1,\ldots,r'_n.T_2 \text{ where }\Omega_1}$$
By induction, we can conclude that $\valueconsist{\langle \langle \otok{k}{\rho},v \rangle\rangle}{T_1}$ and
$\traceconsist{E}$ and $\privconsist{E}{\Phi}$.

For subexpression $e_2$, from [E-UnPack] we have
$$\opsenv[M=M',L=L',S=S']{e_2}{v_2}{E_2}$$ where
$S' = \text{apply}(S,E_2)$, $L' = L[v_1/id]$ and $M' = M[\rho_1/r_1,\ldots,\rho_k/r_k]$.
From [T-Unpack] we have
$$\typeenv[G=\Gamma',O=\Omega']{e_2}{T_3}$$
where
$\Gamma' = \Gamma[T_2[r_1/r'_1,\ldots,r_k/r'_k] / id]$ and
$\Omega' = \Omega \cup \Omega_1[r_1/r'_1,\ldots,r_k/r'_k]$.
To establish the induction hypothesis for $e_2$, we must show
$\mapconsist[M']{\Omega'}$ and
$\localconsist[M']{L'}{\Gamma'}$ and
$\storeconsist{S'}{H}$.

To show $\mapconsist[M']{\Omega'}$, consider a constraint $r \leq r' \in \Omega'$.  There are two cases. If $r \leq r' \in \Omega_1[r_1/r'_1,\ldots,r_k/r'_k]$
then $M'(r) \subseteq M'(r')$ follows from $\valueconsist{\langle \langle \otok{k}{\rho},v \rangle\rangle}{T_1}$.
If $r \leq r' \not\in \Omega_1[r_1/r'_1,\ldots,r_k/r'_k]$ then $r \leq r' \in \Omega$ and $M'(r) = M(r) \subseteq M(r') = M'(r')$.
The reasoning for constraints $r * r'$ is the same.

To show $\localconsist[M']{L'}{\Gamma'}$, we must show
$\valueconsist[M']{L'(x)}{\Gamma'(x)}$ for all identifiers $x$ in
the domain of $\Gamma'$.  For any $id' \neq id$, the statement is equivalent
to showing $\valueconsist[M']{L(x)}{\Gamma(x)}$, which must hold becuase $M'$ does 
does not change any existing binding in $M$.  For $id$, we have
$\valueconsist[M']{v_1}{T_1[r_1/r'_1,\ldots,r_k/r'_k]}$, which is implied by
$\valueconsist{\langle \langle \otok{k}{\rho},v \rangle\rangle}{T_1}$.

Finally, as in previous cases, $\storeconsist{S'}{H}$ follows from Lemma~\ref{lemma:heapconst:apply}.
Thus, we can conclude that 
$\valueconsist[M']{v_2}{T_3}$ and
$\traceconsist{E_2}$ and 
$\privconsist{E_2}{\Phi}$.

Because $T_3$ cannot include any of the unpacked regions in it, $\valueconsist{v_2}{T_3}$.
From [E-Unpack] we have $$valid\_interleave(S, C, E', E_1, E_2)$$ and using Lemma~\ref{lemma:heapconst:effects2} we
prove $\traceconsist{E'}$.
To finish the case, we note that from $\privconsist{E_1}{\Phi}$, $\privconsist{E_2}{\Phi}$, 
$valid\_interleave(S, C, E', E_1, E_2)$, and using Lemma~\ref{lemma:effsound:effects2} we can conclude $\privconsist{E'}{\Phi}$.

\item Case [E-Part].
By assumption,
$$\opsenv{\text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2}{v}{E'}$$
and
$$\typeenv{\text{partition}\ r_p\text{ using }e_1\text{ as }\otok{k}{r}\text{ in }e_2}{T}$$
and
$\mapconsist{\Omega}$ and 
$\localconsist{L}{\Gamma}$ and 
$\storeconsist{S}{H}$.

For subexpression $e_1$, from [E-Part] we have $$\opsenv{e_1}{K}{E_1}$$ and
from [T-Part] we have $$\typeenv{e_1}{\text{coloring}(r_p)}$$
By induction, we conclude that $\valueconsist{K}{\text{coloring}(r_p)}$ and
$\traceconsist{E_1}$ and 
$\privconsist{E_1}{\Phi}$.
For subexpression $e_2$, from [E-Part] we have
$$\opsenv[M=M',S=S']{e_2}{v}{E_2}$$
where 
$M' = M[\rho_1/r_1, \ldots, \rho_k/r_k]$ and
$S' = \text{apply}(S, E_1)$.
From [T-Part] we have
$$\typeenv[O=\Omega']{e_2}{T}$$ 
where
$\Omega' = \Omega \wedge \bigwedge_{i \in [1,k]} r_i \leq r_p \wedge \bigwedge_{1 \leq i < j \leq k} r_i * r_j$.    
To establish the induction hypothesis for $e_2$, we must show
$\mapconsist[M']{\Omega'}$ and
$\localconsist[M']{L}{\Gamma}$ and
$\storeconsist{S'}{H}$.

For $\mapconsist[M']{\Omega'}$, we observe that all of the constraints are either present previously
(hold in $\mapconsist{\Omega}$) or are established explicitly by the partition operation.  Note that
the type of the coloring carries the name of the region being partitioned and that all introduced region names are
not in use in $M$.  For $\localconsist[M']{L}{\Gamma}$, we simply observe that $\localconsist{L}{\Gamma}$ holds
and $M'$ is an extension of $M$.  
Finally, as in previous cases, $\storeconsist{S'}{H}$ follows from Lemma~\ref{lemma:heapconst:apply}.

Therefore, we can conclude by induction
$\valueconsist[M']{v}{T}$ and
$\traceconsist{E_2}$ and 
$\privconsist{E_2}{\Phi}$.
Because $T$ may not mention any of the regions introduced in the partitioning, we have
$\valueconsist{v}{T}$.
From [E-Part] we have $$valid\_interleave(S, C, E', E_1, E_2)$$ and using Lemma~\ref{lemma:heapconst:effects2} we
prove $\traceconsist{E'}$.
To finish the case, we note that from $\privconsist{E_1}{\Phi}$, $\privconsist{E_2}{\Phi}$, 
$valid\_interleave(S, C, E', E_1, E_2)$, and using Lemma~\ref{lemma:effsound:effects2} we can conclude $\privconsist{E'}{\Phi}$.

\item Case [E-Call].
From [E-Call] we have
$$\opsenv{id[\otok{k}{r}](\oton{e})}{v_{n+1}}{E''}$$
and from [T-Call] we have
$$\typeenv{id[\otok{k}{r}](\oton{e})}{T_r[r_1/r'_1,\ldots,r_k/r'_k]}$$

For subexpression $e_1$, from [E-Call] we have $\opsenv{e_1}{v_1}{E_1}$ and from [T-Call] we have
$\typeenv{e_1}{T_1[r_1/r'_1,\ldots,r_k/r'_k]}$.
By induction, 
$\valueconsist{v_1}{T_1[r_1/r'_1,\ldots,r_k/r'_k]}$,
$\traceconsist{E_1}$ and 
$\privconsist{E_1}{\Phi}$.

For the other subexpressions that are arguments to the call the reasoning is very similar; we show
only the last argument $e_n$.  First, 
$$S_n = \text{apply}(E_{n-1},\text{apply}(E_{n-2},\ldots \text{apply}(E_1,S) \ldots )$$
From Lemma~\ref{lemma:heapconst:apply} it follows that $\storeconsist{S_n}{H}$.
We also have $\opsenv{e_n}{v_n}{E_n}$ from [E-Call] and
$\typeenv{e_n}{T_n[r_1/r'_1,\ldots,r_k/r'_k]}$ from [T-Call].
By induction, 
$\valueconsist{v_n}{T_n[r_1/r'_1,\ldots,r_k/r'_k]}$,
$\traceconsist{E_n}$ and 
$\privconsist{E_n}{\Phi}$.

Now, from [E-Call] we have
$valid\_interleave(S, C, E', \oton{E})$
and furthermore
$M' = \{ (r'_1, M(r_1)), \ldots (r'_k, M(r_k)) \}$  and
$L' = \{ (a_1, v_1), \ldots, (a_n, v_n) \}$ and
$S' = \text{apply}(S, E')$ and
$\opsenv[M=M',L=L',S=S',C=C']{e_{n+1}}{v_{n+1}}{E_{n+1}}$.
From [T-Program] we know
$\typeenv[G=\Gamma',O=\emptyset,P=\Phi']{e_{n+1}}{T_r}$,
where $e_{n+1}$ is the body of the function.
To apply the induction hypothesis to $e_{n+1}$ we must
show 
\begin{itemize}
\item $\mapconsist[M']{\emptyset}$,
which is immediate.

\item $\localconsist[M']{L'}{\Gamma'}$ where
$\Gamma'(a_i) = T_i$and $\storeconsist{S'}{H}$.
\end{itemize}

$\Gamma(id) = \forall r'_1, \ldots r'_k.(\oton{T}),\Phi', Q' \rightarrow T_r$ 
$\text{function }id[\otok{k}{r'}](a_1 : T_1, \ldots, a_n : T_n), \Phi', Q' : T_r = e_{n+1}$

$E'_{n+1} = mark\_coherence(E_{n+1}, M' \llbracket Q' \rrbracket)$ 
$valid\_interleave(S, C, E'', E', E_{n+1})$

$\typeenv{e_i}{T_i[r_1/r'_1,\ldots,r_k/r'_k]}$
$\Phi'[r_1/r'_1,\ldots,r_k/r'_k] \subseteq \Phi^*$

\end{itemize}
\end{proof}

\section{Coherence}
\label{sec:coherence}

Theorem~\ref{thm:effects} guarantees that a function call produces a value of the
correct type and does so within the regions of memory for which privileges are held.
It does not, however, address {\em coherence}, what concurrent operations on the regions are permitted.
An innovation in Legion is that programmers can control coherence on a per-region basis.  Intuitively,
the different coherence modes permit different interleavings of memory traces.  For example,
in addition to being an interleaving of the constituent traces, exclusive access requires interleaved memory traces to be:
\begin{itemize}
\item {\em coherent} - a read of location $l$ should see the effect of all writes and reductions to $l$
occurring earlier in the interleaved trace, and
\item {\em sequentially equivalent} - the value read by each read 
operation and the result of applying the interleaved trace to the original store should be the same as if each constituent trace
was applied to the store in a valid sequential execution of the program.
\end{itemize}
These requirements are relaxed for either a non-empty clobber set (representing
operations that may occur concurrently as part of evaluating other functions) or for coherence
modes other than {\em excl}.

\begin{figure*}
$$
\begin{array}{lll}
valid\_interleave(S, C, E', \oton{E}) & = & any\_interleave(E', \oton{E}) \wedge \\
& & coherent(S, L_{excl}(E', C), L_{excl}(E', C), E')\ \wedge \\
& & seq\_equiv(S, L_{excl}(E', C), L_{excl}(E', C), E', \oton{E})\ \wedge \\
& & coherent(S, L_{atomic}(E', C), \emptyset, E')\ \wedge \\
& & \exists \oton{\pi}, (\oton{\pi}) \text{ is a permutation of } (1,\ldots,n)\ \wedge \\
& & \ \ \ seq\_equiv(S, L_{atomic}(E', C), \emptyset, E', E_{\pi_1}, \ldots, E_{\pi_n}) \\
coherent(S, L_1, L_2, []) & = & \text{true} \\
coherent(S, L_1, L_2, [ \epsilon ] \dplus E) & = &\
\begin{cases}
(l \in L_2 \Rightarrow S(l) = v)\ \wedge \\
\ \ \ coherent(S, L_1, L_2, E), & \text{if $\epsilon = read(l, c, v)$} \\
coherent(apply(S, \epsilon), L_1, L_2 \cup \{ l \}, E), & \text{if $\epsilon = write(l, c, v)$ and $l \in L_1$} \\
coherent(apply(S, \epsilon), L_1, L_2, E), & \text{otherwise}
\end{cases} \\
seq\_equiv(S, L_1, L_2, E', \oton{E}) & = & coherent(S, L_1, L_2, E_1 \dplus \ldots \dplus E_2)\ \wedge \\
& & \forall l \in L_1. apply(S, E')(l) = apply(S, E_1 \dplus \ldots \dplus E_n)(l) \\
L_{excl}(E, C) & = & \{ l : op(l, excl, v) \text{ in } E  \} \setminus C \\
L_{atomic}(E, C) & = & \{ l : op(l, atomic, v) \text{ in } E \} \setminus (C \cup L_{excl}(E, C)) \\
\end{array}
$$
\caption{Valid Interleaving Test}
\label{fig:validinterleave}
\vspace{-5mm}
\end{figure*}

Figure~\ref{fig:validinterleave} formally defines {\em
valid\_interleave}, which we briefly explain.  Consider
traces $\oton{E}$.  The {\em exclusive location set} $L_{excl}$ is the
set of locations used in an operation with coherence mode $excl$ in
some trace $E_i$, excepting any locations in the clobber set $C$.
Similarly, the {\em atomic location set} $L_{atomic}$ is the set of
locations in an operation with coherence mode $atomic$, but no
operations with coherence mode $excl$, again excepting any location in
the clobber set.  The $coherent(S,L_1,L_2,E)$ test checks that all
reads of locations in $L_2$, and any read of a location $l \in L_1$
that comes after a write to $l$, return the value most recently written
to the store.  (If $L_2 = \emptyset$ initially, atomic reads without a preceding write are not 
checked, but are deferred to a point where traces that might be ordered before $E$ are included.)  The $seq\_equiv$
test verifies every read in $E'$ with a location in $L_1$ 
is coherent if operations are performed in 
program order and that the store that results from applying $E'$
matches the store resulting from applying the
constituent traces in program order for all locations in $L_1$.  Program order
is implicitly defined in the operational semantics, and is equivalent to a 
postorder traversal of an expression's tree.  For
the atomic location set, we allow the constituent traces to be
permuted, but all locations must be checked with the same permutation.

Flexibility in interleaving operations is the key to achieving parallel execution, but
implementing the precise criteria for checking a valid interleaving is
much too expensive.  We now consider practical mechanisms that generate valid interleavings and 
under what conditions these approaches may be used.

The first option is to serialize the evaluation of the subexpressions that
make up an expression and attempt to generate a valid interleaving by simply
concatenating the constituent traces in program order.  However, there is a
complication related to the non-determinism that is allowed for reads.  Although
each subexpression's trace will have been checked for coherence, the checking
of some operations may have been deferred due to non-empty
clobber sets.  It is possible that when an interleaving of these traces
is attempted in the parent expression's context, there may no valid interleaving
if the wrong guesses were made in the derivation of a subexpression.
The solution in the case of serial execution is to not expand the clobber
set for subexpressions, instead using the parent expression's clobber set.

Only a single rule is modified for this alternate execution - $e_{n+1}$ in
{\sc E-Call} uses the original $C$ instead of $C'$.  This change
can result in there being no terminating execution for an expression in
which the only terminating executions required a particular (non-program-)
ordering of subexpression evaluation.  However, the Legion core language never
guarantees parallel execution, so the evaluation of such an 
expression is unlikely to terminate even under the original clobber set.
We can, however, show that any terminating execution that results from using
the restricted clobber set is a valid execution under the original:

\begin{lem}
\label{lem:clobberreduce}
\rm 
Let $\typeenv{e}{T}$ and $\opsenv{e}{v}{E}$ and $\storeconsist{S}{H}$.  If $C \subseteq C'$, then
$\opsenv[C=C']{e}{v}{E}$.
\end{lem}

The proof of this (and other lemmas in this section) is included in \cite{LegionTypes12}.
We now show that the concatenation of the memory traces from these serially
executed subexpressions is always a valid interleaving:

\begin{thm}
\label{thm:sequential}
\rm
Let $\oton{e}$ be expressions such that 
%$\opsenv[S={S_{i-1}}]{e_i}{v_i}{E_i}$, 
$M, L, H, S_{i-1},\linebreak C \vdash e_i \mapsto v_i, E_i$,
where $S_i = apply(S_{i-1}, E_i)$.  If $E' = E_1 \dplus \ldots \dplus E_n$, then
$valid\_interleave(S_0, C, E', \oton{E})$.
\end{thm}

\begin{proof}
First, we must show that $any\_interleave(E', \oton{E})$.  This is done by
induction on the length of $E_1$, followed by $E_2$, and so on.  Next, we must
show that $E'$ is coherent for both $L_{excl}$ and $L_{atomic}$.  This is done by
matching the structure of $E'$ with $\oton{E}$ and the use of the following
lemma: $apply(S, E_1 \dplus E_2) = apply(apply(S, E_1), E_2)$.  With the
question of coherence addressed, the sequential equivalence of $E'$ is trivial
for both the exclusive and atomic cases.
\end{proof}

Next we examine how to modify the serialized interleaving to produce valid interleavings allowing parallel execution. 
We consider when we may swap the order of two adjacent memory operations in a valid interleaving.
\begin{itemize}
\item The two operations must have come from different constituent
traces, or the swap would violate $any\_interleave$.
\item If one is a read of
a location $l$ (that is not in the clobber set), the other may not be a write or reduce to $l$, 
as changing the order would almost certainly change the value read.
\item If one is a write of a location $l$, the other may not be a write or reduce to $l$,
as the store that would result from applying those operations in the opposite order may
differ in $l$.
\item Finally, if both operations are reductions to the same location, they may be swapped if they
use the same reduction function, as each reduction function is assumed to commute with itself.  
%(We 
%follow the example of other work (i.e. DPJ) and leave proofs of commutativity to the programmer,
%possibly with the aid of other tools.)  
However, if the reduction functions have different names, swapping the two operations might result in a different store.
\end{itemize}

To summarize this, we define a {\em non-interference} operator on two memory operations that have been
annotated with which constituent trace they came from as follows:
$$
\begin{array}{l@{ }l}
op_1^{E_1}(l_1, c_1,& v_1)\ \nonint\ op_2^{E_2}(l_2, c_2, v_2) \Leftrightarrow E_1 \not= E_2\ \wedge \\
& \big( l_1 \not= l_2 \vee (op_1 = read \wedge op_2 = read)\ \vee \\
& (op_1 = reduce_{id_1} \wedge op_2 = reduce_{id_2} \wedge id_1 = id_2) \big)
\end{array}
$$

\begin{lem}
\label{lem:nonintswap}
\rm
Let $S$ be a store, $C$ a clobber set, $\oton{E}, E'_a, E'_b$ memory traces, and
$\epsilon_1, \epsilon_2$ two annotated memory operations such that $\epsilon_1 \nonint \epsilon_2$.
Then,
$$valid\_interleave(S, C, E'_a \dplus [ \epsilon_1, \epsilon_2 ] \dplus E'_b, \oton{E})$$
if and only if
$$valid\_interleave(S, C, E'_a \dplus [ \epsilon_2, \epsilon_1 ] \dplus E'_b, \oton{E}).$$
\end{lem}
\begin{proof}
As before, we consider the three conditions needed to be a valid interleaving.  We describe the
implication in the ``forward'' direction, but the arguments are all symmetric.
\begin{enumerate}
\item If $E'_a \dplus [ \epsilon_1 ] \dplus [ \epsilon_2 ] \dplus E'_b$ is an interleaving of $\oton{E}$,
the proof must first ``pop off'' all of $E'_a$, leaving $[ \epsilon_1 ] \dplus [ \epsilon_2 ] \dplus E'_b$ for
the interleaved subtrace.  The next two steps must pop $\epsilon_1$ and $\epsilon_2$, leaving $E'_b$.
They cannot be in the same subtrace because $\epsilon_1 \nonint \epsilon_2$, so we could also pop
$\epsilon_2$ and then $\epsilon_1$, resulting in a proof that $E'_a \dplus [ \epsilon_2 ] \dplus [ \epsilon_1 ] \dplus E'_b$ is also an interleaving of the constituent traces.
\item For coherence, we must show the value read (if any) by $\epsilon_1$ does not change, the value
read (if any) by $\epsilon_2$ does not change, and that any read in $E'_b$ sees no change.  With the
help of Lemma~\ref{lemma:applychain}, we can eliminate the common $E'_a$ from all sequences by
using $S' = apply(S, E'_a)$.  Similarly to show that $apply(S', [ \epsilon_1  ] \dplus [ \epsilon_2 ] \dplus E) = apply(S', [ \epsilon_2 ] \dplus [ \epsilon_1 ] \dplus E)$ for any $E$, we need only to show that it holds
for $E = []$.  The proof is now reduced to showing the following:
\begin{enumerate}
\item $\epsilon_1 = read(l, c, v) \Rightarrow S'(l) = apply(S',\epsilon_2)(l)$ - since $\epsilon_1$
is a read, either $\epsilon_2$ must be a read (in which case $apply(S, \epsilon_2) = S'$) or must be
to a different location, so the result of applying just it to $S'$ cannot change the value in 
location $l$.
\item $\epsilon_2 = read(l, c, v) \Rightarrow apply(S', \epsilon_1)(l) = S'(l)$ - this is the same
as above, with $\epsilon_1$ and $\epsilon_2$ switched.
\item $apply(S', \epsilon_1 \dplus \epsilon_2) = apply(S', \epsilon_2 \dplus \epsilon_1)$ - we must
show that the two store operations $apply(\bullet, \epsilon_1)$ and $apply(\bullet, \epsilon_2)$
commute.  Modifications to two different locations commute, as do two reads (which make no change
to the store).  Finally, two reductions using the same function to the same location also commute.
\end{enumerate}
\item The proof that sequential equivalence is maintained comes down to showing the final store
is identical after the swap of the two operations, and this is merely a special case of the third
piece of the coherence proof.
\end{enumerate}
\end{proof}

We now extend the non-interference operator to entire memory traces, providing a way to say that
no operation in a trace interferes with any operation in another:
$$E_1 \nonint E_2 \Leftrightarrow \displaystyle\bigwedge_{\epsilon_1 \text{ in } E_1, \epsilon_2 \text{ in } E_2} \epsilon_1 \nonint \epsilon_2$$
and observe that this property permits arbitrary ordering (i.e. parallel execution) of the 
corresponding subexpressions.

\begin{lem}
\label{lem:noninteffects}
\rm
Let $S$ be an initial store, $C$ be a clobber set, $\oton{E}$ be memory traces such that $E_i \nonint E_j$
for every $1 \leq i < j \leq n$.  Then any trace $E'$ satisfying $any\_interleave(E', \oton{E})$
is a valid interleaving.
\end{lem}
\begin{proof}
We will use Theorem~\ref{thm:sequential} and Lemma~\ref{lem:nonintswap} to make repeated swaps
of adjacent operations to transform $E'$ into $E_1 \dplus \ldots \dplus E_n$.  The symmetry of
Lemma~\ref{lem:nonintswap} is then used to show that the validity of $E_1 \dplus \ldots \dplus E_n$
under Theorem~\ref{thm:sequential} implies the validity of $E'$.

The swapping algorithm as follows: Associate with each operation in $E'$ the index of that operation
in $E_1 \dplus \ldots \dplus E_n$.  Any consecutive pair of operations $\epsilon_1$ and $\epsilon_2$
for which $index(\epsilon_1) > index(\epsilon_2)$ is a {\em misordered pair}.  For any such pair,
$\epsilon_1$ and $\epsilon_2$ will be from different constituent traces (if they were from the same
trace, their misordering would mean that $E'$ was not an interleaving of $\oton{E}$).  Every pair of
traces is non-interfering, so $\epsilon_1 \nonint \epsilon_2$ and they can be swapped while preserving
validity of the trace.  At most $n^2/2$ swaps are needed to eliminate all misordered pairs.
\end{proof}

We next extend non-interference to privileges.  We 
use the region mapping $M$ to translate the logical regions named in privileges into sets of
locations (i.e. physical regions), and define non-interference between individual privileges and
sets of privileges:
$$
\begin{array}{@{}l@{}}
priv_1(r_1) \nonint[M] priv_2(r_2) \Leftrightarrow (M(r_1) \cap M(r_2) = \emptyset)\ \vee \\
\hspace{0.35in} (priv_1 = reads \wedge priv_2 = reads)\ \vee \\
\hspace{0.35in} (priv_1 = reduces_{id_1} \wedge priv_2 = reduces_{id_2} \wedge id_1 = id_2)
\end{array}
$$
%% 
%% \begin{center}
%%  \begin{tabular}{llc}
%%  \multicolumn{1}{c}{$\phi_1$} & \multicolumn{1}{c}{$\phi_2$} & $\phi_1 \nonint[M] \phi_2$? \\
%%  \multirow{3}{*}{$reads(r_1)$} & $reads(r_2)$ & true \\
%%  & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  & $reduces_{id_2}(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  \cline{1-1}
%%  \multirow{3}{*}{$writes(r_1)$} & $reads(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  & $reduces_{id_2}(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  \cline{1-1}
%%  \multirow{3}{*}{$reduces_{id_1}(r_1)$} & $reads(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  & $writes(r_2)$ & $M(r_1) \cap M(r_2) = \emptyset$ \\
%%  & $reduces_{id_2}(r_2)$ & $\begin{array}{c}id_1 = id_2 \vee \\ M(r_1) \cap M(r_2) = \emptyset\end{array}$ \\
%%  \end{tabular}
%%  \end{center}

$$\Phi_1 \nonint[M] \Phi_2 \Leftrightarrow \displaystyle\bigwedge_{\phi_1 \in \Phi_1, \phi_2 \in \Phi_2} \phi_1 \nonint[M] \phi_2$$

\begin{lem}
\label{lem:nonintpriv}
\rm
Let $M$ be a region mapping and $E_1$ and $E_2$ two memory traces such that $\privconsist{E_1}{\Phi_1}$ and $\privconsist{E_2}{\Phi_2}$.  If $\Phi_1$ and $\Phi_2$ are non-interfering under $M$, then $E_1$ and $E_2$ must be non-interfering.
\end{lem}

\begin{proof}
The proof is by contradiction.  Assume $E_1 \cancel\nonint E_2$.  Then there must be some $\epsilon_1$ in $E_1$ and some $\epsilon_2$ in $E_2$ such that $\epsilon_1 \cancel\nonint \epsilon_2$.  This requires that
$\epsilon_1$ and $\epsilon_2$ be operations to the same location $l$ and not both be reads or both
reductions using the same function name.  Assume $\epsilon_1 = read(l, c_1, v_1)$ and $\epsilon_2 = write(l, c_2, v_2)$.  The definition of $\privconsist{E_1}{\Phi_1}$ tells us that there must be some $r_1$ satisfying $l \in M(r_1) \wedge reads(r_1) \in \Phi_1$.  Similarly, there must be some $r_2$ satisfying $l \in M(r_2) \wedge writes(r_2) \in \Phi_2$.  Letting $\phi_1 = reads(r_1)$ and $\phi_2 = writes(r_2)$, we have $M(r_1) \cap M(r_2) \not= \emptyset$ (it contains at least $l$), so $\phi_1 \cancel{\nonint[M]} \phi_2$ and $\Phi_1 \cancel{\nonint[M]} \Phi_2$.  There are seven other cases to consider (using $id_1 \not= id_2$ for the reduce-reduce case), but all yield the same result.
\end{proof}

We can now state the key result that allows the Legion runtime to safely evaluate subexpressions as
separate tasks in parallel.

\begin{thm}
\label{thm:parallelexec}
\rm
Let $\oton{e}$ be well-typed Legion expressions, each with its own privileges $\Phi_i$.
Let $M$ be a region mapping, $L$ a local value mapping,
$H$ a heap typing, and $S$ be an initial store satisfying
$\mapconsist{\Omega}$, $\localconsist{L}{\Gamma}$, and $\storeconsist{S}{H}$.
If $\Phi_i \nonint[M] \Phi_j$ for $1 \leq i < j \leq n$, then any parallel execution of expressions
$\oton{e}$ results in a valid interleaving of memory operations.
\end{thm}

The proof follows directly from Lemmas~\ref{lem:noninteffects}
and \ref{lem:nonintpriv}.  This result holds even if the
clobber set $C$ is non-empty, allowing two ``locally independent''
subtasks to run in parallel even if one or both of them is interacting
(in a programmer-permitted way) with a third subtask.

We now highlight an important aspect of the Legion
implementation that is different from other systems and relies on the
soundness of privileges.  Non-interference of memory operations is
something that can only be determined after evaluation of an
expression is completed, and only at great expense, as illustrated at
one extreme by work on transactional
memory \cite{Harris05}.  At the other extreme are systems
that check non-interference statically, such as Jade \cite{Rinard98} and
DPJ \cite{Bocchino11}.  In contrast, Legion checks non-interference of privileges
at runtime, which is much simpler and more efficient than checking
non-interference of dynamic memory traces.  Even though the privileges
themselves are static, the region mapping $M$ is dynamic.  Testing
non-interference on the privileges with physical regions allows
parallel execution in many more cases than a purely static analysis
can achieve.

\subsection{Relaxed Coherence}
\label{subsec:relaxed}

The analysis above does not consider the coherence modes of operations,
making the worst-case assumption that all operations use $excl$ mode.  However, if two or more
subexpressions contain calls to functions using $atomic$ or $simult$ coherence modes
on some regions, executions besides serial program order may 
result in valid interleavings even if constituent traces have locations in common.  We discuss
two cases of interest.  In both, the previous section's analysis 
is repeated with minor modifications, so for brevity we limit
discussion to a description of the differences rather than complete proofs.

The first important case occurs when all apparently-interfering operations have coherence modes
of either $atomic$ or $simult$.  In this case, the conditions required for
coherence and sequential equivalence are trivially satisfied for locations in $L_{excl}$.  The
requirements for locations in $L_{atomic}$ allow any permutation of the constituent traces,
so the runtime is free to serialize the execution of the subexpressions in any order, not just
program order.

To convert the required analysis to one that can be performed on regions, we modify
the non-interference operator for memory operations to consider only locations in $L_{excl}$ and show that
traces that are non-interfering under this modified check satisfy the requirement for atomic
execution.  We then modify the non-interference operator on privileges to allow 
$M(r_1) \cap M(r_2) \not= \emptyset$ if either $atomic(r_1)$ or $simult(r_1)$ is in $Q_1$ and
either $atomic(r_2)$ or $simult(r_2)$ is in $Q_2$. We then show that subexpressions with 
privileges that are {\em atomically non-interfering} (under a dynamic region mapping
$M$) always have memory traces compatible with atomically-reordered execution.  Finally,
in the special case where each subexpression has $atomic(r)$ or $simult(r)$ in $Q_i$ for every $r$
in $\Phi_i$, we can eliminate the dynamic check entirely.

The second case is a subset of the first, in which all apparently-interfering operations have
the $simult$ coherence mode.  In this case, all the coherence and sequential equivalence 
checks are trivially satisfied, and any interleaving of the constituent traces is valid.
We again modify the non-interference criteria for memory operations, traces, and privileges, this time
only allowing $simult$ coherence annotations. We show that subexpression with privileges
that are {\em simultaneously non-interfering} (under mapping $M$) can be executed
in parallel, which includes all the cases covered by Theorem~\ref{thm:parallelexec} as well
as those where the programmer has used the $simult$ coherence mode to explicitly allow parallel
execution.  Again, a special (but important) case exists where all privileges for each
subexpression are paired with $simult$ coherence mode declarations, allowing the safety of
parallel execution to be determined at compile time.

\section{Hierarchical Scheduling}
\label{sec:scheduling}

The enforcement of a (locally) valid interleaving at every step in the operational semantics
captures the hierarchical nature of the scheduler used by the Legion runtime.  As the cost of
a scheduling decision is quadratic in the number of function calls (more precisely, the number of 
privileges for those calls), it should be clear that a hierarchical scheduling algorithm will
be vastly more efficient than a global algorithm that must consider the privileges of every
concurrently executing expression.
However, a scheduling algorithm that makes decisions hierarchically must do so with limited
information, and must respect all required scheduling dependencies while ideally imposing as
few extra dependencies as possible.  The following theorem forms the basis of the 
Legion scheduler.

Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$.  Now let $e_3$ be some subexpression of $e_1$ and $e_4$ be some 
subexpression of $e_2$, with corresponding privileges $\Phi_3$ and $\Phi_4$.
By Theorem~\ref{thm:parallelexec}, if the runtime can verify $\Phi_1 \nonint[M] \Phi_2$, then 
expressions $e_1$ and $e_2$ may be evaluated in parallel.  The evaluation of the subexpressions
$e_3$ and $e_4$ will therefore be allowed to run in parallel, and we must show that this is
permissible.  We do so by proving that any memory traces that can result from the evaluation
of the subexpression must be non-interfering (allowing parallel execution by
Lemma~\ref{lem:noninteffects}).

\begin{thm}
\label{thm:hiersched}
\rm
Let $e_1$ and $e_2$ be well-typed Legion expressions using privileges $\Phi_1$ and
$\Phi_2$ respectively, where $\Phi_1 \nonint[M] \Phi_2$.  Let $e_3$ be a subexpression of $e_1$ and $e_4$ be a 
subexpression of $e_2$.  Any memory traces $E_3$ of $e_3$ and $E_4$ of $e_4$ resulting from evaluation of
$e_1$ and $e_2$ (with the usual consistent $M$, $L$, $H$, and $S$), must be
non-interfering.
\end{thm}

\begin{proof}
Let $E_1$ be the result of an evaluation of $e_1$, with $E_3$ being the memory trace from the
evaluation of $e_3$ within the expression tree.  $E_1$ will be formed from a tree of valid
interleavings, each of which must include all the memory operations of each constituent trace.
By induction over the number of intermediate subexpressions between $e_1$ and $e_3$, we
can show that any memory operation $\epsilon_3$ that is in $E_3$ must also be in $E_1$.
Similarly, any memory operation $\epsilon_4$ that is in $E_4$ must also be in $E_2$.  We have
$E_1 \nonint E_2$ from Lemma~\ref{lem:nonintpriv}, so and therefore $\epsilon_3 \nonint \epsilon_4$.  Since
this holds for all $\epsilon_3$ in $E_3$ and all $\epsilon_4$ in $E_4$, we have $E_3 \nonint E_4$.
\end{proof}

 The Legion task scheduler uses Theorem~\ref{thm:hiersched} in the
following way.  Sibling function calls (those invoked within the same
function body) are checked for non-interference of their (dynamic)
privileges.  If they interfere they are serialized; otherwise they are
considered for execution as parallel subtasks.  Although determining that $e_1$
and $e_2$ are non-interfering guarantees $e_3$ and $e_4$ are
non-interfering, the converse is not true.  It is possible for $e_1$
and $e_2$ to be judged to be potentially interfering and be serialized
even when the subexpressions are non-interfering.  This loss of
precision can impact performance, and did in a few instances in our
experiments.  
%When the performance impact is significant, two options
%exist to improve matters.  
In cases where $e_1$ and $e_2$ are in fact
non-interfering, region partitions can be further refined to make the
disjointness visible in the privileges.  Alternately, the code can be
refactored to lift the subexpressions up to the same level, allowing
the dynamic test based on the region mapping to infer non-interference
in more cases.  
%Although this can increase the overhead of scheduling
%decisions, judicious use of this technique in practice has not shown
%any measurable performance degradation.