### 2012-07-17 doc/asplos2013/micro.tex
Getting the asplos paper off the ground

\section{Microbenchmark Evaluation}
\label{sec:micro}