### 2012-07-20 doc/asplos2013/conclusion.tex
Adding a complete introduction

\section{Conclusion}
\label{sec:conclusion}

Conclusion goes here.