### 2012-07-20 doc/asplos2013/intro.tex

\section{Introduction}
\label{sec:intro}

% Big machines -> latencies are growing
% Processors can't be blocked, need to be able to continue handling work
% Operations need to be composable
% deferred execution model

% Contributions:
% - An event system where every operation can take an event and returns an event
% - Handles are globally valid
% - More flexible synchronization primitive for this environment: deferred locks
% - Physical regions for data movement including special versions for reductions

% Parallel programs have three parts
%  Parallel control
%  Data movement in the memory hierarchy
%  Synchronization

% Bulk-synchronous
% asynchronous components are not composable

With the multicore revolution, modern supercomputers have continued to scale
by increasing the number of processors available for computation.  The
price of this scaling is complex interconnection networks and
deep memory hierarchies.  These parts of the system are designed to
minimize latency, but the growing scale of supercomputers has
caused average and worst-case latencies to grow by orders of magnitdue.
Recently, some of these latencies have grown to the point that supercomputers
share many of the same performance properties as machines commonly
classified as distributed systems.  Despite this major evolution in the
performance of supercomputers, most programs for supercomputers are still
written using programming models and libraries that were designed
for machines many orders of magnitude smaller\cite{MPI,COARRAY_FORTRAN,UPC99}.

The approach that current programming models provide is commonly reffered to
as a {\em bulk-synchronous} model\cite{Valiant90}.  Bulk-synchronous models
break computations down into three kinds of phases: parallel computation, data
communication, and synchronization.  At any given point in time all the processors
in the system are engaged in one of the three kinds of phase.  The bulk-synchronous
approach to parallel programming makes it easy to reason about parallel execution, but
has a fundamental limitation: it is very difficult to hide long-latency communication
or memory movement phases.  In older machines this was not a problem because slower
processors and smaller machines meant computation phases were much larger than
communication or synchronization phases.  However, today's machines are much
larger with faster processors which cause increasingly large percentages of time
to be spend in the communication and synchronization phases.

To combat this problem, parallel programming libraries have introduced
asynchronous constructs which allow the programmer to overlap phases.  For example,
MPI has asynchronous {\em send} and {\em receive} operations for sending
messages in parallel with computation.  There are two problems with this approach.
First, not all phases support asynchronous constructs.  Second,
the asynchronous operations do not {\em compose}.  There is no way to express
that an asynchronous computation should begin as soon as an asynchronous
communication operation has completed.  The result is that
at some point there must be blocking operations to coordinate synchronous
and asynchronous phases.  If blocking calls are not accurately placed in the
program they can lead to processor stalls that expose latency.  Correct
placement of blocking calls can be dependent on input data, algorithmic decisions like blocking
factors, or even the underlying hardware.  
To hide as much latency as possible and to minimize processor stalls, regardless
of inputs, algorithmic choices, or underlying hardware,
all three major aspects of parallel programs (computation, data movement, synchronization)
should be made asynchronous and composable.

We propose a new interface for programming modern supercomputers 
that is fully asynchronous to hide the large latencies associated with current 
interconnection networks and memory hierarchies.  Our interface makes it possible
to asynchronously compose computation, data movement, and synchronizatioan.  The 
fundamental mechanism that enables us to compose all three aspects of parallel
programs are {\em events}.  Events provide a mechanism for naming a point in the future
when an operation (i.e. computation, data copy, synchronization) will complete.  
Every call to perform an operation in our runtime returns immediately with an event that will 
{\em trigger} when the operation has completed.  Furthermore, every call
to perform an operation can take a precondition event that must trigger before
the operation can begin.  Using events programmers can compose large
chains of dependent operations.  The runtime is free to execute operations
in any order that obeys the event dependences.  With this freedom the
runtime can schedule operations in a way that optimizes throughput
while automatically hiding the long latency operations.

The following sections describe our contributions:
\begin{itemize}
\item We present an interface for asynchronously composing
parallel computation, data movement, and synchronization based on events.  
To support synchronization in an event system we describe a novel synchronization
primitive called a {\em deffered lock}.  Our interface also supports efficient
mechanisms for performing distributed reduction operations (Section~\ref{sec:interface}).

\item We describe a distributed implementation of our interface that requires no
global coordination and is capable of running on current supercomputers.  We
describe optimizations for the implementation of events, deffered locks, and
reductions (Section~\ref{sec:impl}).

\item We show the result of micro-benchmarks that are designed to stress
our implementation.  The results indiciate that our implementation approaches the performance of
the underlying hardware (Section~\ref{sec:micro}).

\item We demonstrate that our interface is powerful enough to support a higher-level
runtime system and that real applications running on this stack of software
are capable of achieving high-performance (Section~\ref{sec:apps}).
\end{itemize}

In Section~\ref{sec:related} we describe related work and in Section~\ref{sec:conclusion}
we conclude.