### 2012-07-20 doc/asplos2013/related.tex
Adding some related work and updating the bibliography

\section{Related Work}
\label{sec:related}

MPI is the current industry standard for programming large
supercomputers\cite{MPI}.  While MPI supports asynchronous
communication operations there is no mechanism for composing
asynchronous operations.  Co-array Fortran, UPC, and Titanium
are array based languages that implement bulk-synchronous
programming models similar to MPI that allow for asynchronous
exchanges of data\cite{COARRAY_FORTRAN,UPC99,JV:Yel98}.  Like
MPI however, none of these languages allow for general composition
of asynchronous tasks, communication, and synchronization.

Both POSIX threads and OpenMP\cite{OPENMP98} are used for intra-node parallel
programming in large machines, but neither support any asynchronous
operations.  CUDA\cite{CUDA} and OpenCL\cite{Khronos:OpenCL} support
the compostition of asynchronous kernel launches and asynchronous copies
between a host node and a single GPU.  However, the only synchronization
options available in both interfaces are blocking operations on either
a stream or a work queue.  OpenCL has events similar to our interface for
expressing ordering, but has no synchronization primitive for expressing
more relaxed properties such as atomicity that can be described by a
deferred lock.  OpenCL events are not valid globally, but can only be
used with a single GPU context.

Cilk is a parallel programming model that has demonstrated the power of
asynchronous function calls\cite{CILK95}.  The Cilk {\em work-first}
principle provides a compelling theoretical argument for parallel programming
interfaces such as ours that optimize for throughput at the potential
expense of adding additional latency to the critical 
dependence path\cite{Frigo98}.

Chapel\cite{Chamberlain:Chapel} and X10\cite{X1005} are high-level parallel
programming languages that support asynchronous operations.  The constructs
introduced in these languages are higher-level and have complex 
semantics.  We view our interface as providing a potential target
for implementing these higher-level asynchronous operations in an efficient
manner similar to Legion\cite{Legion12}.

The implementation of our interface shares many similarities with large
distributed systems.  Many distributed system implement a publish/subscribe
abstraction for supporting communication that is very similar to how
we manage events and event waiters\cite{Aguilera99,Carzaniga01}.  Work has
also been done on using object oriented languages to build 
distributed systems \cite{Eugster01,Harrison97,Chang91}.  In these cases
callbacks are registered to run when event operations are triggered
from remote locations.