### 2012-07-20 doc/asplos2013/legion-llr.tex

\documentclass{sig-alternate}

\usepackage{amsmath}
%\usepackage{amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{graphics}
\usepackage{url}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}

\begin{document}

\title{High-Performance Language Primitives for Distributed Architectures using Asynchronous Events}
\numberofauthors{3}
\author{
\alignauthor Sean Treichler\\
             \affaddr{Stanford University} \\
             \email{sjt@cs.stanford.edu}
\alignauthor Michael Bauer \\
             \affaddr{Stanford University} \\
             \email{mebauer@cs.stanford.edu}
\alignauthor Alex Aiken \\
             \affaddr{Stanford University} \\
             \email{aiken@cs.stanford.edu}
}
\maketitle

\begin{abstract}
%The limit is 12 pages, including bibliography and appendices.
%Deadlines:
%\begin{tabular}{ll}
%Abstract deadline & July 16, 2012 3pm PDT \\
%Paper deadline & July 23, 2012 3pm PDT \\
%Author response & around October 15, 2012 \\
%Author notification & November 9, 2012
%\end{tabular}
We present a low-level runtime interface for current highly distributed, heterogeneous
parallel machines that provides a fully asynchronous 
programming model, allowing all types of runtime actions, including synchronization,
to be non-blocking.  Our interface includes mechanisms for the creation and movement
of data through the memory hierarchy as well as {\em deferred locks}, a novel synchronization primitive.
Asynchrony is exposed by the the interface using a distributed event system.  All
operations return an event to be triggered when the operation completes.  Events
can be chained together and used to create dependences between operations.

We describe an implementation of our interface capable of running on
a large supercomputer with both CPU and GPU processors and a high speed interconnect.
We employ several micro-benchmarks to show that our implementation approaches
the performance limits of the underlying hardware.
We also demonstrate that the interface is sufficiently powerful to serve as the foundation 
of an advanced parallel programming system.  We measure the performance of three real-world applications
that use our system on the Keeneland supercomputer.
\end{abstract}

%\input{outline}
\input{intro}
\input{interface}
\input{impl}
\input{micro}
\input{results}
\input{related}

\input{conclusion}

{
\small
\bibliographystyle{abbrv}
\bibliography{bibliography}
}

\end{document}