### 2012-08-04 doc/sc2012/legion.tex
Adding some stuff for the final version of the SC paper

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts 
\bibliographystyle{IEEEtran}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{graphics}
\usepackage{url}
\usepackage{minibox}

\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{observation}[theorem]{Observation}
\IEEEoverridecommandlockouts

\begin{document}

\pdfpagewidth=8.5in
\pdfpageheight=11in

\title{Legion: Expressing Locality and Independence with Logical
  Regions\thanks{This work was supported by the Army High Performance
    Research Center (DARPA contract HR0011-10-9-0008), a subcontract
    from Los Alamos National Laboratory for the DOE Office of Science,
    Advanced Scientific Computing Research (DE-AC52-06NA25396), and NVIDIA.  The views and
    conclusions contained in this document are those of the authors
    and should not be interpreted as representing the official
    policies, either expressed or implied, of the U.S. Government.}}
%\author{ \IEEEauthorblockN{Michael Bauer\IEEEauthorrefmark{1},
%%\authorinfo(Stanford University, email: \mbox{mebauer@cs.stanford.edu}
%Sean Treichler\IEEEauthorrefmark{1}, 
%%\authorinfo(Stanford University, email: \mbox{sjt@cs.stanford.edu}
%Elliott Slaughter\IEEEauthorrefmark{1}, 
%% \authorinfo(Stanford University, email: \mbox{slaughter@cs.stanford.edu}
%Alex Aiken\IEEEauthorrefmark{1}  
%%\authorinfo(Stanford University, email: \mbox{aiken@cs.stanford.edu}
%%\email{aiken@cs.stanford.edu}
%}
%\IEEEauthorblockA{
%\IEEEauthorrefmark{1}{Computer Science Department, Stanford University}
%}
%}

\author{
\IEEEauthorblockN{Michael Bauer}
\IEEEauthorblockA{Stanford University \\ mebauer@cs.stanford.edu}
\and
\IEEEauthorblockN{Sean Treichler}
\IEEEauthorblockA{StanfordUniversity \\ sjt@cs.stanford.edu}
\and
\IEEEauthorblockN{Elliott Slaughter}
\IEEEauthorblockA{Stanford University \\ slaughter@cs.stanford.edu}
\and
\IEEEauthorblockN{Alex Aiken}
\IEEEauthorblockA{Stanford University \\ aiken@cs.stanford.edu}
}

\IEEEpubid{\makebox[\columnwidth][l]{\minibox{SC12, November 10-16, 2012, Salt Lake City, Utah, USA \\ 978-1-4673-0806-9/12/ \$31.00~\copyright~2012 IEEE}}
\hspace{\columnsep}
\makebox[\columnwidth]{}
}

\maketitle

\begin{abstract}
Modern parallel architectures have both heterogeneous processors and
deep, complex memory hierarchies.  We present Legion, a programming
model and runtime system for achieving high performance on these
machines.  Legion is organized around {\em logical regions}, which
express both locality and independence of program data, and {\em
  tasks}, functions that perform computations on regions.  We describe
a runtime system that dynamically extracts parallelism from
Legion programs, using a distributed, parallel scheduling algorithm
that identifies both independent tasks and nested parallelism.  Legion
also enables explicit, programmer controlled movement of data through
the memory hierarchy and placement of tasks based on locality
information via a novel mapping interface.  We evaluate our Legion
implementation on three applications: fluid-flow on a regular grid,
a three-level AMR code solving a heat diffusion equation, and 
a circuit simulation.
%
% FIXME: put some performance numbers here
%
\end{abstract}

\section{Introduction}
\label{sect:intro}
%Put intro text here \cite{Fatahalian06}

Modern parallel machines are increasingly complex, with deep,
distributed memory hierarchies and heterogeneous processing units.  Because the
costs of communication within these architectures vary by several orders of magnitude, 
the penalty for mistakes in the placement of data or computation is usually very poor performance.  Thus,
to achieve good performance the programmer
and the programming system must reason about {\em locality} (data resident
close to computation that uses it) and {\em independence} (computations
operating on disjoint data, and therefore not requiring communication and able to be placed in possibly distant parts of the machine).  
Most contemporary
programming systems have no facilities for the programmer to express locality
and independence.  The few languages that do focus primarily on
array-based locality \cite{Fatahalian06,CHAPEL04,UPC99} and 
avoid irregular pointer data structures.  
%
% FIXME
% Sean would like a bit more on why other systems don't support pointer data structures (e.g., the difficulty of
% alias analysis.  I thought about it and feel it is too much of a digression: mentioning that in the first paragraph
% makes it sound like it will be an important part of the paper.
%
%encode such structures in arrays, which 
%data structures encoded in arrays which are impossible for compilers
%and runtime systems to reason about with regard to locality and independence.

In this paper we describe Legion, a parallel programming system based
on using {\em logical regions} to describe the organization of data
and to make explicit relationships useful for reasoning about locality
and independence.
A logical region names a set of objects.  Logical regions are first-class 
values in Legion and may be dynamically allocated, deleted and stored
in data structures. Regions can also be passed as arguments
to distinguished functions called {\em tasks} that access the data in
those regions, providing locality information.  Logical regions may be
{\em partitioned} into disjoint or aliased (overlapping) {\em subregions}, providing
information for determining independence of computations.
Furthermore, computations access logical regions with particular {\em
  privileges} ({\em read-only}, {\em read-write}, and {\em reduce}) and {\em coherence} (e.g., {\em exclusive access} and {\em
  atomic access}, among others).  Privileges express how a task may use
its region arguments, providing data dependence information that
is used to guide the extraction of parallelism.  For example, if two tasks
access the same region with read-only privileges the two tasks
can potentially be run in parallel.
Coherence properties express the required semantics of concurrent region accesses.
For example, if the program executes $f_1(r); f_2(r)$ and tasks $f_1$
and $f_2$ both require {\em exclusive} access to region $r$, then Legion guarantees the result will
be as if $f_1(r)$ completes before $f_2(r)$ begins.   On the other hand, if
the tasks access $r$ with {\em atomic} coherence, then Legion
guarantees only atomicity of the tasks with respect to $r$: either task $f_1(r)$
appears to run entirely before $f_2(r)$ or vice versa.

Logical regions do not commit to any particular layout of the data or
placement in the machine.  At runtime, each logical region has one or
more {\em physical instances} assigned to specific memories.  It is often useful to have multiple
physical instances of a logical region (e.g., to replicate read-only data, or to allow
independent reductions that are later combined).

To introduce the programming model, we present
a circuit simulation in
Section~\ref{sec:ex}, illustrating regions, tasks,
permissions and coherence properties, the interactions between them,
and how these building blocks are assembled into a Legion
program.  Subsequent sections each describe a contribution in the
implementation and evaluation of Legion:

\input{code_ex}

\begin{itemize}

\item We define a {\em software out-of-order processor}, or SOOP, for
  scheduling tasks with region arguments in a manner analogous to how
  out-of-order hardware schedulers process instructions with register
  arguments (Section~\ref{sec:soop}).  In addition to pipelining the execution of tasks over
  several stages, our SOOP is distributed across the machine and is
  also hierarchical and naturally extracts nested
  parallelism (because tasks may recursively spawn subtasks).

%\item We use a {\em deferred execution} model that extracts parallelism
%even from highly irregular, unbalanced computations.  Instead of blocking,
%an operation $o$ that cannot immediately execute because it is dependent
%on some other operation $o'$ waits on an event that occurs when $o'$ completes.
%Events can be created for the completion of any operation in Legion, and 
%the scheduler can build long chains of deferred operations.  In Section~\ref{sec:exp}
%we show that dataflow-like deferred execution gives us a significant performance improvement
%over bulk synchronous global synchronization primitives such as barriers.

\item Of central importance is how
tasks are assigned (or {\em mapped}) to processors and how physical instances of logical regions are
mapped to specific memory units (Section~\ref{sec:mapping}).  Often using
application-specific information results in better
mappings than a generic mapping strategy.  We describe
a {\em mapping interface} that allows programmers to give the SOOP a specification
of how to map tasks and regions for a specific application, or even part of an
application.  This mapping API is designed so that
any user-supplied mapping strategy can only affect the performance of applications,
not their correctness.  
%Thus, user mappings only affect the time and resources required
%to run the program and cannot change the program's output.

\item We present results of experiments on three applications:
  fluid-flow on a regular grid, a three-level AMR code solving a
  heat diffusion equation and a circuit simulation.
  We compare each application with the best reference
  versions on three different clusters of multicore processors with
  GPUs, including the Keeneland supercomputer \cite{Keeneland}.
%
% FIXME Give the same numbers as in the abstract.
%

\end{itemize}

\input{circuit}
\input{part_fig}
\input{soop}
\input{mapping}
\input{experiments}
\input{related}
%\input{conclusion}

\bibstyle{IEEEtran}
\bibliography{bibliography}

\end{document}