### 2012-09-12 doc/popl2013/response.txt
couple more edits, answers to the last open questions, spell check

We thank the reviewers for their very detailed comments.  Most of the
comments are specifically about presentation; we acknowledge the
issue, and if the paper is accepted we will address the concerns
raised.

Many of the reviewers raise similar questions, which we have grouped
together in a single response below in addition to answering issues
raised by individual reviewers.  To keep the response to a reasonable
length we have tried to address the most common and significant
issues.  In many cases, questions were asked which were addressed in
our prior published work [1], which may have been unavailable to the
reviewers previously but can now be found at
http://theory.stanford.edu/~aiken/publications/papers/sc12.pdf .

Reviewers A, B, and C all asked for additional details about colorings.
In Core Legion, colorings are explicit mappings of region elements (named by their location)
to integer "colors".  This represents the most general case of describing the partitioning of
a region.  More concise descriptions exist in the full Legion language (e.g. using a pure
function to determine the color of an element on demand), but a single form was selected for
Core Legion for simplicity.  As described in [1], Legion does not attempt to solve the (in many cases still-open)
problem of partitioning data for any given application.  In our example, the body of the "color_circuit"
task would require the programmer to compute a partition of the circuit graph and return a coloring
that assigns each circuit element a color corresponding to its piece of the partition.  The same
set of colors are used to identify the subregions that result from the partitioning.
In production scientific codes this is often done by a call 
to an external library (e.g. PARMETIS for graph-based data structures).  
A coloring is merely a way to capture the result of the partitioning 
computation in a general format understood by the Legion runtime.

The principle reason to support multiple partitions of the same data is to allow precise descriptions
of the sets of elements that will be accessed by a given task.  These sets are often different in 
different phases of an algorithm, and the increased precision is necessary to safely parallelize
applications such as the circuit example.  Although the subregions of different partitions are often
accessed with different privileges, this is a property of an algorithm rather than a Legion requirement.

Reviewer A asked how "downregion" operations are performed and used.  The "downregion" operation
requires a dynamic check that the given location is valid in the provided logical region.  To
perform these checks the runtime remembers which locations have been allocated in each logical
region and dynamically checks that the location has been allocated for the given region.  The 
"build_lists" task is an example of an optimization done to reduce the number of "downregion" calls.
While a task could iterate over the elements of a subregion through repeated uses of "downregion" on
each element of the parent region, by building a list typed on the subregion up front, the cost 
of those "downregion" checks can be amortized over multiple iterations of the simulation.

Both reviewers A and B inquired how the dynamic disjointness tests are performed.  This
is a primary subject of our prior work [1].

Reviewer B commented that the Legion type system is similar in several ways to existing region-based
typed systems.  While we reuse many features of prior region languages, we believe that Legion
represents a fundamentally new composition of these features.  Primarily Legion allows for first-class
logical regions which permit elements to exist in many regions simultaneously via multiple dynamic
partitions of regions.  Furthermore, the way the region information is leveraged both at compile-time 
and at runtime represent a new point in the design space of region languages.  While the proof of
soundness does not introduce novel proof techniques, it is also far from immediate that the type system
is sound and we feel the proof is necessary given large differences in the semantics of logical regions 
from previous region systems.  The soundness proof was also a prerequisite for our later proofs
concerning coherence modes and hierarchical scheduling.

Both reviewers B and C thought coherence modes were an interesting technical contribution, but
would have liked a higher-level description to build intuition.
A description of the usage and intuition behind coherence modes is presented in [1] and
was omitted from this paper due to space constraints.  Our goal in this paper was to present
two new results about coherence modes that are not in [1].  First, we present a formal semantics 
of the possible interleavings allowed by different coherence modes and their interactions.  
Second, we prove that coherence modes do not impact the soundness of our type system or the 
correctness of our hierarchical scheduling algorithm.  A brief summary of coherence modes is:
"exclusive" coherence corresponds to apparently-sequential execution, "atomic" corresponds 
to transactional-memory-like execution, and "simultaneous" corresponds to
"all bets are off" pthreads-style parallelism.  However, unlike other implementations, 
Legion's coherence modes only allow for reordering of subtasks with the same parent task, 
permitting the containment of non-determinism to a portion of the task hierarchy.

Reviewer A asked if the operational semantics would be simpler if the only relaxation of sequential
semantics came from an explicit "par(e1,e2)" construct.  While the semantics would indeed be simpler,
the use of a static construct (rather than one able to consider dynamic region mappings) would limit
the expressible forms of parallelism to a (well-studied) subset of what Legion programs can achieve.

Reviewer D asked what is currently implemented.  Legion currently consists of a static type 
checker for programs written in Core Legion and a C++ runtime library for fully-featured Legion programs.

Finally, reviewers B, C, and D expressed confusion about the nature of the clobber set.
The choice of large-step operational semantics was made to match the hierarchical and distributed
scheduling used by the Legion runtime.  When coherence modes other than exclusive are used, there may be
multiple tasks modifying the same location of memory - the clobber set is the set of all such locations for
a given task.  The clobber set is used to bound the allowed non-determinism within a given task, and the
fact that the top-level task (which has no sibling tasks) has an empty clobber set guarantees that any 
execution of the program is consistent with a standard sequentially consistent memory model.

An advantage of the large-step semantics with the use of clobber sets is that our soundness result holds for
any scheduling of tasks that satisfies the "valid_interleave" rules.  We considered the use of small-step semantics
that explicitly track the state of every parallel task's execution, but the complexity of the constraints used
to decide which task(s) were able to take a step at any given point were onerous and would limit our soundness
guarantees to only a particular scheduling algorithm.