### 2012-09-13 doc/popl2013/response.txt
udpate

We thank the reviewers for their very detailed comments.  Most of the
comments are specifically about presentation.  If the paper is accepted 
we will certainly address all the concerns raised; below we focus on what
appear to be the most common or important points.

In many cases, questions were asked which were addressed in our prior
published work [1], which may have been unavailable to the reviewers
previously but can now be found at
http://theory.stanford.edu/~aiken/publications/papers/sc12.pdf .

Reviewers A, B, and C asked for additional details about colorings.
In Core Legion, colorings are explicit mappings of region elements
(named by their location) to integers ("colors"), which represents the
most general case of describing the partitioning of a region.  More
concise descriptions exist in the full Legion language (e.g. using a
pure function to determine the color of an element on demand), but a
single form was selected for Core Legion for simplicity.  As described
in [1], Legion does not attempt to solve the (in many cases
still-open) problem of optimally partitioning data for any given
application, but instead provides a mechanism for programmers to write
their own partitioning code and communicate the partitioning decisions
to the Legion runtime.  In our example, the body of the
"color_circuit" task would require the programmer to compute a
partition of the circuit graph and return a coloring that assigns each
circuit element a color; each color identifies a separate subregion of
the partition consisting of all the circuit elements with that color.
In production scientific codes partitioning is often done by a call to
an external library (e.g. PARMETIS for graph-based data structures).
A coloring is a way to capture the result of the partitioning
computation in a general format understood by the Legion runtime.
More information on colorings can be found in [1].

Reviewer C correctly points out that regions are central to
performance as well as for the type structure of Legion programs.
Regions are in Legion to allow programmers to describe locality and
independence for structuring parallel computations.  The programmer's
choice of egions can affect load balancing, communication, and the
granularity of parallel tasks.  In the case of the circuit example, no
partitioning scheme that does not look at the actual structure of the
graph being partitioned can pick a partition that will lead to good
performance; that is, for performance reasons the partitioning should
be done dynamically.  Assuming Legion has a sound type system,
however, any partitioning should be correct.  The paper focuses on the
correctness issues for the Legion implementation: that there are no
wild pointers to the wrong regions and that tasks observe both
privilege and coherence requirements.

Reviewer C asked for clarification concerning the the relationship
between multiple partitions and privileges.  The principle reason to
support multiple partitions of the same data is to allow precise
descriptions of the sets of elements that will be accessed by a given
task.  These sets are often different in different phases of an
algorithm, and both naming the precise set of nodes and the privileges
required are needed to safely parallelize applications such as the
circuit code.  In some cases (like the circuit code) it is possible in
theory to have a single partitioning of the data by taking the product
of all the overlapping partitions, but this results in many, many
unintuitive distinct subregions and is extremely unpleasant
to program with (we tried it once).

Reviewer A asked how "downregion" operations are performed and used.
The "downregion" operation requires a dynamic check that the given
location is valid in the provided logical region.  To perform these
checks the runtime remembers which locations have been allocated in
each logical region and dynamically checks that the location has been
allocated for the given region.  The "build_lists" task is an example
of an optimization done to reduce the number of "downregion" calls.
While a task could iterate over the elements of a subregion through
repeated uses of "downregion" on each element of the parent region, by
building a list typed on the subregion up front, the cost of those
"downregion" checks can be amortized over multiple iterations of the
simulation.

Both reviewers A and B inquired how the dynamic disjointness tests are
performed.  This is a primary subject of our prior work [1].  Briefly,
the runtime keeps track of the tree of region partitions and the tasks
that are currently using those regions (and with what
privileges/coherence).  When a new task is launched its region
arguments (and required privileges/coherence) are checked against the
current region tree to decide which, if any, of the already running
tasks must complete before the new task can execute.  This test can be
made provably efficient, even when scheduling is concurrent and
distributed, but relies crucially on the soundness of the type system
for correctness.

Reviewer B commented that the Legion type system is similar in several
ways to existing region-based typed systems.  While we reuse many
features of prior region languages, we believe that Legion represents
a new composition of these features.  Primarily Legion allows for
first-class logical regions which permit elements to exist in many
regions simultaneously via multiple dynamic partitions of regions.
Furthermore, the way the region information is leveraged both at
compile-time and at runtime represent a new point in the design space
of region languages.  We would like to point out specifically the way
in which existential types are used to read (packed) regions from the
heap and open them with permission to read/write/reduce values in
those regions - there is a subtle interaction between free variables
in existential types and bounded quantification of privileges, but
this is what allows Legion to check statically that a task can, say,
iterate over an unbounded list L of subregions and access elements of
those regions so long as there is a constraint that shows the task has
the same permissions on some ancestor region of every subregion in L.
And while the proof of soundness does not introduce novel proof
techniques, it is far from immediate that the type system is sound and
we felt the proof was necessary given the large differences in the
semantics of logical regions from previous region systems.  The
soundness proof was also a prerequisite for our later proofs
concerning coherence modes and hierarchical scheduling.

Both reviewers B and C thought coherence modes were an interesting
technical contribution, but would have liked a higher-level
description to build intuition.  A description of the usage and
intuition behind coherence modes is presented in [1] and was omitted
from this paper due to space constraints.  In this paper we present
two new results about coherence modes that are not in [1].  First, we
present a formal semantics of the possible interleavings allowed by
different coherence modes and their interactions.  Second, we prove
that coherence modes do not impact the soundness of our type system or
the correctness of our hierarchical scheduling algorithm.  A brief
summary of coherence modes is: "exclusive" coherence corresponds to
apparently-sequential execution, "atomic" corresponds to
transactional-memory-like execution, and "simultaneous" corresponds to
"all bets are off" pthreads-style parallelism.  However, unlike other
implementations, Legion's coherence modes only allow for reordering of
subtasks with the same parent task, permitting the containment of
non-determinism to a portion of the task hierarchy.

Reviewer A asked if the operational semantics would be simpler if the
only relaxation of sequential semantics came from an explicit
"par(e1,e2)" construct.  While the semantics would indeed be simpler,
the use of a static construct (rather than one able to consider
dynamic region mappings) would limit the expressible forms of
parallelism to a (well-studied) subset of what Legion programs can
achieve.

Reviewer D asked what is currently implemented.  Legion currently
consists of a static type checker for programs written in Core Legion
and a C++ runtime library for fully-featured Legion programs.

Finally, reviewers B, C, and D expressed confusion about the nature of
the clobber set.  The choice of large-step operational semantics was
made to match the hierarchical and distributed scheduling used by the
Legion runtime.  When coherence modes other than exclusive are used,
multiple tasks may modify the same location of memory - the clobber
set for a task is the set of all such locations that may be updated by
other tasks.  The clobber set is used to bound the allowed
non-determinism within a given task, and the fact that the top-level
task (which has no sibling tasks) has an empty clobber set guarantees
that any execution of the program is consistent with a standard
sequentially consistent memory model.

An advantage of the large-step semantics with the use of clobber sets
is that our soundness result holds for any scheduling of tasks that
satisfies the "valid_interleave" rules.  We considered the use of
small-step semantics that explicitly track the state of every parallel
task's execution, but the complexity of the constraints used to decide
which task(s) were able to take a step at any given point were
onerous; to manage this complexity we were concerned we would have to
limit our soundness guarantees to a particular scheduling algorithm.

Reviewer A asked why the E-NewColor rule allows the initial coloring
to be non-empty.  There are two reasons for this.  First, in the
large-step semantics we use, the region mapping and heap typing cannot
be changed by an E-New rule, so a partitioning of a region must
include extra elements of the appropriate heap type to support
allocation within a subregion.  Second, it allows implementations in
which an existing element that was not colored by the programmer is
included in (at most one) arbitrary subregion, which can improve the
efficiency of data transfers at runtime.