### 2012-10-17 doc/asplos2013/response.txt
edits

We would like to thank the reviewers for the detailed feedback.  Below we will...

Reviewers 1, 2, and 4 requested clarification of which components of the the API are
novel and which are similar to prior work.  Although events, locks, barriers, and reductions
have appeared in many forms, we believe the following contributions are novel:
  a) the use of generational events to efficiently describe the very large number of dependencies
that exist in a distributed application
  b) the deferred lock construct in which access to a critical resource can be mediated without
blocking operations that can waste execution resources
  c) the ability to modify a barrier's expected arrival count which allows nested task subdivision
without requiring a parent task to explicitly wait for subtasks to finish
  d) a general treatment of reductions that allows arbitrary reduction operations and improved
efficiency through specialized reduction instances that allow individual reduction operations to be
deferred and applied in bulk
  e) the ability to asychronously compose all of these constructs

Reviewers 1 and 5 had several questions about our handling of reductions.  One source of confusion
was the use of "commutativity".  The traditional use of these terms is limited to binary operations
in which both the left and right operands are of the same type.  Our implementation allows reductions
that have different left and right operand types (e.g. scaling of a vector by a scalar) and treats
each reduction operation as a unary function applied (atomically) to the current value (e.g. V *= 3
becomes (*3)(V) ).  When multiple reductions are applied to the same value, these unary functions
can be composed, and it's the commutativity of that composition
(e.g. [(*3) o (*2)](X) = [(*2) o (*3)](X) for all X) that allows the application of these
reduction operations to be deferred and reordered.  Nearly all "reasonable" reduction operations
that one might propose satisfy this form of commutativity.  Most also satisfy our criterion of foldability
(e.g. [(*3) o (*2)](X) = (*6)(X) for all X), allowing the use of our reduction fold instances to further
improve performance in most cases.

Reviewer 1 noted that use "Distributed Architectures" in our title might be confusing.  Our goal was
to distinguish our implementation from those that are limited to shared-memory systems.  A more
accurate description would be "Distributed Memory Architectures."  We would be willing to change the 
title to clarify our meaning if the submission is accepted.

Reviewer 2 noted that OpenCL supports events for asynchronously launching tasks and moving data. 
As we noted in our related work, OpenCL events are only valid in a single address space (e.g. node),
whereas our events which are valid in every address space on distributed memory machines.  Furthermore, the
synchronization primitives in OpenCL do not use events and cannot compose with other operations.

Reviewer 3 expressed doubts about the performance of the Legion version of the applications 
relative to hand-coded applications and whether the applications fit well into the bulk-synchronous model.  
The fourth citation in our paper [4] (which will appear at Supercomputing 2012) compares these same 
applications written in Legion to hand-coded reference implementations, and shows the Legion versions outperforming
the existing implementations on larger machine sizes.  All three reference implementations, two of which
were written by third parties, were coded in a bulk-synchronous style.  In this paper, we aimed to quantify 
what portion of that speedup results from the composable, asynchronous nature of our 
runtime system.  To factor out any performance benefits gained by the Legion programming model 
independent from our runtime, we compared against bulk-synchronous version of the applications from [4] 
written in Legion instead of the reference implementations.

Reviewer 4 requested a comparison to existing dataflow programming models.  A more apt comparison
for dataflow models would be the Legion programming system which is beyond the scope of this paper.  
Instead, we view the runtime presented in this paper as orthogonal work to dataflow programming models.   
The runtime presented here provides a lower-level layer of abstraction that could serve as the 
foundation for many different dataflow programming systems.

Reviewer 4 was curious about the difference between our deferred locks and I-structures from the dataflow
literature.  I-structures support a producer-consumer, write-once semantics which is much closer
to the functionality provided by events in our runtime.  In contrast, deferred locks provide
unordered atomic semantics, permitting multiple writes to an object and allowing programmers to describe
properties like serializability without the requirement of ordering.  We are unaware of any feature of
dataflow implementations that support a similar semantics to deferred locks.

Reviewer 1 commented that blocking APIs do not necessarily stall processors if there are enough 
runnable threads to keep the processor busy.  While this is true, the overhead of having many threads
to support this kind of computation is high due to the cost of software threads on CPUs.  (On GPUs, threads
are much lighter-weight, but current programming models do not allow these threads to be suspended.)
Instead our system provides lighter-weight tasks which the programmer
can launch at a much finer granularity with incurring the overhead of a thread.  Due to the minimal
overhead of tasks, programmers can express a much finer granularity of parallelism which the runtime
can then leverage if there are available hardware resources.  Existing programming systems like MPI
and P-threads entail too much overhead to leverage this granularity of task-parallelism.

Reviewer 1 and reviewer 5 asked for clarification regarding barriers and the need for a dynamic change
in arrival count.  In our experiencing writing programs to our runtime system, we've found that
the number of expected arrivals at a barrier is not known at barrier creation-time (intuitively this
is necessary to support nested parallelism in Legion).  To support this paradigm, we added the feature of
being able to dynamically alter the arrival count.  To implement this feature on the distributed 
machines we're targetting, we use the well-studied vector clock algorithm from the distributed systems
literature (see citation [16] for additional details on the semantics and implementation of vector clocks).

Reviewer 1 and reviwer 5 both inquired about how the runtime works with GPUs.  Our runtime uses CUDA to 
control the GPUs in the cluster.  A GPU is exposed as another kind of processor in our system, and launches
asynchronous CUDA grids for each task.  GPU memories (framebuffer and zero-copy) are exposed in the same way
as CPU main memory and the globally accessible "GASNet" memory.  Copies of data between GPU and other
memories are explicitly requested by the programmer (like all other copies), and are performed with
asynchronous DMA requests to the CUDA driver.  The CUDA driver provides device events, which are exposed
in our runtime as general events that allow GPU operations to be composeable as well.