### 2012-10-29 doc/types-pldi2013/legion-types.tex
Updating the abstract and title

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}
% preprint for dates
\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{multirow}
\usepackage{ifmtarg}
\usepackage{url}
\usepackage{bcprules}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{chngpage}

%\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}
\newcommand{\texcomment}[1]{}

\begin{document}

\title{A Type System for Structured Aliasing in Parallel Programs}
%\title{Scalable and Safe Parallelism using Logical Regions}
\authorinfo{Sean Treichler \and Michael Bauer \and Alex Aiken}{Stanford University}{\{sjt,mebauer,aiken\}@cs.stanford.edu}
%\authorinfo{Sean Treichler}{Stanford Unviersity}{sjt@cs.stanford.edu}
%\authorinfo{Michael Bauer}{Stanford University}{mebauer@cs.stanford.edu}
%\authorinfo{Alex Aiken}{Stanford University}{aiken@cs.stanford.edu}
\maketitle

\begin{abstract}
Aliasing introduces a significant impediment to reasoning about the
structure and usage of data in parallel programs.  Traditional approaches
have either disallowed aliasing to enable static analyses, or have permitted
aliasing at the cost of dynamic runtime analyses.  In this paper we introduce a
type system for the Legion programming model that uses a hybrid static and
dynamic approach to manage aliasing with minimal runtime overhead.  We prove 
that our type system can soundly reason about the privilege and coherence 
properties in Legion programs. Furthermore, we prove that type soundness 
enables a hierarchical, distributed scheduling algorithm for Legion programs 
even in the presence of aliasing. We show that an implementation of a Legion 
type checker improves the performance of Legion programs up to 71\% by 
eliding memory checks that the type system proves safe.  We also demonstrate 
that an implementation of the Legion runtime based on the hierarchical 
scheduling algorithm scales well for three real-world applications, achieving 
up to 62.5X speedup on 96 GPUs on the Keeneland supercomputer.
%Legion is a programming model that introduces
%{\em logical regions} as a mechanism for targeting large heterogeneous machines 
%with complex memory hierarchies.  Logical regions are first-class 
%values in Legion, and programmers may {\em partition} regions in multiple different
%ways into subregions.  We present
%a type system with privileges and coherence properties for reasoning about logical regions.  
%We prove the soundness of the type system and use it to show
%that all memory accesses in Legion are safe.  Furthermore, we prove that
%type soundness enables a scalable, hierarchical
%scheduling algorithm for Legion programs.  We show that an implementation of the type checker
%for Legion programs improves performance up to 71\% by eliding
%memory checks the type system proves safe.  We also demonstrate that an implementation of the Legion 
%runtime based on the hierarchical scheduling algorithm scales well for
%three real-world applications.

%Limit is 12 pages, including bibliography and appendices.
%Deadlines: \\

%\begin{tabular}{ll}
%Paper registration & July 7, 2012 3:59am PDT \\
%Paper submission & July 11, 2012 3:59am PDT \\
%Author response & September 10 - September 13, 2012 \\
%Author notification & October 1, 2012
%\end{tabular}
\end{abstract}

\input{intro}

\input{example}
%\input{outline}

\input{langdef}

\input{evaluation}

\input{related}

\input{conclusion}

{
\small
\bibliography{bibliography}
}

\end{document}