### 2012-11-02 doc/types-pldi2013/legion-types.tex

\documentclass[9pt,nocopyrightspace,preprint]{sigplanconf}
% preprint for dates
\bibliographystyle{plainnat}
\usepackage{amsmath,amsthm}
\usepackage{listings}
\usepackage{tikz}
\usepackage{subfigure}
\usepackage{multirow}
\usepackage{pgfplots}
\usepackage{booktabs}
\usepackage{dcolumn}
\usepackage{cancel}
\usepackage{multirow}
\usepackage{ifmtarg}
\usepackage{url}
\usepackage{bcprules}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{chngpage}

%\newcommand{\infrule}[2]{\displaystyle\frac{\displaystyle\strut{#1}}{\displaystyle\strut {#2}}}
\newcommand{\deref}{\ast}
\newcommand{\rread}[1]{\mbox{\em Read}(#1)}
\newcommand{\rwrite}[1]{\mbox{\em Write}(#1)}
\newcommand{\lca}[2]{#1 \sqcup #2}
\newcommand{\rleq}{\leq}
\newcommand{\interval}[1]{\mbox{\em interval}(#1)}
\newcommand{\context}[1]{\mbox{\em context}(#1)}
\newtheorem{theorem}{Theorem} 
\newtheorem{lemma}[theorem]{Lemma}
\newcommand{\texcomment}[1]{}

\begin{document}

%\title{A Type System for Structured Aliasing in Parallel Programs}
\title{Supporting Dynamic Data Partitioning in Parallel Programs}
%\title{Scalable and Safe Parallelism using Logical Regions}
\authorinfo{Sean Treichler \and Michael Bauer \and Alex Aiken}{Stanford University}{\{sjt,mebauer,aiken\}@cs.stanford.edu}
%\authorinfo{Sean Treichler}{Stanford Unviersity}{sjt@cs.stanford.edu}
%\authorinfo{Michael Bauer}{Stanford University}{mebauer@cs.stanford.edu}
%\authorinfo{Alex Aiken}{Stanford University}{aiken@cs.stanford.edu}
\maketitle

\begin{abstract}
Applications written for distributed-memory architectures must
partition their data in order to enable parallel execution.
Current language designs either analyze the partitioning of data
statically, disallowing any aliasing of data, or use expensive dynamic
checks to detect aliasing and ensure correct execution.
We describe Legion, a programming system that relies on a hybrid system
of static and dynamic analyses to support reasoning about potential aliasing introduced
by dynamic data partitions with minimal overhead.  We present the novel aspects of the Legion design, 
prove the soundness of the Legion type system, and show Legion type checking 
improves performance by up to 71\% by eliding provably safe memory checks.  
Furthermore, we show how the language design and type safety
enables a hierarchical, distributed runtime scheduling algorithm, and how
a system of data coherence annotations improves performance in the presence
of aliased data.  We report results for three real-world applications running
on distributed memory machines, achieving up to 62.5X speedup on 96 GPUs 
on the Keeneland supercomputer.
%Legion is a programming model that introduces
%{\em logical regions} as a mechanism for targeting large heterogeneous machines 
%with complex memory hierarchies.  Logical regions are first-class 
%values in Legion, and programmers may {\em partition} regions in multiple different
%ways into subregions.  We present
%a type system with privileges and coherence properties for reasoning about logical regions.  
%We prove the soundness of the type system and use it to show
%that all memory accesses in Legion are safe.  Furthermore, we prove that
%type soundness enables a scalable, hierarchical
%scheduling algorithm for Legion programs.  We show that an implementation of the type checker
%for Legion programs improves performance up to 71\% by eliding
%memory checks the type system proves safe.  We also demonstrate that an implementation of the Legion 
%runtime based on the hierarchical scheduling algorithm scales well for
%three real-world applications.

%Limit is 12 pages, including bibliography and appendices.
%Deadlines: \\

%\begin{tabular}{ll}
%Paper registration & July 7, 2012 3:59am PDT \\
%Paper submission & July 11, 2012 3:59am PDT \\
%Author response & September 10 - September 13, 2012 \\
%Author notification & October 1, 2012
%\end{tabular}
\end{abstract}

\input{intro}

\input{example}
%\input{outline}

\input{langdef}

\input{evaluation}

\input{related}

\input{conclusion}

{
\small
\bibliography{bibliography}
}

\end{document}