### 2012-11-08 doc/types-pldi2013/alt_example.tex
Minor edits

\section{Circuit Example}
\label{sec:example}

\lstset{
  captionpos=b,
  language=Haskell,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  morekeywords={function,rr,int,float,bool,isnull,partition,as,downregion,upregion,reads,writes,rdwrs,reduces,read,write,reduce,using,unpack,pack,coloring,multicoloring,color,newcolor,atomic,simultaneous},
  deletekeywords={head,min,max}
}

We begin by introducing an example program 
that motivates many of the features in Legion programming model
and provides an intuition for many of our design decisions.  The example program
executes a simulation of an arbitrary graph of electrical circuit elements.  
The simulation runs for many time steps and performs three computations 
for each time step: calculate new currents, distribute charge, and update voltages.
Due to space constraints we only introduce small code examples. 
The source code for the full circuit example written in the core Legion
language (see Section~\ref{sec:legioncore}) can be found in the
submitted appendix.

The simulation takes as input an arbitrary graph of circuit elements
that have been dynamically allocated into two logical regions: {\tt all\_nodes} 
and {\tt all\_wires}.  The types of nodes and wires are the tuples
shown in Listing~\ref{lst:tuples}.  The first two fields in the
{\tt CircuitWire} tuple are pointers that specify the two end points of
a wire.  Pointers are denoted by the @ symbol
followed by the set of region(s) into which a pointer can point.
The {\tt CircuitWire} type is parametrized by the local region
names {\tt rn} and {\tt rg} which contain {\tt CircuitNode} elements.
Our type system enforces the correctness of pointer types.

\begin{lstlisting}[label={lst:tuples},caption={Tuples and Pointers Example}]
--                        voltage,current,charge,capacitance,piece ID
type CircuitNode        = <float,float,float,float,int>
--                      owned node, owned or ghost node, resistance, current
type CircuitWire<rn,rg>  = <CircuitNode@rn, CircuitNode@(rn,rg),float,float>
\end{lstlisting}

The most important decision in any Legion program is the choice of
how to partition program data.  The choice of a partitioning algorithm
can depend on a multitude of factors including the type of data structure
and input data.  Therefore
Legion requires applications to select their own partitioning algorithms.
In this example we invoke a third-party dynamic graph partitioning library
called METIS\cite{Metis98} as shown in Listing~\ref{lst:metis}.

\begin{lstlisting}[label={lst:metis},caption={Partition Computation}]
type NodeList<rl,rn>       = < CircuitNode@rn, node_list<rl,rn>@rl >
type WireList<rl,rw,rn,rg>= < CircuitWire<rn,rg>@rw, wire_list<rl,rw,rn,rg>@rl >
function extern_metis[rl,rn,rw](node_list : NodeList<rl,rn>@rl,
          wires_list : WireList<rl,rw,rn,rn>@rl), reads(rl,rn,rw), writes(rn) : bool;
\end{lstlisting}

METIS takes as arguments lists of all the nodes and all the wires in
the graph.  It then breaks the graph into pieces by annotating each 
{\tt CircuitNode} with a piece ID (the last field in the tuple).
The function {\tt extern\_metis} is parametrized on the regions it will
access; in this case {\tt rn} the region containing nodes, {\tt rw} the region
contining wires, and {\tt rl} the region containing the lists of pointers.  The function
must also specify the {\em privileges} (e.g read) with which it will access
these regions.  {\tt extern\_metis} will read all three regions and write
the {\tt rn} region since it will be writing to the last field of each
{\tt CircuitNode}.  Our type system enforces the Legion 
requirement that each function can only be invoked with a subset of the 
privileges contained in its calling context.

\begin{figure}[t]
\centering
\subfigure[Node region tree.]{
\label{sfig:part_fig:tree}
\includegraphics[scale=0.50]{figs/OwnedTree.pdf}
}
\subfigure[Circuit piece coloring.]{
\label{sfig:part_fig:pieces}
\includegraphics[scale=0.67]{figs/Owned_Nodes.pdf}
}
\subfigure[Ghost {\tt rg0} coloring.]{
\label{sfig:part_fig:ghost_zero}
\includegraphics[scale=0.67]{figs/Ghost_RN0.pdf}
}
\subfigure[Ghost {\tt rg1} coloring.]{
\label{sfig:part_fig:ghost_one}
\includegraphics[scale=0.67]{figs/Ghost_RN1.pdf}
}
\vspace{-2mm} 
\caption{Partitions of $r{\_}all{\_}nodes$. \label{fig:part_fig}}
\vspace{-6mm}
%  \caption[My caption
\end{figure}

Figure~\ref{sfig:part_fig:pieces} shows one possible output from
METIS with different colors identifing different piece IDs.  To capture
the result of a partitioning algorithm, Legion uses {\em colorings}.
Colorings are a mapping from pointers to {\em colors} (e.g. integers).  In
the circuit example, the piece IDs are the designated colors. 
Listing~\ref{lst:coloring} shows the function {\tt create\_piece\_coloring}
that constructs a coloring from a list of nodes which have been 
annotated with piece IDs.

\begin{lstlisting}[label={lst:coloring},caption={Coloring Construction}]
function create_piece_coloring[rl,rn](node_list: NodeList<rl,rn>@rl),
            reads(rl,rn) : coloring(rn) = 
    if isnull(node_list) then newcolor(rn) else
    let list_elem : NodeList<rl,rn> = read(node_list) in
    let part_coloring : coloring(rn) = create_node_coloring(list_elem.2) in
    let node : CircuitNode = read(list_elem.1) in
        color(part_coloring, list_elem.1, node.5)
\end{lstlisting}

Colorings are the input for partition operations in Legion.  The partition
statement shown in Listing~\ref{lst:partition} partitions the region
{\tt all\_nodes} into logical subregions with a separate output region
for each color in the coloring.  Due to space constraints we, only show
code for {\tt rn0} and {\tt rn1} pieces for the remainder of the example.  We also
do not show the coloring or partitioning of the {\tt all\_wires} region as
it can be computed in a straightforward manner from the partitioning of
nodes into pieces.

\begin{lstlisting}[label={lst:partition},caption={Partition Operation Example}]
let piece_coloring : coloring(all_nodes) = create_piece_coloring[rl,all_nodes](node_list) in
partition all_nodes using piece_coloring as rn0,rn1... in ...
\end{lstlisting}

Since colorings can only color each pointer at most with a single color
(dynamically checked), the partition operation results in 
{\em disjoint} subregions.  Furthermore, the partition operation
introduces constraints into the static checking enviornment describing
both the disjointness of the subregions (i.e. {\tt rn0 $*$ rn1})
and the inclusion of each subregion in the original region
(i.e. {\tt rn0 $\leq$ all\_nodes}).

In addition, to partitioning the graph to describe {\em owned} nodes
for each piece, the circuit example also
needs to be able to name the {\em ghost nodes} for each piece.  Ghost node
regions are used for accessing the nodes of wires that span
two different pieces. (This is also the reason {\tt CircuitWire} is parametrized on
two regions).  Figure~\ref{sfig:part_fig:ghost_zero} and 
Figure~\ref{sfig:part_fig:ghost_one} show colorings for the ghost regions
corresponding to {\tt rn0} and {\tt rn1} respectively.  Note that two 
nodes are colored in both colorings.  Since nodes can appear in multiple
subregions, we use a {\em multicoloring} (implemented in core Legion
by multiple single colorings) to describe {\em aliased} partitions.  Aliased
partitions still introduce subregion constraints into the static checking
environment, but not disjointness constraints.  Figure~\ref{sfig:part_fig:tree}
shows the resulting logical region tree for node regions.

The first-class nature of logical regions in Legion allows them to
be written into the heap.  To avoid losing important static properties of
logical regions in this scenario, we use {\em region
relationships}.  Region relationships are a bounded existential type
that allows programmers to {\em pack} a group of regions together and remember
properties about them such as disjointness and subregion relationships.
Listing~\ref{lst:rr} shows an example region relationship for a {\tt CircuitPiece} 
describing its region of wires {\tt rpw}, region of 
nodes {\tt rpn}, region of ghost nodes {\tt rg}, and important constraints.

\begin{lstlisting}[label={lst:rr},caption={Region Relationship Example}]
type CircuitPiece<rl,rw,rn> = rr[rpw,rpn,rg]
                            < WireList<rl,rpw,rpn,rg>@rl, NodeList<rl,rpn>@rl >         
                            where rpn #$\le$# rn and rg #$\le$# rn and rpw #$\le$# rw and
                                  rn * rw and rl * rn and rl * rw
\end{lstlisting}

The Legion type system verifies properties of region relationships when packing;
when region relationships are unpacked the same properties are
reintroduced into the static checking environment.
A key feature underlying the soundness of the Legion
type system is that privileges cannot be packed in a region
relationship---privileges belong to functions. When a function unpacks
a region relationship it must already hold privileges for the unpacked regions.
Listing~\ref{lst:loop} shows the main time step loop run by the 
recursive function {\em execute\_time\_steps} (line 11)
as well the leaf task functions (lines 2-7).  
For each time step, the loop unpacks 
two previously packed circuit pieces, giving new names to the subregions introduced
by each region relationship.  The {\em execute\_time\_steps} function
will have read/write privileges for the newly named regions, such as {\tt rn0},
because it has read/write privileges for {\tt rn} and the {\tt CircuitPiece} 
region relationship declares {\tt rn0 $\leq$ rn}.  Note this does require
a dynamic check that {\tt rn0} is a subregion of {\tt rn}, but the cost of the 
check will be ammortized over all the uses of {\tt rn0}.

The {\tt execute\_time\_steps} function   
illustrates the importance of having different partitions provide 
multiple views onto the same logical region.  Both {\tt calc\_new\_currents} 
and {\tt distribute\_charge} (lines 2-5)
use the owned and ghost regions of a piece, which are from different partitions. In
{\tt calc\_new\_currents} these regions only need read
privileges, allowing both instances of the task to be
run in parallel.  In {\tt distribute\_charge} the
privilege is for a reduction which can also be done in parallel
because of the atomic and commutative nature of reductions.  Finally,
the {\tt update\_voltage} function (line 6) modifies only the disjoint owned regions, 
permitting each piece to be updated in parallel.  No single partition of the nodes
describes these data sharing patterns.

\lstset{
  captionpos=b,
  language=Haskell,
  basicstyle=\scriptsize,
  numbers=left,
  numberstyle=\tiny,
  columns=fullflexible,
  stepnumber=1,
  escapechar=\#,
  keepspaces=true,
  belowskip=-10pt,
  literate={<}{{$\langle$}}1 {>}{{$\rangle$}}1,
  morekeywords={function,rr,int,float,bool,isnull,partition,as,downregion,upregion,reads,writes,rdwrs,reduces,read,write,reduce,using,unpack,pack,coloring,multicoloring,color,newcolor,atomic,simultaneous},
  deletekeywords={head,min,max}
}
\begin{lstlisting}[float={t},label={lst:loop},caption={Main Simulation Loop}]
-- Leaf Task Declarations (implementations in appendix)
function calc_new_currents[rl,rw,rn,rg] ( ptr_list : WireList<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn,rg), writes(rw) : bool
function distribute_charge[rl,rw,rn,rg] ( ptr_list : WireList<rl,rw,rn,rg>@rl ), 
      reads(rl,rw,rn), reduces(reduce_charge,rn,rg), atomic(rn,rg) : bool 
function update_voltage[rl,rn] ( ptr_list : NodeList<rl,rn>@rl ), 
      reads(rl,rn), writes(rn) : bool
-- Reduction function for distribute charge
function reduce_charge ( node : CircuitNode, current : float ) : CircuitNode

-- Time Step Loop
function execute_time_steps[rl,rw,rn] ( p0 : CircuitPiece<rl,rw,rn>, 
      p1 : CircuitPiece<rl,rw,rn>, steps : int ) , reads(rn,rw,rl), writes(rn,rw) : bool = 
  if steps #$<$# 1 then true else
  unpack p0 as piece0 : CircuitPiece<rl,rw,rn>[rw0,rn0,rg0] in 
  unpack p1 as piece1 : CircuitPiece<rl,rw,rn>[rw1,rn1,rg1] in
  let _ : bool = calc_new_currents[rl,rw0,rn0,rg0](piece0.1) in
  let _ : bool = calc_new_currents[rl,rw1,rn1,rg1](piece1.1) in
  let _ : bool = distribute_charge[rl,rw0,rn0,rg0](piece0.1) in
  let _ : bool = distribute_charge[rl,rw1,rn1,rg1](piece1.1) in
  let _ : bool = update_voltage[rl,rn0](piece0.2) in
  let _ : bool = update_voltage[rl,rn1](piece1.2) in
      execute_time_steps[rl,rw,rn](p0,p1,steps-1)
\end{lstlisting}

The {\tt execute\_time\_steps} function also demonstrates the need to 
dynamically discover parallelism through region disjointess.  
To run the two instances of 
{\tt calc\_new\_currents} in parallel requires knowing that {\tt rw0}
and {\tt rw1} are disjoint (lines 17-18).  There is a similar requirement for parallel
execution of the two {\tt update\_voltage} calls with {\tt rn0} and {\tt rn1} (lines 21-22).  In
both cases, this knowledge was statically available when the subregions were created 
and could have been captured in a region relationship at the cost
of much more complicated code.  Instead, we use a simpler region relationship
and rely on an inexpensive dynamic disjointness
check by the Legion runtime to (re)discover the parallelism.

In addition to privileges, functions can also specify {\em coherence}
on regions.  By default, coherence on all regions is {\em exclusive}, meaning
the function must appear to execute in program order relative to other function
calls that access non-disjoint data.  Programmers can allow other
execution orderings using relaxed coherence.  Line 5 of Listing~\ref{lst:loop} shows
an example of {\tt atomic}, a relaxed coherence mode requiring
that operations to {\tt rn} and {\tt rg} appear atomic relative
to other functions using regions which may alias.  The most relaxed coherence
mode is {\tt simult}; simultaneous coherence allows concurrent access 
to the region by all functions that are using the region
in a simultaneous mode.  The interaction between {\em tasks} (which are functions 
chosen to execute in parallel) using the same region with different coherence
modes is formalized in Section~\ref{sec:coherence}.

%The most important decision in any Legion program is the choice of
%how to partition program data into logical regions.  Logical regions
%are the granularity at which the Legion runtime will reason
%about the independence of tasks, and therefore how much parallelism
%is available.  Logical regions also define the granularity at which
%the Legion runtime will move data.  Coarser regions will lead
%to more efficient bulk copies of data throgh the memory hierarchy due
%to better locality.  There is therefore a natural tension: partitioning 
%should be fine-grained enough to express desired parallelism, but not 
%so fine-grained as to result in inefficient data movement.

%The first step in the circuit example is to partition the arbitrary graph
%of circuit elements into pieces that can be handled independently.
%In the case of this example all nodes in the graph have initially
%been allocated in the region {\tt all\_nodes} (Figure~\label{sfig:part_fig:tree}).
%The decision of how to partition the graph is the responsibility
%of the programmer and can be done by an external library (e.g. METIS
%in this case).  Once the decision of how to partition is made, the
%result must be communicated to the Legion runtime to 
%generate the corresponding logical regions.  Partitioning decisions
%are described via {\em colorings}.  Colorings are parametrized by
%the region they are partitioning and capture a mapping from pointers
%into the region to specific colors (e.g. integers).  Figure~\ref{sfig:part_fig:pvs}
%illustrates the coloring used for generating the {\tt p\_nodes\_pvs}
%partition.